### 2016-03-29 doc/thesis/preamble.tex
Writing stuff

\documentclass[12pt]{report}
\usepackage[osf]{garamondx}
\usepackage[garamondx,cmbraces]{newtxmath}
\usepackage{marginnote}
\usepackage{fullpage}
\usepackage{setspace}
\usepackage[backend=biber, style=numeric-comp, sorting=none]{biblatex}
\usepackage{nicefrac}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{multicol}
\usepackage{multirow}
\usepackage[hidelinks]{hyperref}
\usepackage[acronym,nonumberlist]{glossaries}
\usepackage{glossary-mcols}
\usepackage{cleveref}
\usepackage{array}
\usepackage{textcomp}
\usepackage{float}

\newcommand{\comment}[1]{(\textbf{#1})}

\newcommand{\defn}[1]{\textit{#1}}
\newcommand{\software}[1]{\textit{#1}}
\newcommand{\etal}{\textit{et al.}}
\newcommand{\eg}{\textit{e.g.}}

\newcommand{\set}[1]{\left\lbrace#1\right\rbrace}
\newcommand{\sett}[1]{\{#1\}}
\renewcommand{\emptyset}{\varnothing}
\renewcommand{\vec}[1]{\mathbf{#1}}
\renewcommand{\star}[1]{#1^*}
\renewcommand{\d}{\mathrm{d}\,}

\newcommand{\N}{\mathcal{N}}
\renewcommand{\L}{\mathcal{L}}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

\DeclareMathOperator{\Exponential}{Exponential}
\DeclareMathOperator{\Uniform}{Uniform}
\DeclareMathOperator{\E}{E}
\DeclareMathOperator{\Var}{Var}

\DeclareMathOperator{\inc}{in}
\DeclareMathOperator{\out}{out}

\newfloat{data}{hb}{lod}
\floatname{data}{Data}
\crefname{data}{data}{data}

\newacronym{MCMC}{MCMC}{Markov chain Monte Carlo}
\newacronym{ER}{ER}{Erd\H{o}s-R\'enyi}
\newacronym{BA}{BA}{Barab\'asi-Albert}
\newacronym{WS}{WS}{Watts-Strogatz}
\newacronym{SI}{SI}{susceptible-infected}
\newacronym{SVM}{SVM}{support vector machine}
\newacronym{PCA}{PCA}{principal components analysis}
\newacronym{SMC}{SMC}{sequential Monte Carlo}
\newacronym{ABC}{ABC}{approximate Bayesian computation}
\newacronym{nltt}{nLTT}{normalized lineages-through-time}
\newacronym{ltt}{LTT}{lineages-through-time}
\newacronym{pdf}{pdf}{probability density function}
\newacronym{rv}{rv}{random variable}
\newacronym{ML}{ML}{maximum likelihood}
\newacronym{MAP}{MAP}{maximum \textit{a posteriori}}
\newacronym{GSL}{GSL}{GNU scientific library}
\newacronym{SIR}{SIR}{susceptible-infected-recovered}
\newacronym{IS}{IS}{importance sampling}
\newacronym{SIS}{SIS}{sequential importance sampling}
\newacronym{HMM}{HMM}{hidden Markov model}
\newacronym{iid}{i.i.d.}{independent and identically distributed}
\newacronym{HPD}{HPD}{highest posterior density}
\newacronym{HIV}{HIV}{human immunodeficiency virus}
\newacronym{SARS}{SARS}{severe acute respiratory syndrome}
\newacronym{ODE}{ODE}{ordinary differential equation}
\newacronym{GLM}{GLM}{generalized linear model}

\newglossaryentry{alpha}{
  name=$\alpha$,
  description=preferential attachment power parameter in Barab\'asi-Albert networks
}
\newglossaryentry{m}{
  name=$m$,
  description=number of edges added per vertex when constructing a Barab\'asi-Albert network
}
\newglossaryentry{N}{
  name=$N$,
  description=number of nodes in the network
}
\newglossaryentry{I}{
  name=$I$,
  description=number of nodes which are eventually infected
}
\newglossaryentry{gamma}{
  name=$\gamma$,
  description=exponent of power-law degree distribution in scale-free networks
}
\newglossaryentry{p}{
  name=$p$,
  description=edge density parameter in Erd\H{o}s-R\'enyi networks
}
\newglossaryentry{lambda}{
  name=$\lambda$,
  description=decay factor meta-parameter for tree kernel
}
\newglossaryentry{sigma}{
  name=$\sigma$,
  description=radial basis function variance meta-parameter for tree kernel
}
\newglossaryentry{n}{
  name=$n$,
  description=number of particles used for ABC-SMC
}
\newglossaryentry{M}{
  name=$M$,
  description=number of simulated datasets per particle in ABC-SMC
}
\newglossaryentry{tmrca}{
  name=$t_{\mathrm{mrca}}$,
  description=time to most recent common ancestor of a group of extant taxa
}

\graphicspath{{../figures/}}
\newcommand{\tablepath}{../tables}
\frenchspacing
\onehalfspacing

\title{Phylodynamic inference of contact network parameters through approximate
Bayesian computation}
\author{Rosemary M. McCloskey and Art F.Y. Poon}

\addbibresource{papers.bib}
\makeglossaries