### 2016-04-16 doc/thesis/appendix.tex
More writing yay

\begin{figure}
  \centering
  \includegraphics{kernel-I-tree.pdf}
  \caption{
    Transmission trees simulated over \gls{BA} networks with varying values of
    $I$, the number of infected nodes when the epidemic simulation was stopped.
  }
  \label{fig:Itrees}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{kernel-m-tree.pdf}
  \caption{
    Transmission trees simulated over \gls{BA} networks with varying values of
    $m$, the number of edges added per vertex.
  }
  \label{fig:mtrees}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{kernel-N-tree.pdf}
  \caption{
    Transmission trees simulated over \gls{BA} networks with varying values of
    $N$, the number of nodes in the network.
  }
  \label{fig:Ntrees}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{kernel-alpha-crossv.pdf}
  \label{fig:alphacrossv}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{kernel-I-crossv.pdf}
  \label{fig:Icrossv}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{kernel-m-crossv.pdf}
  \label{fig:mcrossv}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{kernel-N-crossv.pdf}
  \label{fig:Ncrossv}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{kernel-alpha-kpca.pdf}
  \label{fig:alphakpca}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{kernel-I-kpca.pdf}
  \label{fig:Ikpca}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{kernel-m-kpca.pdf}
  \label{fig:mkpca}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{kernel-N-kpca.pdf}
  \label{fig:Nkpca}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{gridsearch-alpha-kernel.pdf}
  \caption{                                              
    Grid search kernel scores for testing trees simulated under various             
    \gls{alpha} values. The other \gls{BA} parameters were fixed at \gls{I} =
    1000, \gls{N} = 5000, and \gls{m} = 2. 
  }        
  \label{fig:gridalpha}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{gridsearch-I-kernel.pdf}
  \caption{                                              
    Grid search kernel scores for testing trees simulated under various             
    \gls{I} values. The other \gls{BA} parameters were fixed at \gls{alpha} =
    1.0, \gls{N} = 5000, and \gls{m} = 2. 
  }        
  \label{fig:gridI}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{gridsearch-m-kernel.pdf}
  \caption{                                              
    Grid search kernel scores for testing trees simulated under various             
    \gls{m} values. The other \gls{BA} parameters were fixed at \gls{alpha} =
    1.0, \gls{I} = 1000, and \gls{N} = 5000.
  }        
  \label{fig:gridm}
\end{figure}

\begin{figure}
  \centering
  \includegraphics{gridsearch-N-kernel.pdf}
  \caption{                                              
    Grid search kernel scores for testing trees simulated under various             
    \gls{N} values. The other \gls{BA} parameters were fixed at \gls{alpha} =
    1.0, \gls{I} = 1000, and \gls{m} = 2. 
  }        
  \label{fig:gridN}
\end{figure}