### 2016-04-18 doc/thesis/intro.tex
More writing

\section{Phylogenetics and phylodynamics}
\label{sec:phylo}

\subsection{Phylogenetic trees}

In evolutionary biology, a \defn{phylogeny}, or \defn{phylogenetic tree}, is a
graphical representation of the the evolutionary relationships among a group of
organisms or species (generally, \defn{taxa})~\autocite{haeckel1866generelle}.
The \defn{tips} of a phylogeny, that is, the nodes without any descendants,
correspond to \defn{extant}, or observed, taxa. The \defn{internal nodes}
correspond to their (usually extinct) common ancestors. The edges or
\defn{branches} of the phylogeny connect ancestors to their descendants.
Phylogenies may have a \defn{root}, which is a node with no descendants
distinguished as the most recent common ancestor of all the extant
taxa~\autocite{harding1971probabilities}. When such a root exists, the tree is
referred to as being \defn{rooted}; otherwise, it is \defn{unrooted}. The
structural arrangement of nodes and edges in the tree is referred to as its
\defn{topology}~\autocite{cavalli1967phylogenetic}.

The branches of the tree may have associated lengths, representing either
evolutionary distance or calendar time between ancestors and their descendants.
The term ``evolutionary distance'' is used here imprecisely to mean any sort of
quantitative measure of evolution, such as the number of differences between
the DNA sequences of an ancestor its descendant, or the difference in average
body mass or height. A phylogeny with branch lengths in calendar time units is
often referred to as \defn{time-scaled}. In a time-scaled phylogeny, the
internal nodes can be mapped onto a timeline by using the tips of the tree,
which usually correspond to the present day, as a reference
point~\autocite{nee1992tempo}. The corresponding points on the timeline are
called \defn{branching times}, and the rate of their accumulation is referred
to as the \defn{branching rate}. Rooted trees whose tips are all the same
distance from the root are called \defn{ultrametric}
trees~\autocite{buneman1974note}. These concepts are illustrated in
\cref{fig:speciestree}.

\begin{figure}[ht]
  \centering
  \includegraphics{speciestree.pdf}
  \caption[Illustration of a rooted, ultrametric, time-scaled phylogeny]
    {Illustration of a rooted, ultrametric, time-scaled phylogeny. The tips of
      the tree, which represent extant taxa, are placed at the present day on
      the time axis. Internal nodes, representing extinct common ancestors to
      the extant taxa, fall in the past. The topology of the tree indicates
      that cats and dogs are the most closely related pair of species, whereas
      fish is most distantly related to any other node in the tree.}
  \label{fig:speciestree}
\end{figure}

\subsection{Transmission trees}

In epidemiology, a \defn{transmission tree} is a graphical representation of an
epidemic's progress through a population. Like phylogenies, transmission trees
have tips, nodes, edges, and branch lengths. However, rather than recording an
evolutionary process (speciation), they record an epidemiological process
(transmission). The tips of a transmission tree represent the removal of
infected hosts, while internal nodes correspond to transmissions from one host
to another.Transmission trees generally have branch lengths in units of
calendar time, with branching times indicating times of transmission. The root
of a transmission tree corresponds to the initially infected patient who
introduced the epidemic into the network, also known as the \defn{index case}.
The internal nodes may be labelled with the donor of the transmission pair, if
this is known. The tips of the tree, rather than being fixed at the present
day, are placed at the time at which the individual was removed from the
epidemic, such as by death, recovery, isolation, behaviour change, or
migration. Consequently, the transmission tree may not be ultrametric, but may
have tips located at varying distances from the root. Such trees are said to
have \defn{heterochronous} taxa~\autocite{drummond2003measurably}, in contrast
to the \defn{isochronous} taxa found in most phylogenies of macro-organisms. A
transmission tree is illustrated in \cref{fig:contactnet} (right).

Each infected individual in an epidemic may appear in the transmission tree
more than once. This is different from the transmission \emph{network}, in
which each infected individual appears exactly once, and edges are in
one-to-one correspondence with transmissions~\autocite{welch2011statistical,
keeling2005networks}. Transmission networks are discussed further in
\cref{sec:contactnet}, and the distinction between the two objects is
illustrated in \cref{fig:contactnet}. However, since transmission networks
generally have no cycles (unless re-infection occurs), they are trees in the
graph theoretical sense, and hence are sometimes also referred to as
transmission trees~\autocite[\eg][]{kenah2015algorithms}. In this work, we
reserve the term ``transmission tree'' for the objects depicted on the right
side of \cref{fig:contactnet}, following \eg~\autocite{stadler2013uncovering}.
The term ``transmission network'' is taken to mean the subgraph of the contact
network along which transmissions occurred, following
\eg~\autocite{welch2011statistical, keeling2005networks}.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{contactnet.pdf}
    \caption[Illustration of a contact network and transmission tree]{
      Illustration of epidemic spread over a contact network, and the
      corresponding transmission tree. (Left) A contact network with five
      hosts, labelled $a$ through $e$. Thick shaded edges indicate symmetric
      contacts among the hosts. The transmission network is indicated by
      coloured arrows. The epidemic began with node $a$, who transmitted to
      nodes $b$ and $c$. Node $c$ further transmitted to node $d$. Node $e$ was
      not infected. (Right) The transmission tree corresponding to this
      scenario, with a timeline of transmission and removal times.
    }
    \label{fig:contactnet}
\end{figure}

Since transmission trees are essentially a detailed record of an epidemic's
progress, they contain substantial epidemiological information. As a basic
example, the \gls{ltt} plot~\autocite{nee1992tempo}, which plots the number of
lineages in a phylogeny against time, can be used to quantify the incidence of
new infections over the course of an epidemic~\autocite{holmes1995revealing}.
However, in all but the most well-studied of epidemics, transmission trees are
not possible to assemble through traditional epidemiological
methods~\autocite{welch2011statistical}. The time and effort to conduct
detailed interviews and contact tracing of a sufficient number of infected
individuals is usually prohibitive, and may be additionally be confounded by
misreporting~\autocite{eames2015six}. However, it turns out that for viral
epidemics, some of the epidemiological information contained in the
transmission tree leaves a mark on the viral genetic material circulating in
the population. A family of methods called
\defn{phylodynamics}~\autocite{grenfell2004unifying} addresses the challenge of
estimating epidemiological parameters from viral sequence
data~\autocite{volz2013viral}.

\subsection{Phylodynamics: linking evolution and epidemiology}

%Most modern analyses use model-based methods, which simultaneously
%estimate the phylogeny with branch lengths and the parameters of a model of
%evolution. Although they usually work well in practice, the estimated topology
%can vary based on the model used and, in the case of Bayesian analysis, the
%priors. In addition, intra-host viral populations are genetically
%heterogeneous, so choosing a single representative genotype per host is
%necessarily imprecise. One can use either the genotype of a specific virion
%sampled from the host, or a synthetic genotype, such as a consensus or
%reconstructed ancestral sequence.

In general, \defn{viral phylogenies} are simply phylogenetic trees relating
virus strains. \defn{Inter-host} phylogenies relate one representative viral
genotype from each host in a population. In this work, when we use the phrase
``viral phylogeny'', we are referring to an inter-host phylogeny unless
otherwise stated. Phylogenetic inference is a complex topic which we will not
discuss here (see \eg \autocite{nei2000molecular} for a full review). We assume
here that the viral phylogeny is known, a common assumption in the development
and application of phylodynamic methods.

The crux of phylodynamics is the fact that the epidemiological processes
recorded in transmission trees, and the evolutionary processes recorded in
viral phylogenies, occur on similar time scales for RNA
viruses~\autocite{drummond2003measurably}. Phylodynamic methods have been used
to investigate epidemiological parameters such as transmission rate, recovery
rate, and basic reproductive number~\autocite{pybus2009evolutionary,
volz2013viral}. The majority of phylodynamic studies attempt to infer the
parameters of an epidemiological model for which the likelihood of an observed
phylogeny can be calculated. Most often, this is some variation of the
birth-death~\autocite{kendall1948generalized, stadler2012estimating} or
coalescent~\autocite{kingman1982coalescent, volz2012complex} models.

Due to the relationship between the aforementioned processes, there is a close
relationship between viral phylogenies and transmission trees. In particular,
the transmission process is quite similar to \defn{allopatric
speciation}~\autocite{coyne2004speciation}, where genetic divergence follows
the geographic isolation of a sub-population of organisms. Thus, transmission,
which is represented as branching in the transmission tree, causes branching in
the viral phylogeny as well. Similarly, the removal of an individual from the
transmission tree causes the extinction of their viral lineage in the
phylogeny. Due to these relationships, the topology of the viral phylogeny is
often used as a proxy for the topology of the transmission tree. However, there
are several complications and caveats which must be kept in mind when
estimating the transmission tree in this manner.

First are the issues of rooting and time-scaling. Modern likelihood-based
methods of phylogenetic reconstruction~\autocite[\eg][]{price2010fasttree,
stamatakis2014raxml} produce unrooted trees whose branch lengths measure
genetic distance in units of expected substitutions per site. On the other
hand, transmission trees are rooted, and have branches measuring calendar
time~\autocite{pybus2009evolutionary}. It is generally assumed that sampling a
virus from the individual also corresponds to their removal from the
transmission tree, so the positions of the tips in time are fixed. Therefore,
to transform a viral phylogeny into an estimated transmission tree, we must
find a root and an assignment of branch lengths such that the tips are placed
at their proscribed times. In addition, the branch lengths should be chosen
such that the variation in \defn{evolutionary rate}, which is the ratio of a
branch's length in genetic distance to its length in calendar time, is low
across the tree. While there is some variation among hosts, due to
immunological and other factors, we generally expect to observe globally
similar evolutionary rates. Methods for time-scaling a phylogeny include
root-to-tip regression~\autocite{shankarappa1999consistent, korber2000timing,
drummond2003inference}, which we apply in this work, and least-square
dating~\autocite{to2015fast}. Both of these methods can be used to root the
tree, by simply trying all possible root positions and choosing the one which
minimizes a loss function ($1-R^2$, or root-mean-square error, for root-to-tip
regression; sum of squared errors for least-square dating). Alternatively, the
tree may be rooted separately with an outgroup~\autocite{li1988rates} before
time-scaling.

A second, perhaps more insidious problem is the fact that the correspondence
between the topologies of the viral phylogeny and transmission tree is not
necessarily exact. Due to intra-host diversity, the viral strain which is
transmitted may have split from another lineage within the donor long before
the transmission event occurred. Hence, the branching point in the viral
phylogeny may be much earlier than that in the transmission tree. Another
possibility is that one host transmitted to two or more recipients, but the
lineages they each received originated within the donor host in a different
order than that in which the transmissions occurred. In this case, the topology
of the transmission tree and the viral phylogeny will be mismatched. Although
phylodynamics is quite new, these phenomena have been studied in evolutionary
biology for some time. Viral phylogenies are a specific version of a more
general class of trees called \defn{gene trees}, which represent the
evolutionary history of a section of genetic material. Transmission trees, on
the other hand, are highly analogous to \defn{species trees}, whose tips are
species and internal nodes are common ancestors. This analogy derives from the
functional similarity between transmission and allopatric speciation. Hence,
the potential discordance between transmission trees and viral phylogenies is
the similar to that between gene and species trees, which is called
\defn{incomplete lineage sorting}. In practice, this discordance has not proven
an insurmountable problem: for example, \textcite{leitner1996accurate} were
able to accurately recover a known transmission tree using a viral phylogeny.

What we have just discussed is a two-step procedure for estimating the
transmission tree. First, a viral phylogeny is constructed from genetic
sequence data, and then it is rooted and time-scaled into a transmission tree.
This approach is straightforward, frequently used, and has the advantage of
leveraging tried-and-true tools for phylogenetic inference. However, it also
has drawbacks, perhaps the most obvious being the multiplication of errors
produced by the separate steps. One commonly used alternative method is to
directly estimate a time-scaled phylogeny by simultaneously inferring the tree
topology, its root and branch lengths, and the parameters of a \defn{molecular
clock} model. A molecular clock is a hypothesis about the evolutionary rates
along the branches of the tree, such as that they are all equal (a \defn{strict
clock}) or that they are \gls{iid} from a common distribution (a \defn{relaxed
clock}). This inference is usually done in a Bayesian framework using
\gls{MCMC}, so that prior information (including the tip dates) can be included
in the analysis, and the so-called nuisance parameters of the molecular clock
model can be marginalized out. Software packages for performing these analyses
include BEAST~\autocite{bouckaert2014beast} and
MrBayes~\autocite{ronquist2012mrbayes}.

Several other authors have developed methods tailor-made for inferring
transmission trees. \textcite{didelot2014bayesian} develop a Bayesian version
of the two-step approach which allows transmissions to occur anywhere along the
branches of a transmission tree, rather than being constrained to the branching
points in the viral phylogeny. The method requires sampling of every infected
individual, although the authors indicate that it could be extended to relax
this assumption. \textcite{cottam2008integrating} describe a likelihood-based
method which enumerates all transmission trees consistent with an established
phylogeny, assigning each a likelihood based on other epidemiological data. 
This approach is novel in its integration of data from multiple sources,
however because it enumerates a large portion of the tree space, it is unlikely
to scale to larger epidemics. \textcite{ypma2012unravelling} develop a joint 
likelihood function integrating temporal, geographic, and genetic observations, 
and use Bayesian \gls{MCMC} to estimate both the tree and the parameters of the
likelihood function. Their approach can handle missing data and produces high
resolution transmission trees when multiple types of data are available. A
different approach is undertaken by \textcite{jombart2011reconstructing}, who
describe a method to build transmission trees directly from sequence data,
contingent on the common ancestors also being sampled. This makes the method
attractive for slow-evolving pathogens, but less practical for viral outbreaks
where samples from common ancestors are unlikely to be available.

\subsection{Tree shapes}
\label{subsec:treeshape}

The aim of viral phylodynamics is to glean some kind of knowledge, about the
epidemic, the virus, or its hosts and their behaviour, by studying a phylogeny,
most often an estimated transmission tree~\autocite{pybus2009evolutionary,
volz2013viral}. What is informative about a phylogeny, beyond the demographic
characteristics of the individuals it relates, is its \defn{shape}. The shape
of a phylogeny has two components: the topology, and the distribution of branch
lengths~\autocite{mooers1997inferring}. Methods of quantifying tree shape fall
into two categories: summary statistics, and pairwise measures.

Summary statistics assign a numeric value to each individual tree. One of the
most widely used is Sackin's index~\autocite{shao1990tree}, which measures the
imbalance or asymmetry in a rooted tree. For the $i$th tip of the tree, we
define $N_i$ to be the number of branches between that tip and the root. The
unnormalized Sackin's index is defined as the sum of all $N_i$. It is called
unnormalized because it does not account for the number of tips in the tree.
Among two trees having the same number of tips, the least-balanced tree will
have the highest Sackin's index. However, among two equally balanced trees, the
larger tree will have a higher Sackin's index. This makes it challenging to
compare balances among trees of different sizes. To correct this,
\textcite{kirkpatrick1993searching} derive the expected value of Sackin' index
under the Yule model~\autocite{yule1925mathematical}. Dividing by this expected
value normalizes Sackin's index, so that it can be used to compare trees of
different sizes.

Rather than assigning numbers to individual trees, pairwise measures associate
a numeric value to each \textit{pair} of trees, indicating how different the
trees are from each other. Distance measures allow us to identify groups of
related phylogenies, for example, local epidemics which are undergoing a
similar pattern of expansion. One such distance measure is the
\gls{nltt}~\autocite{janzen2015approximate}, which compares the
\gls{ltt}~\autocite{nee1992tempo} plots of two trees. Specifically, the two
\gls{ltt} plots are normalized so that they begin at $(0, 0)$ and end at $(1,
1)$, and the difference between the two plots is integrated between 0 and 1. In
the context of infectious diseases, the \gls{ltt} is related to the
prevalence~\autocite{holmes1995revealing}, so large values may indicate that
the trees being compared are the products of different epidemic
trajectories~\autocite{janzen2015approximate}.

\textcite{poon2013mapping} developed an alternative pairwise measure which
applies the concept of a \defn{kernel function} to phylogenies. Kernel
functions, originally developed for \glspl{SVM}~\autocite{burges1998tutorial}, 
compare objects by mapping them into a feature space of very high, possibly
even infinite, dimension.

Another tree distance measure is the \defn{phylogenetic kernel}, or ``tree
kernel'' developed by \textcite{poon2013mapping}. As opposed to the \gls{nltt},
the tree kernel is maximized when the two trees being compared are the same.
The basis of the tree kernel is the kernel trick originally developed for
\glspl{SVM}~\autocite{burges1998tutorial}. The idea of the kernel trick is to
compare objects by mapping them into a feature space of very high, possibly
even infinite, dimension. The similarity between objects is taken to be their
dot product in the feature space. It is called a ``trick'' because this dot
product is computed using a \defn{kernel function} without explicitly mapping
the objects to the feature space, which would be computationally prohibitive.
In the case of the tree kernel, the feature space is the space of all possible
\defn{subset trees}, which are subtrees that do not necessarily extend all the
way to the tips. The subset-tree kernel was originally developed for comparing
parse trees in natural language processing~\autocite{collins2002new} and did
not incorporate branch length information. The version developed by
\textcite{poon2013mapping} includes a radial basis function to compare the
differences in branch lengths, thus incorporating both the trees' topologies
and their branch lengths in a single similarity score. The tree kernel was
later shown to be highly effective in differentiating trees simulated under a
compartmental model with two risk groups of varying contact
rates~\autocite{poon2015phylodynamic}. In that paper,
\citeauthor{poon2015phylodynamic} used the tree kernel as the distance function
in \gls{ABC} (see \cref{sec:abc}), to fit epidemiological models to observed
trees. This is an example of kernel-ABC~\autocite{nakagome2013kernel}, which
will be discussed further in \cref{sec:abc}.

\section{Contact networks}
\label{sec:contactnet}

\subsection{Overview}
\label{subsec:netoverview}

Epidemics spread through populations of hosts through \defn{contacts} between
those hosts. The definition of contact depends on the mode of transmission of
the pathogen in question. For an airborne pathogen like influenza, a contact
may be simple physical proximity, while for \gls{HIV}, contact could be via
unprotected sexual relations or blood-to-blood contact (such as through needle
sharing). A \defn{contact network} is a graphical representation of a host
population and the contacts among its members~\autocite{klovdahl1985social,
morris1993epidemiology, keeling2005networks}. The \defn{nodes} in the network
represent hosts, and \defn{edges} or \defn{links} represent contacts between
them. A contact network is shown in \cref{fig:contactnet} (left). Contact
networks are a particular type of \defn{social
network}~\autocite{moreno1953shall, barnes1954class}, which is a network in
which edges may represent any kind of social or economic relationship. Social
networks are frequently used in the social sciences to study phenomena where
relationships between people or entities are important \autocite[for a review
see][]{wasserman1994social}.

Edges in a contact networks may be \defn{directed}, representing one-way
transmission risk, or \defn{undirected}, representing symmetric transmission
risk. For example, a network for an airborne epidemic would use undirected
edges, because the same physical proximity is required for a host to infect or
to become infected. However, an infection which may be spread through
blood-to-blood contact through transfusions transfusions would use directed
edges, since the donor has no chance of transmitting to the recipient. Directed
edges are also useful when the transmission risk is not equal between the
hosts, such as with HIV transmission, where acting as the receptive partner
carries a higher risk of infection than acting as the insertive partner. In
this case, a contact could be represented by two directed edges, one in each
direction between the two hosts, with the edges annotated by what kind of risk
they imply~\autocite{wasserman1994social}. An undirected contact network is
equivalent to a directed network where each contact is represented by two
symmetric directed edges. The \defn{degree} of a node in the network is how
many contacts it has. In directed networks, we may make the distinction between
\defn{out-degree} and \defn{in-degree}, which count respectively the number
incoming and outgoing edges. The \defn{degree distribution} of a network
denotes the probability that a node has any given number of links. The set of
edges attached to a node are referred to as its \defn{incident} edges.

Epidemiological models most often assume some form of contact homogeneity. The
simplest models, such as the \gls{SIR} model, assume a completely homogeneously
mixed population, where every pair of contacts is equally likely. More
sophisticated models partition the population into groups with different
contact rates between and among each group. However, these models still assume
that every possible contact between a member of group $i$ and a member of group
$j$ is equally likely. This assumption is clearly unrealistic for the majority
of human communities, and can lead to significant errors in predicted epidemic
trajectories when there is substantial heterogeneity
present~\autocite{bansal2007individual, volz2007susceptible}. Contact networks
provide a way to relax this assumption by representing individuals and their
contacts explicitly. It is important to note that, although panmixia is an
unrealistic modelling assumption, it has not proven a substantial hurdle to
epidemic modelling in practice~\autocite{anderson1992infectious}. Using this
assumption, researchers have been able to derive estimates of the transmission
rate and the basic reproductive number of various outbreaks, which have agreed
with values obtained by on-the-ground data collection. Therefore, if one is
interested only in these population-level variables, the additional complexity
of contact network models may not be warranted. Rather, these models are most
useful when we are interested in properties of the network itself, such as
centrality, structural balance, and
transitivity~\autocite{wasserman1994social}.

From a public health perspective, knowledge of contact networks has the
potential to be extremely useful. On a population level, network structure can
dramatically affect the speed and pattern of epidemic
spread~\autocite[\eg][]{barthelemy2005dynamical, volz2008sir}. For example,
epidemics are expected to spread more rapidly in networks having the ``small
world'' property, where the average path length between two nodes in the
network is relatively low~\autocite{watts1998collective}. Some sexually
transmitted infections would not be expected to survive in a homogeneously
mixed population, but their long-term persistence can be explained by contact
heterogeneity~\autocite{anderson1992infectious, pastor2001epidemic}. Hence, the
contact network can provide an idea of what to expect as an epidemic unfolds.
In terms of actionable information, vaccination strategies which would
eradicate an epidemic in a random network might not work if the network is
scale-free~\autocite[][see \cref{subsec:pa}]{keeling2005networks}. On a local
level, contact networks can be informative about the groups or individuals who
are at highest risk of acquiring or transmitting infection, and would therefore
benefit most from public health interventions~\autocite{wang2015targeting,
little2014using}.

Contact networks are a challenging type of data to collect, requiring extensive
epidemiological investigation in the form of contact
tracing~\autocite{morris1993epidemiology, welch2011statistical,
keeling2005networks}. Therefore, it has been necessary to explore less
resource-intensive alternatives which still contain information about
population structure. For instance, it is possible to obtain limited
information about the contact network by individual interviews without contact
tracing. Variables which can be estimated in this fashion are referred to as
\defn{node-level} measures~\autocite{wasserman1994social}. One of the most
well-studied of these is the degree distribution, which can be estimated by
simply asking each person how many contacts they had in some interval of
time~\autocite{liljeros2001web, schneeberger2004scale, colgate1989risk}.

An alternative approach has been the analysis of other networks, which can be
estimated with phylogenetic methods from viral sequence data. Some work focuses
on the \defn{phylogenetic network}, in which two nodes are connected if the
genetic distance between their viral sequences is below some threshold.
Primarily, this work has focused on the detection of \defn{phylogenetic
clusters}, which are groups of individuals whose viral sequences are
significantly more similar to each other's than to the general population's.
The phylogenetic network is informative about ``hotspots'' of transmission and
can be used to identify demographic groups to whom targeted interventions are
likely to have the greatest effect~\autocite{poon2015impact}.  However, this
network may show little to no agreement with a contact data obtained through
epidemiological methods~\autocite{yirrell1998molecular, resik2007limitations,
robinson2013dynamics}, and therefore may be a poor proxy for the contact
network. Other studies~\autocite{brown2011transmission} have investigated the
\defn{transmission network}, which is the subgraph of the contact network
consisting of infected nodes and the edges which led to their
infections~\autocite{welch2011statistical} (\cref{fig:contactnet}, left). It is
possible to estimate the transmission network phylogenetically, although the
methods required for doing so are more sophisticated than for estimating the
phylogenetic network~\autocite{brown2011transmission}. These studies again
mostly focusing on clustering, and also on degree distributions.

Other statistical methods have been developed to infer contact network
parameters strictly from the timeline of an epidemic, using neither genetic
data nor reported contacts. \textcite{britton2002bayesian} developed a Bayesian
method to infer the $p$ parameter of an \gls{ER} network, along with the
transmission and removal rate parameters of the \gls{SI} model, using observed
infection and optionally removal times. However, it was designed for only a
small number of observations, and was unable to estimate $p$ independently from
the transmission rate.  \textcite{groendyke2011bayesian} significantly updated
and extended the methodology of \citeauthor{britton2002bayesian}, and applied
it to a measles outbreak affecting 188 individuals. They were able to obtain a
much more informative estimate of $p$, although this data set included both
symptom onset and recovery times for all individuals, and was unusual in that
the entire contact network was presumed to be infected. \textcite{volz2008sir}
developed differential equations describing the dynamics of the \gls{SIR} model
on a wide variety of random networks defined by their degree distributions.
Although the topic of estimation was not addressed in the original paper,
\citeauthor{volz2008sir}'s method could in principle be used to fit such models
to observed epidemic trajectories, similar to what is done with the ordinary
\gls{SIR} model. \textcite{volz2007susceptible} later extended the method to
dynamic contact networks and applied it to a sexual network relating 99
individuals investigated during a syphilis outbreak.

\subsection{Scale-free networks and preferential attachment}
\label{subsec:pa}

A \defn{scale-free} network is one whose degree distribution follows a power
law, meaning that the number of nodes in the network with degree $k$ is
proportional to $k^{-\gamma}$ for some constant
$\gamma$~\autocite{barabasi1999emergence}. Scale-free networks are
characterized by a large number of nodes of low degree, with relatively few
``hub'' nodes of very high degree. Epidemiological surveys have indicated that
human sexual networks tend to be scale-free~\autocite{liljeros2001web,
schneeberger2004scale, colgate1989risk}. Interestingly, many other types of
network, including computer networks, biological neural networks, metabolic
networks~\autocite{jeong2000large}, and academic co-author networks, also have
the scale-free property.

Several properties of scale-free networks are relevant in epidemiology.  The
high-degree hub nodes are known as
\defn{superspreaders}~\autocite{kemper1980identification}, which have been
postulated to contribute in varying degree to the spread of diseases such as
\gls{HIV}~\autocite{stadler2013uncovering} and
\gls{SARS}~\autocite{shen2004superspreading}. Scale-free networks have no
epidemic threshold~\autocite{pastor2001epidemic}, meaning that diseases with
arbitrarily low transmissibility can persist at low levels indefinitely. This
is in contrast with homogeneously mixed populations, in which transmissibility
below the epidemic threshold would result in exponential decay in the number of
infected individuals and eventual extinction of the pathogen (Anderson \& May,
I think).

One mechanism which has been shown to lead to scale-free networks is
\defn{preferential attachment}~\autocite{simon1955class,
barabasi1999emergence}. The simplest preferential attachment model is known as
the \acrlong{BA} (BA) model after its
inventors~\autocite{barabasi1999emergence}. Under this model, networks are
formed by starting with a small number $m_0$ of nodes. New nodes are added one
at a time until there are a total of $N$ in the network. Each time a new node
is added, $m \geq 1$ edges are added from it to other nodes in the graph. In
the original formulation, the partners of the new node are chosen with
probability linearly proportional to their degree. However,
\citeauthor{barabasi1999emergence} suggest extending the model such that the
probability of choosing a partner of degree $d$ is proportional to $d^\alpha$
for some constant $\alpha$, and we use this extension here. When $m = 1$, the
network takes on the distinctive shape of a tree, that is, it does not contain
any cycles. Cycles are present in the network for all all other $m$ values.

There has been some contention of the idea that contact networks are
scale-free. \textcite{handcock2004likelihood} fit several stochastic models of
partner formation to empirical degree distributions derived from population
surveys of sexual behaviour. They found that a negative binomial distribution,
rather than a power law, was the best fit to five out of six datasets, although
the difference in goodness of fit was extremely small in four out of these
five. \textcite{bansal2007individual} found that an exponential distribution,
rather than a power law, was the best fit to degree distributions of six social
or sexual networks.

\subsection{Relationship between network structure and transmission trees}

The contact network underlying an epidemic constrains the shape of the
transmission network, which in turn determines the topology of the transmission
tree relating the infected hosts (\cref{fig:contactnet}). The index case who
introduces the epidemic into the network becomes the root of the tree. Each
time a transmission occurs, the lineage corresponding to the donor host in the
tree splits into two, representing the recipient lineage and the continuation
of the donor lineage. \Cref{fig:contactnet} illustrates this correspondence.
It's important to note that, although the order and timing of transmissions
determines the tree topology uniquely, the converse does not hold. That is, for
any given topology, there are in general many transmission networks which would
lead to that topology. In other words, it impossible to distinguish who
transmitted to whom from a transmission tree alone~\autocite{bernard2007hiv}.

A number of studies have made progress in quantifying the relationship between
contact networks and transmission trees. \textcite{o2010contact} simulated
epidemics over networks with four types of degree distribution. They then
estimated the Bayesian skyride~\autocite{minin2008smooth} population size
trajectory in two ways: from the phylogeny, using \gls{MCMC}; and from the
incidence and prevalence trajectories, using the method developed by 
\textcite{volz2009phylodynamics}. They found that the concordance between
the two skyrides, as well as the relationship between the skyride and
prevalence curve, was qualitatively different for each degree distribution.
\textcite{leventhal2012inferring} investigated the relationship between
transmission tree imbalance and several epidemic parameters under four contact
network models, and found that these relationships varied considerably
depending on which model was being considered. \textcite{welch2011network}
simulated transmission trees over networks with varying degrees of community
structure. They found that transmission trees simulated under networks with low
clustering could not generally be distinguished from those simulated under
highly clustered networks, and concluded that contact network clusters do not
affect transmission tree shape. However, more recently,
\textcite{villandre2016assessment} investigated the correspondence between
contact network clusters and transmission tree clusters, and found a moderate
correspondence between the two.

In summary, studies in this group have demonstrated that network structure
profoundly influences tree shape, but have not attempted to quantitatively
infer network parameters from observed trees.

%\section{Approximate Bayesian computation}

%Consider a model $M$, with parameters $\theta$, which we wish to fit to some
%observed data $D$. By ``fit'', we often mean that we want to find particular
%values $\hat{\theta}$ for the parameters which optimize the likelihood of our
%data given those parameters and the model,
%\[
%  \hat{\theta} = \argmax_\theta \Pr(D | M, \theta).
%\]
%This $\hat{\theta}$ is the \defn{maximum likelihood} parameter estimate. Note
%that we are employing a common abuse of notation here, where $\Pr(\cdots)$ is
%being unsed to refer to a probabilty \emph{density} rather than a true
%probability. Alternative to maximum likelihood, we may be interested less in a
%point estimate and more in the posterior distribution of possible values of
%$\theta$ given our data, $\Pr(\theta \mid M, D)$. This will be expanded upon
%below, but for the moment, for illustrative purposes, we restrict our attention
%to the maximum likelihood problem.
%
%If the model we are fitting is sufficiently simple, it may be possible to
%calculate $\hat{\theta}$ directly, using calculus. Most models do not admit
%analytic maximum likeilhood solutions, but if the likelihood any set of
%parameters can be calculated up to a normalizing constant, then
%${\Pr(D \mid M,\theta)}$ can be optimized numerically. A wide range of
%optimization strategies exist, the choice of which to use depending on the
%complexity of the model and whether or not we have access to the gradients of
%the likelihood function with respect to each of the parameters. The majority of
%modelling problems fit into this category, and numerical optimization is
%well-developed and extremely widely used.
%
%However, there are some cases, often when the observed data is of a complex
%type, that explicitly calculating the likelihood of some observed data is
%impossible, even up to a normalizing constant. For example, suppose that we
%want to model a chess player's behaviour. We will set up a simple one-parameter
%model which describes the chess playing process.  The parameter, $a \in [0, 1]$
%indicates the player's eagerness to remove his opponent's pieces from the
%board. We can write down an algorithm for the player's behaviour under such a
%model.
%
%% TODO: don't break this over the page
%
%\begin{algorithmic}
%  \While{the game is not over}
%    \If{I can capture an opponent's piece and $\Uniform(0, 1) < a$}
%      \State{capture the piece}
%    \Else
%      \State{make any other move at random}
%    \EndIf
%  \EndWhile
%\end{algorithmic}
%
%Suppose the observed data are the ending configurations of the board, after the
%player has concluded a game against an opponent with a known value of $a$. The
%model we have designed is very simple, but it is not obvious how to calculate
%the likelihood of a particular ending configuration. Indeed, it seems that the
%only way is to enumerate every possible path the game could have taken, and
%tabulate the ending configurations of each. Clearly, this is infeasible.
%Approximate Bayesian computation was designed for situations like these, where
%exact likelihoods are not available, perhaps due to the model involving an
%algorithm or generative process.

\section{Sequential Monte Carlo}
\label{sec:smc}

\Gls{SMC} is a statistical inference method which samples from a sequence of
probability distributions in a fixed order~\autocite{del2006sequential}, based
on a technique called \gls{SIS}.

To fully describe \gls{SMC}, we will introduce some notation and terminology.
For a sequence $x_1, \ldots, x_d$, we will write $\vec{x_i}$ to mean the
partial sequence $x_1, \ldots, x_i$.

A \defn{Markov kernel} is the continuous analogue of the transition matrix in a
finite-state Markov model. For some spaces $X$ and $Y$, $K \maps X \times Y \to
[0, 1]$ such that
\begin{align}
    \label{eq:mk}
    \int_Y K(x, y) \d y = 1
\end{align}
for all $x \in X$. This is an ``operational'' definition of Markov kernel which
will be suitable for our purposes. A more precise definition can be found in
\textbf{TODO}. Note that a Markov kernels have nothing to do with the kernel
functions defined in \cref{subsec:treeshape}, other than sharing a name.
Unfortunately, the word ``kernel'' is ubiquitous in mathematics and refers to
many unrelated concepts.

\subsection{Sequential importance sampling}
\label{subsec:sis}

The basis of \gls{SIS} is \gls{IS}, which is a method of estimating summary
statistics of distributions which are known only up to a normalizing constant,
and therefore cannot be sampled from directly. That is, if $\pi$ is such a
distribution and $f$ is any real-valued function, \gls{IS} is concerned with
estimating
\[
  \pi(f) = \int f(x)\pi(x)\d x = \int f(x) \frac{\gamma(x)}{Z} \d x,
\]
where the integral is over the space on which $\pi$ is defined, $\gamma(x)$ is
known pointwise, and $Z = \int \gamma(x) \d x$ is the unknown normalizing
constant. Suppose we have at hand another distribution $\eta$, called the
\defn{importance distribution}, from which we are able to sample. Define the
\defn{importance weight} as the ratio ratio $w(x) = \gamma(x)/\eta(x)$. We can
express the normalizing constant $Z$ in terms of the importance weight and
distribution, $Z = \int w(x) \eta(x) \d x$, and in turn write the expectation
of interest as
\[
  \int f(x) \pi(x) \d x = \frac{\int f(x) \gamma(x) \d x}
                               {\int w(x) \eta(x) \d x}.
\]
If we sample a large number of points from $\eta$, then $\eta(x)$ can be
approximated by a Monte Carlo estimate. Since the remaining quantities $f$,
$\gamma$, and $w$ can all be evaluated pointwise, these are all the ingredients
we need to obtain an estimate of $\pi(f)$. Although this is a simple and
elegant approach, the drawback is that the variance of the estimate is
proportional to the variance of the importance weights~\autocite{liu2008monte},
which may be quite large if $\eta$ and $\gamma$ are very different. Therefore,
the practical use of \gls{IS} on its own is limited, since it depends on
finding an importance distribution similar to $\pi$, which we usually know very
little about \textit{a priori}.

\gls{SIS} applies \gls{IS} to the problem of sampling from a distribution $\pi$ 
on an high-dimensional space, say $\pi(\vec{x}) = \pi(x_1, \ldots, x_d)$. The
objective is to build up an importance distribution $\eta$ for $\pi$
sequentially. Let $\vec{x_i} = (x_1, \ldots, x_i)$, for $1 \leq i \leq d$. By
the general product rule,
\[
  \pi(\vec{x}) 
  = \pi(x_1) \pi(x_2 \mid x_1) \cdots
    \pi(x_{d-1} \mid \vec{x_{d-2}}) \pi(x_d \mid \vec{x_{d-1}}).
\]
This decomposition is natural in many contexts, particularly for on-line
estimation. For example, in a stateful model like an \gls{HMM}, $x_i$ may
represent the state at time $i$, with $\pi(\vec{x})$ being the posterior
distribution over possible paths. The importance distribution $\eta$ for $\pi$
will be constructed using a similar decomposition,
\[
  \eta(\vec{x}) 
  = \eta(x_1) \eta(x_2 \mid x_1) \cdots
    \eta(x_{d-1} \mid \vec{x_{d-2}}) \eta(x_d \mid \vec{x_{d-1}}).
\]
The importance weights for $\eta$ with respect to $\pi$ are then defined
recursively as
\begin{align}
  \label{eq:sisw}
  w(\vec{x_i}) = \frac{\pi(\vec{x_i})}{\eta(\vec{x_i})}
  = \frac{\pi(x_i \mid \vec{x_{i-1}})\pi(\vec{x_{i-1}})}
         {\eta(x_i \mid \vec{x_{i-1}})\eta(\vec{x_{i-1}})}
  = \frac{\pi(x_i \mid \vec{x_{i-1}})}
         {\eta(x_i \mid \vec{x_{i-1}})}\cdot w(\vec{x_{i-1}}).
\end{align}
Thus, we can choose $\eta(x_i \mid \vec{x_{i-1}})$ such that the variance of
the importance weights is as small as possible at every step, eventually
arriving at a full importance distribution. The process for choosing $\eta$
will differ according to the problem at hand. One potential choice for
$\eta(x_i \mid \vec{x_{i-1}})$ is simply $\pi(x_i \mid \vec{x_{i-1}})$, if it
is possible to compute. For example, in the case of an \gls{HMM}, $\pi(x_i \mid
\vec{x_{i-1}})$ is simply the transition probability from $x_{i-1}$ to $x_i$.
In a Bayesian setting, $\eta$ may be taken to be the prior distribution. In
some applications, it is possible to compute or approximate the form of $\eta$
which minimizes the variance of the weights. The resulting importance
distribution is known as the \defn{optimal
kernel}~\autocite{cappe2007overview}. Choosing each element of the
decomposition of $\eta$ is a problem-specific task (see \textit{e.g.}
\autocite{smith2013sequential,liu2008monte} for many examples).

The recursive definition \cref{eq:sisw} suggests an algorithm for obtaining a
sample from $\pi$. We begin with a sample of $n$ ``particles'' which have been
sampled from the importance distribution $\eta(x_0)$ for $\pi(x_0)$. The
particles are updated and reweighted $d$ times, corresponding to the $d$
elements of the decomposition of $\pi$. At the $i$th step, each particle is
extended to include $x_i$ drawn from the chosen $\eta(x_i \mid \vec{x_{i-1}})$,
and the importance weights are recalculated and normalized.

Of course, $\eta$ is merely an approximation to $\pi$, and may be a fairly poor
one depending on the application. Try as we might to keep the variances of the
weights low, the cumulative errors at each sequential step tend to push many of
the weights to very low values. This results in a poor approximation to $\pi$,
since only a few particles retain high importance weights after all $d$
sequential steps. To mitigate this problem, we periodically apply a resampling
step when the variance in the importance weights becomes too high. Several
different criteria have been proposed for when to resample, but we focus here
on the one described by \textcite{liu2008monte}, namely the decay of the
\gls{ESS} below a prescribed threshold. The \gls{ESS} of the population of
particles is defined as
\[
  \ESS(w) = \frac{n}{1 + \Var(w)},
\]
where $n$ is the number of particles~\autocite{liu2008monte}. When the
\gls{ESS} drops below the threshold, we resample the particles according to
their weights. This results in the removal of low-weight particles from the
population, and also equalizes all the weights. Various resampling strategies
beyond the basic sampling with replacement have been proposed, but we will not
discuss those here. The full \gls{SIS} algorithm with resampling is given as
\cref{alg:sis}. We use a superscript $^{(k)}$ to refer to the $k$th particle,
and for brevity write $w$ for $w(\vec{x_i})$. The variable $T$ is the threshold
for the \gls{ESS}, below which the population is resampled. It is conventional
to take $T = n/2$.

\begin{algorithm}
  \caption{Sequential importance sampling.}
  \begin{algorithmic}
    \For {$k = 1$ to $n$}
      \State Sample $x_0^{(k)}$ from $\eta(x_0)$
      \Comment{Initialize the $k$th particle}
      \State $w^{(k)} \gets \dfrac{\pi\left(x_0^{(k)}\right)}{\eta\left(x_0^{(k)}\right)}$
    \EndFor
    \For {$i = 1$ to $d$}
      \For {$k = 1$ to $n$}
        \State Sample $x_i^{(k)}$ from $\eta\left(x_i \mid \vec{x_{i-1}^{(k)}}\right)$
        \Comment Extend the $k$th particle
        \State $w^{(k)} \gets \dfrac{\pi\left(x_i^{(k)} \mid \vec{x_{i-1}^{(k)}}\right)}{\eta\left(x_i^{(k)} \mid \vec{x_{i-1}^{(k)}}\right)} \cdot w^{(k)}$
      \EndFor
      \State Normalize the weights so that $\sum w = 1$
      \If {$\ESS(w) < T$}
        \State Resample the particles according to $w$
        \For {$k = 1$ to $n$}
          \State $w^{(k)} \gets 1/n$
        \EndFor
      \EndIf
    \EndFor
    \State Sample $n$ particles with probabilities $w$
  \end{algorithmic}
  \label{alg:sis}
\end{algorithm}

\subsection{Sequential Monte Carlo samplers}
\label{subsec:smcsamp}

The \gls{SIS} algorithm described above aims to sample from a high-dimensional
distribution $\pi(x)$, by sequentially sampling from $d$ distributions of lower
but increasing dimension. \textcite{del2006sequential} developed an
\defn{\gls{SMC} sampler} with an alternative objective: to sample sequentially
from $d$ distributions $\pi_1, \ldots, \pi_d$, all of the same dimension and
defined on the same space. The $\pi_i$ are assumed to form a related sequence,
such as posterior distributions attained by sequentially considering new
evidence. As with \gls{SIS}, we assume that $\pi_i(x) = \gamma_i(x) / Z_i$,
where $\gamma_i$ is known pointwise and the normalizing constant $Z_i$ is
unknown.

Both algorithms involve progression through a sequence of related
distributions. For \gls{SIS}, these distributions are lower-dimensional
marginals of the target distribution, while for the \gls{SMC} sampler, they are
of the same dimension and constitute a smooth progression from an initial to a
final distribution. In both cases, the neighbouring distributions in the
sequence are related to each other in some way, and we can take advantage of
that relationship to create a sequence of importance distributions alongside
the sequence of targets. In \gls{SIS}, the neighbouring marginals
$\pi(\vec{x_i})$ and $\pi(\vec{x_{i+1}})$ were related by the conditional
density $\pi(x_i \mid \vec{x_{i-1}})$, which we used to inform the importance
distribution. In \gls{SMC}, the relationship between subsequent distributions
is less explicit, but it is assumed that they are related closely enough that
an importance distribution for $\pi_i$ can be easily transformed into one for
$\pi_{i+1}$. In particular, the sequence of importance distributions $\eta_i$
is constructed as
\begin{align}
  \label{eq:impint}
  \eta_i(x) = \int \eta_{i-1}(\xi) K_i(\xi, x) \d\xi,
\end{align}
where $K_i$ is a Markov kernel and the integral is over the space on which the
$\pi_i$ are defined. The choice of $K_i$ should be based on the percieved
relationship between $\pi_{i-1}$ and $\pi_i$. \textcite{del2006sequential}
propose the use of a \gls{MCMC} kernel with equilibrium distribution $\pi_i$.
That is,
\[
  K_i(\xi, x) = \max\left(1, \frac{q(x, \xi)\pi_i(\xi)}{q(\xi, x)\pi_i(x)}\right),
\]
where $q(\xi, x)$ is a proposal function such as a Gaussian distribution
centered at $\xi$ (see \cref{subsec:mfit}).

Although this method of building up $\eta$ appears straightforward, the
drawback is that the importance distribution itself becomes intractible. In
particular, evaluating $\eta_i(x)$ involves a $i$-dimensional integral of the
type in \cref{eq:impint}. As it is necessary to evaluate $\eta(x)$ pointwise to
perform \gls{IS}, this construction appears to have defeated the purpose of
providing an importance distribution for each $\pi_i$.
\textcite{del2006sequential} overcome this problem with two ``artificial''
objects. First, they propose the existence of \textit{backward} Markov kernels
$L_{i-1}(x_i, x_{i-1})$. For now, these kernels are arbitrary, and will be
precisely defined on a problem-specific basis. Second, they define an
alternative sequence of target distributions
\[
  \tilde{\pi}_i(\vec{x_i}) = \pi_i(x_i) \prod_{k=1}^{i-1} L_k(x_{k+1}, x_k)
\]
of increasing dimension. This brings us back to the setting described above in
\cref{subsec:sis}, namely of building up an importance distribution of
dimension $d$ sequentially through lower-dimensional distributions. We can
write $\tilde{\pi}_i$ in terms of $\tilde{\pi}_{i-1}$ by noticing that
\begin{align*}
  \frac{\tilde{\pi}_i(\vec{x_i})}{\tilde{\pi}_{i-1}(\vec{x_{i-1}})} 
  = \frac{\pi_i(x_i) \prod_{k=1}^{i-1} L(x_{k+1}, x_k)}
  {\pi_{i-1}(x_{i-1}) \prod_{k=1}^{i-2} L(x_{k+1}, x_k)}
  = \frac{\pi_i(x_i) L(x_i, x_{i-1})}{\pi_{i-1}(x_{i-1})},
\end{align*}
and hence
\[
  \tilde{\pi}_i = \frac{\pi_i(x_i) L(x_i, x_{i-1})}{\pi_{i-1}(x_{i-1})} \cdot \tilde{\pi}_{i-1}.
\]
Therefore, the importance weights for these new targets are defined recursively as
\begin{align}
  w(\vec{x_i}) 
    &= \frac{\tilde{\pi}_i(\vec{x_i})}{\eta_i(\vec{x_i})} \\
    &= \frac{\tilde{\pi}_{i-1}(\vec{x_{i-1}}) \pi_i(x_i) L(x_i, x_{i-1})}
           {\eta_{i-1}(\vec{x_{i-1}}) \pi_{i-1}(x_{i-1}) K_i(x_{i-1}, x_i)} \\
    &= w(\vec{x_{i-1}}) \cdot
      \frac{\pi_i(x_i) L_{i-1}(x_i, x_{i-1})}
           {\pi_{i-1}(x_{i-1}) K_i(x_{i-1}, x_i)} \\
    &\propto w(\vec{x_{i-1}}) \cdot
      \frac{\gamma_i(x_i) L_{i-1}(x_i, x_{i-1})}
           {\gamma_{i-1}(x_{i-1}) K_i(x_{i-1}, x_i)}.
    \label{eq:smcwt}
\end{align}
The final key piece of information is to notice that, because the $L_i$ are
Markov kernels, $\pi_i$ is simply the marginal in $\vec{x_{i-1}}$ of
$\tilde{\pi}$. Therefore, a sample from $\tilde{\pi}_i$ automatically gets us a
sample from $\pi_i$, by considering only the $i$th component of $\vec{x_i}$.
These are all the ingredients we need to apply \gls{SMC}. The sequences of
kernels $L$ and $K$ should be chosen based on the problem at hand to inimize
the variance in the importance weights as well as possible. For a fixed choice
of $K_i$, the backward kernels $L_i$ which minimize this variance are called
the \defn{optimal} backward kernels. The full \gls{SMC} sampler algorithm is
presented as \cref{alg:smcsamp}.

\begin{algorithm}
  \caption{Sequential Monte Carlo sampler of \textcite{del2006sequential}.}
  \begin{algorithmic}
    \For {$k = 1$ to $n$}
      \State Sample $x_1^{(k)}$ from $\eta_1(x_1)$
      \Comment{Initialize the $k$th particle}
      \State $w^{(k)} \gets \dfrac{\gamma_1\left(x_1^{(k)}\right)}{\eta_1\left(x_1^{(k)}\right)}$
      \State Normalize the weights so that $\sum w = 1$
    \EndFor
    \For {$i = 2$ to $d$}
      \For {$k = 1$ to $n$}
        \State Sample $x_i^{(k)}$ from $K(x_{i-1}^{(k)}, x_i)$
        \Comment Extend the $k$th particle
        \State $w^{(k)} \gets w^{(k)} \cdot \dfrac{\gamma_i(x_i) L_{i-1}(x_i, x_{i-1})}{\gamma_{i-1}(x_{i-1}) K_i(x_{i-1}, x_i)}$
      \EndFor
      \State Normalize the weights so that $\sum w = 1$
      \If {$\ESS(w) < T$}
        \State Resample the particles according to $w$
        \For {$k = 1$ to $n$}
          \State $w^{(k)} \gets 1/n$
        \EndFor
      \EndIf
      \State Sample the $i$th component of $n$ particles with probabilities $w$
    \EndFor
  \end{algorithmic}
  \label{alg:smcsamp}
\end{algorithm}

\section{Approximate Bayesian computation}
\label{sec:abc}

\glsreset{ABC}

\subsection{Model fitting}
\label{subsec:mfit}

A \defn{mathematical model} is a formal description of a hypothesized
relationship between some observed data, $\vec{x} = \set{x_1, \ldots, x_n}$,
and outcomes, $\vec{y} = \set{y_1, \ldots, y_n}$. A \defn{parametric} model
defines a family of possible relationships between data and outcomes, indexed
by one or more numeric parameters $\theta$. A \defn{statistical} model
describes the relationship between data and outcomes in terms of probabilities.
Statistical models define, either explicitly or implicitly, the probability of
observing $\vec{y}$ given $\vec{x}$ and, if the model is parametric, $\theta$.
In this context, the observed outcomes are taken to be realizations of random
variables $\vec{Y} = \set{Y_1, \ldots, Y_n}$. Note that it is entirely possible
to have no data $\vec{x}$, only observed outcomes $\vec{y}$. In this case, a
model would describe the process by which $\vec{y}$ is generated.

To illustrate these concepts, consider the well-known linear model. For
clarity, we will restrict our attention to the case of one-dimensional data and
outcomes where each $x_i$ and $y_i$ is a single real number. The linear model
postulates that the outcomes are linearly related to the data, modulo some
noise introduced by measurement error, environmental fluctuations, and other
external factors. Formally, $y_i = \beta x_i + \varepsilon_i$, where $\beta$ is
the slope of the linear relationship, and $\varepsilon_i$ is the error
associated with measurement $i$. We can make this model a statistical one by
hypothesizing a distribution for the error terms $\varepsilon_i$; most
commonly, it is assumed that they are normally distributed with variance
$\sigma$. In mathematical terms, $Y_i \sim \beta x_i + \N(0, \sigma^2)$, where
``$\sim$'' means ``is distributed as''. We can see from this formulation that
the model is parametric, with parameters $\theta$ = ($\beta$, $\sigma$).
Moreover, we can write down the probability density of observing outcome $y_i$
given the parameters,
\[
  \Pr(Y_i = y_i \mid \beta, \sigma) = 
  \Pr\left(\N(0, \sigma^2) = y_i - \beta x_i\right).
\]
Note that we have followed the standard statistical abuse of notation and used
``$\Pr$'' to refer to a probability \emph{density}, rather than a probability.
Also, we are treating the $x_i$ as fixed quantities, and therefore have not
conditioned the probability density on $\vec{x}$. Assuming all the $y_i$ are
independent, the probability density of the entire observed set of outcomes is
the product of the probability density of each individual $y_i$,
\[
  \Pr(\vec{Y} = \vec{y} \mid \beta, \sigma) = 
  \prod_{i=1}^N \Pr(Y_i = y_i \mid \beta, \sigma).
\]

For a general model, the probability density of $\vec{y}$ given the parameters
$\theta$ is also known as the \defn{likelihood}, written $\L$, of $\theta$.
That is, $\L(\theta \mid \vec{y}) = \Pr(\vec{y} \mid \theta)$. The higher the
value of the likelihood, the more likely the observations $\vec{y}$ are under
the model. Thus, the likelihood provides a natural criterion for fitting the
model parameters: we want to pick $\theta$ such that the probability density of
our observed outcomes $\vec{y}$ is as high as possible. The parameters which
optimize the likelihood are known as the \textit{\gls{ML}} estimates, denoted
$\hat{\theta}$. That is,
\[
  \hat{\theta} = \argmax_\theta\; \Pr(\vec{y} \mid \theta).
\]
\Gls{ML} estimation is usually performed with numerical optimization. In the
simplest terms, many possible values for $\theta$ are examined, $\L(\theta \mid
\vec{y})$ is calculated for each, and the parameters which produce the highest
value are accepted. Many sophisticated numerical optimization methods exist,
although they may not be guaranteed to find the true \gls{ML} estimates if the
likelihood function is complex. Occasionally, as is the case with linear
regression, the \gls{ML} estimates can be found explicitly by setting the
likelihood function's derivatives to zero.

\Gls{ML} estimation makes use only of the data and outcomes to estimate the
model parameters $\theta$. However, it is frequently the case that the
investigator has some additional information or belief about what $\theta$ are
likely to be. For example, in the linear regression case, the instrument used
to measure the outcomes may have a well-known margin of error, or the sign of
the slope may be obvious from previous experiments. The Bayesian approach to
model fitting makes use of this information by codifying the investigator's
beliefs as a \defn{prior distribution} on the parameters, denoted
$\Pr(\theta)$. Instead of considering only the likelihood, Bayesian inference
focuses on the product of the likelihood and the prior, $\Pr(\vec{y} \mid
\theta) \Pr(\theta)$. Bayes' theorem tells us that this product is related to
the \textit{posterior distribution} on $\theta$,
\[
  \Pr(\theta \mid \vec{y}) = \frac{\Pr(\vec{y} \mid \theta) \Pr(\theta)}
                                  {\Pr(\vec{y})}.
\]

In principle, $\Pr(\vec{y} \mid \theta) \Pr(\theta)$ can be optimized
numerically just like $\L(\theta \mid \vec{y})$, which would also optimize the
posterior distribution. The resulting optimal parameters are called the
\gls{MAP} estimates. However, from a Bayesian perspective, $\theta$ is not a
fixed quantity to be estimated, but rather a random variable with an associated
distribution (the posterior). Therefore, the \gls{MAP} estimate by itself is of
limited value without associated statistics about the posterior distribution,
such as the mean or credible intervals. Unfortunately, to calculate such
statistics, it is necessary to evaluate the normalizing constant
$\Pr(\vec{y})$, which is almost always an intractable integral.

A popular method for circumventing the normalizing constant is the use of
\gls{MCMC} to obtain a sample from the posterior distribution. \Gls{MCMC} works
by defining a Markov chain whose states are indexed by possible model
parameters. The transition probability from state $\theta_1$ to state
$\theta_2$ is taken to be
\[
  \max\left(1, \frac{\Pr(\vec{y} \mid \theta_2) \Pr(\theta_2) q(\theta_2, \theta_1)}
                    {\Pr(\vec{y} \mid \theta_1) \Pr(\theta_2) q(\theta_1, \theta_2)} \right),
\]
where $q(\theta, \theta')$ is a symmetric \defn{proposal distribution} used in
the algorithm to generate the chain. The stationary distribution of this Markov
chain is equal to the posterior distribution on $\theta$. Therefore, if a long
enough random walk is performed on the chain, the distribution of states
visited will be a Monte Carlo approximation of $\Pr(\theta \mid \vec{y})$, from
which we can calculate statistics of interest. Actually performing this random
walk is straightforward and can be accomplished via the Metropolis-Hastings
algorithm (\cref{alg:mh}).

\begin{algorithm}
  \caption{Metropolis-Hastings algorithm for Markov chain Monte Carlo.}
  \begin{algorithmic}
    \State Draw $\theta$ according to $\Pr(\theta)$
    \Loop
      \State Propose $\theta'$ according to $q(\theta, \theta')$
      \State Accept $\theta \gets \theta'$ with probability
      $\max \left( 1, 
       \dfrac{\Pr(\vec{y} \mid \theta') \Pr(\theta') q(\theta', \theta)}
             {\Pr(\vec{y} \mid \theta\phantom{'}) \Pr(\theta\phantom{'}) q(\theta, \theta')}
       \right)$
    \EndLoop
  \end{algorithmic}
  \label{alg:mh}
\end{algorithm}

\subsection{Overview of ABC}
\label{subsec:abcoverview}

Most mathematical models are amenable to fitting via one or both of the
approaches, \gls{ML} or Bayesian inference, discussed above. However, there are
some, particularly in the domain of population genetics, for which calculation
of either the likelihood or the product of the likelihood and the prior may be
infeasible. For example, one or both of these quantities may be expressible
only as an intractable integral. \Gls{ABC} is designed for such cases, where
standard likelihood-based techniques for model fitting cannot be applied. Such
models are particularly prevalent in population
genetics~\autocite{beaumont2002approximate, beaumont2010approximate}.

Ordinarily, Bayesian inference targets the posterior distribution $\Pr(\theta
\mid \vec{y})$. That is, in the Bayesian framework, model parameters with
higher posterior density are ``better'' in the sense that they offer a more
credible explanation for the observed data. Approximate Bayesian computation
offers an alternative metric for parameter credibility, namely the similarity
of simulated datasets to the observed data. If datasets simulated under the
model closely resemble the real data, it follows that the model is a reasonable
approximation to the real-world process generating the observed data. More
formally, suppose we have a distance measure $\rho$ defined on the space of all
possible data our model could generate. \gls{ABC} aims to sample from the joint
posterior distribution of model parameters and simulated datasets $\vec{z}$
which are within some small distance $\varepsilon$ of the observed data
$\vec{y}$,
\[
  \Pr(\theta, \vec{z} \mid \vec{y}, \varepsilon) =
  \frac{\Pr(\theta) \Pr(\vec{z} \mid \theta) \I_{A_{\varepsilon, \vec{y}}} (\vec{z})}
  {\int_{A_{\varepsilon, \vec{y}} \times \Theta} \Pr(\theta) \Pr(\vec{z} \mid \theta) \d \theta}.
\]
Here, $A_{\varepsilon, \vec{y}}$ is an $\varepsilon$-ball around $\vec{y}$ with
respect to $\rho$, $\Theta$ is the space of all possible model parameters, and
$\I$ is the indicator function~\autocite{marin2012approximate}. As we shall
see in the next section, this distribution can be sampled from exactly. The
word ``approximate'' derives from the assumption that, for a suitably chosen
distance $\rho$ and a small enough $\varepsilon$, the marginal in $\vec{z}$ of
this distribution approximates the posterior of
interest~\autocite{marin2012approximate}. That is,
\[
  \int \Pr(\theta, \vec{z} \mid \vec{y}, \varepsilon) \d\vec{z}
  \approx \Pr(\theta \mid \vec{y}).
\]
To emphasize that the distribution on the left only approximates the posterior,
we will refer to it as the \textit{\gls{ABC} target distribution}. Note that in
many formulations, the distance function $\rho$ is defined as $\rho(S(\cdot),
S(\cdot))$ where $S$ is a function which maps data points into a vector of
summary statistics. This can be useful if the data are high-dimensional or of a
complex type, but it is not strictly necessary. For instance, if the data are
numeric and of low dimension, the distance function may simply be the Euclidian
distance~\autocite{sisson2007sequential}. For more complex data,
\textcite{nakagome2013kernel} proposed the use of a kernel function (defined in
\cref{subsec:treeshape}), an approach they dubbed \defn{kernel-\gls{ABC}}.

\subsection{Algorithms for ABC}
\label{subsec:abcalg}

Algorithms for performing \gls{ABC} fall into one of three categories:
rejection, \gls{MCMC}, and \gls{SMC}. To simplify the math, we shall restrict
the descriptions of these algorithms to the case of one simulated dataset per
parameter particle (the meaning of this will become clear shortly). The
extension to multiple datasets per particle will be given at the end.

Rejection ABC is the simplest method, and also the one which was first
proposed~\autocite{rubin1984bayesianly, tavare1997inferring}. The algorithm,
outlined in \cref{alg:abcrej}, repeats the following steps until a desired
number of samples from the target distribution are obtained. Parameter values
$\theta$ are sampled according to the prior distribution $\Pr(\theta)$. Then, a
simulated dataset $\vec{z}$ is generated from the model with the sampled
parameter values. By definition, the probability density of obtaining the
particular dataset $\vec{z}$ is $\Pr(\vec{z} \mid \theta)$. Finally, the
parameters are sampled if the distance of $\vec{z}$ from the observed data
$\vec{y}$ is less than $\varepsilon$ - that is, with probability
$\I_{A_{\varepsilon, \vec{y}}}(\vec{z})$. Putting this all together, the
parameters $\theta$ are sampled with probability proportional to
\[
  \Pr(\theta) \Pr(\vec{z} \mid \theta) \I_{A_{\varepsilon, \vec{y}}}(\vec{z}),
\]
which is exactly the numerator of the \gls{ABC} target distribution. Thus,
$\theta$ represents an unbiased sample from the approximate posterior.

\begin{algorithm}
  \caption{Rejection \gls{ABC}.}
  \begin{algorithmic}
    \Loop
      \State Draw $\theta$ according to $\Pr(\theta)$
      \State Simulate a dataset $\vec{z}$ from the model with parameters $\theta$
      \If{$\rho(\vec{y}, \vec{z}) < \varepsilon$}
        \State Sample $\theta$
      \EndIf
    \EndLoop
  \end{algorithmic}
  \label{alg:abcrej}
\end{algorithm}

Rejection \gls{ABC} is easy to understand and implement, but it is not
generally computationally feasible. If the posterior is very different from the
prior, a very large number of samples may need to be taken in order to find a
simulated dataset which is close to $\vec{z}$. The inefficiency is compounded
by the curse of dimensionality - the measure of the $\varepsilon$-ball around
$\vec{y}$ decreases exponentially with the number of dimensions.
\gls{ABC}-\gls{MCMC} (\cref{alg:abcmcmc}) was designed to overcome these
hurdles~\autocite{marjoram2003markov}. The approach is similar to ordinary
Bayesian \gls{MCMC} (\cref{subsec:mfit}), except that a distance cutoff
replaces the likelihood ratio. That is, the transition probability between
states $\theta_1$ and $\theta_2$, with simulated datasets $\vec{z_1}$ and
$\vec{z_2}$, is defined as
\[
  \max\left(1, \frac{\Pr(\vec{z_2} \mid \theta_2) q(\theta_2, \theta_1)}
                    {\Pr(\vec{z_1} \mid \theta_1) q(\theta_1, \theta_2)} 
    \cdot \I_{A_{\varepsilon, \vec{y}}}(\vec{z_2}) \right).
\]

\begin{algorithm}
  \caption{ABC-MCMC.}
  \begin{algorithmic}
    \State Draw $\theta$ according to $\Pr(\theta)$
    \Loop
      \State Propose $\theta'$ according to $q(\theta, \theta')$
      \State Simulate a dataset $\vec{z'}$ according to the model with
             parameters $\theta$
      \State Accept $\theta \gets \theta'$ with probability
      $\max \left( 1, 
       \dfrac{\Pr(\theta') q(\theta', \theta)}
             {\Pr(\theta\phantom{'}) q(\theta, \theta')} 
       \cdot \I_{A_{\varepsilon, \vec{y}}}(\vec{z'}) \right)$
    \EndLoop
  \end{algorithmic}
  \label{alg:abcmcmc}
\end{algorithm}

Some of the same computational inefficiencies arise with \gls{ABC}-\gls{MCMC}
as with rejection. For example, in regions of low posterior density, the
probability to simulate a dataset proximal to the observed data is low. Various
strategies have been developed to mitigate this, including reducing the
tolerance level $\varepsilon$ as the chain
progresses~\autocite{ratmann2007using}.

The most recently developed class of algorithm for \gls{ABC} is
\gls{ABC}-\gls{SMC}~\autocite{sisson2007sequential, beaumont2009adaptive}. As
with \gls{ABC}-\gls{MCMC}, the algorithm is a straightforward modification of
an existing Bayesian inference method, in this case the \gls{SMC} sampler
(\cref{subsec:smcsamp}). The sequence of target distributions is defined as
$\pi_i = \Pr(\theta, \vec{z} \mid \vec{y}, \varepsilon_i)$ for a decreasing
sequence of tolerances $\varepsilon_i$. The intention is for the algorithm to
progress smoothly through a sequence of target distributions which ends at the
\gls{ABC} approximation to the posterior. As discussed in
\cref{subsec:smcsamp}, the choices of the kernels $K$ and $L$ is
problem-specific, and so appropriate kernels must be chosen for \gls{ABC}.
Several options have been proposed~\autocite{beaumont2009adaptive,
sisson2007sequential, del2012adaptive}.

All the algorithms discussed in this section can be straightforwardly extended
to sample from the joint distribution
\[
  \pi_\varepsilon(\theta, z_1, \ldots, z_M \mid y),
\]
which is equivalent to associating $M$ simulated datasets to each parameter
particle instead of just one. The simulated dataset $z$ is replaced by
$\vec{z} = z_1, \ldots, z_M$, and the indicator function for the
$\varepsilon$-ball around $y$ is replaced by
\[
  \sum_{k=1}^M \I_{A_{\varepsilon, y}} (z_i).
\]
For \gls{ABC}-\gls{MCMC} and \gls{ABC}-\gls{SMC}, the proposal distribution
$q(\theta, \theta') \Pr(z \mid \theta')$ is replaced by
\[
  q_i(\theta, \theta') \prod_{k=1}^M \Pr(z_i \mid \theta').
\]