### 2016-04-27 doc/thesis/acknowledgements.tex
Thesis almost done

First and foremost I am deeply indebted to my supervisor, Dr. Art Poon, 
for mentoring me through the majority of my academic career.

I also thank my labmates, especially Jeff, Richard, and Thuy. I'll miss our
Thursday lunches.

I am grateful for the mentorship of Dr. Sara Mostafavi and Dr. Ryan Morin, who
accepted me into their labs and always had time for my many questions.  Thanks
also to Dr. Marni Mishna and Dr. Cedric Chauve, for starting me on this path.

I owe thanks to my fellow graduate students of the bioinformatics training
program. Special thanks go to Ogan for keeping me on an even keel when times
got tough.

Finally, I dedicate this thesis to my parents, who have supported me morally
and financially, endured my frustrations, and celebrated my successes.