### 2016-05-03 doc/figures/contactnet-empty.tex
GLM thingies

\documentclass[landscape]{article}
\usepackage{tikz}
\usepackage[T1]{fontenc}
\usepackage{mathptmx}
\usepackage{fullpage}

\usetikzlibrary{positioning}
\usetikzlibrary{calc}
\usetikzlibrary{arrows}
\usetikzlibrary{automata}

\graphicspath{{stock/}}

\begin{document}
\pagestyle{empty}

\begin{tikzpicture}
  [every node/.style = {inner sep=0pt},
   every path/.style = {->, >=stealth, very thick},
   label/.style = {anchor=south, color=white, inner sep=6pt},
   time/.style = {anchor=north west, inner sep=2pt}]
  \def\pht{1.5cm}

  \node at (-6, 5.5) (a) {\includegraphics[height=\pht]{person}};
  \node at (a.south) [label] {\huge a};
  \node [below left=of a] (b) {\includegraphics[height=\pht]{person}};
  \node at (b.south) [label] {\huge b};
  \node [below=of b] (e) {\includegraphics[height=\pht]{person}};
  \node at (e.south) [label] {\huge e};
  \node [below right=of a] (c) {\includegraphics[height=\pht]{person}};
  \node at (c.south) [label] {\huge c};
  \node [below=of c] (d) {\includegraphics[height=\pht]{person}};
  \node at (d.south) [label] {\huge d};

  \fill [shade, opacity=0.2] (a.east) -- (c.north) -- (c.north west) -- (a.south east) -- cycle;
  \fill [shade, opacity=0.2] (a.west) -- (b.north) -- (b.north east) -- (a.south west) -- cycle;
  \fill [shade, opacity=0.2] (a.south) -- (a.south west) -- (e.north) -- (e.north east) -- cycle;
  \fill [shade, opacity=0.2] (b.south west) -- (b.south) -- (e.north) -- (e.north west) -- cycle;
  \fill [shade, opacity=0.2] ($(c.south) + (-8pt, 0)$) -- ++(16pt, 0) -- ($(d.north) + (8pt, 0)$) -- ++(-16pt, 0) -- cycle;
  \fill [shade, opacity=0.2] (b.east) -- (b.south east) -- (d.west) -- (d.north west) -- cycle;
\end{tikzpicture}

\end{document}