### 2016-05-08 doc/thesis/preface.tex
Highlight new things

The initial idea to use kernel approximate Bayesian computation to infer
contact network model parameters was Dr. Poon's. The tree kernel was originally
developed by Dr. Poon, but the version used here was implemented by me to
improve computational efficiency. The idea to apply sequential Monte Carlo was
mine, but Dr. Alexandre Bouchard-C\^ot\'e made me aware of the adaptive version
used in this work. Dr. Sarah Otto suggested the experiments involving a network
with a heterogeneous $\alpha$ parameter and peer-driven sampling. Dr. Richard
H. Liang provided guidance in the development of the Gillespie simulation
algorithm and statistical advice. The \software{netabc} program, and all
supplementary analysis programs, were written by me.

A version of chapter 2 has been submitted for publication with the title
``Reconstructing network parameters from viral phylogenies.'' An oral
presentation entitled ``Phylodynamic inference of contact network parameters
with kernel-ABC'' was given based on chapter 2 to the 23rd HIV Dynamics and
Evolution meeting on April 25, 2016, in Woods Hole, Massachusetts, USA (the
presentation was delivered remotely). 
%A poster based on chapter 2 entitled
%``Likelihood-free estimation of contact network parameters from viral
%phylogenies'' was presented at the Intelligent Systems for Molecular Biology
%meeting in Orlando, Florida, USA.

Use of the BC data is in accordance with an ethics application that was
reviewed and approved by the UBC/Providence Health Care Research Ethics Board
(H07-02559).