### 2016-05-10 doc/thesis/acknowledgements.tex
Underline changes for printing

First and foremost I am deeply indebted to my supervisor, Dr. Art Poon, 
for mentoring me through the majority of my academic career.

I also thank my labmates, especially Jeff, Richard, and Thuy. I'll miss our
Thursday lunches.

I am grateful for the mentorship of my committee members Dr. Alexandre
Bouchard-C\^ot\'e, Dr. Richard Harrigan, and Dr. Sarah Otto. Thanks also to my
rotation supervisors Dr. Sara Mostafavi and Dr. Ryan Morin, who accepted me
into their labs and always had time for my many questions.

I owe thanks to my fellow graduate students of the bioinformatics training
program, especially to Ogan.

Finally, I dedicate this thesis to my parents, who have supported me morally
and financially, endured my frustrations, and celebrated my successes.