### 2016-05-21 doc/thesis/objective.tex

\glsreset{ABC}
\glsreset{BA}
\glsreset{SMC}

The spread of a disease is most often modelled by assuming either a
homogeneously mixed population~\autocite{hamer1906milroy,
kermack1927contribution}, or a population divided into a small number of
homogeneously mixed groups~\autocite{rushton1955deterministic}. This
assumption, also called \defn{mass
action}~\autocite{heesterbeek2000mathematical}, or \defn{panmixia}, implies
that any two individuals in the same compartment are equally likely to come
into contact making transmission possible at some predefined rate. Although
this provides a reasonable approximation in many
cases~\autocite{anderson1992infectious}, the error introduced by assuming a
panmictic population can be substantial when significant contact heterogeneity
exists in the underlying
population~\autocite{bansal2007individual,barthelemy2005dynamical,keeling2005networks}.
Contact network models provide an alternative to compartmental models which do
not require the assumption of panmixia. In addition to more accurate
predictions, the parameters of the networks themselves may be of interest from
a public health perspective. For example, certain vaccination strategies may be
more or less effective in curtailing an epidemic depending on the underlying
network's degree distribution~\autocite{peng2013vaccination, ma2013importance}.
Phylodynamic methods have been used to fit many different types of models to
phylogenetic data~\autocite{pybus2009evolutionary,volz2013viral}, but these
models generally assume a panmictic population. The primary objective of this
work is \emph{to develop a method to fit contact network models in a
phylodynamic framework.}

\newcommand{\G}{\mathcal{G}}
\newcommand{\Nu}{\mathcal{N}}

Calculating the likelihood of the parameters of a contact network models seems
likely to be an intractable problem. We have not proven this is the case, but
some intuition can be provided by examining the process involved in the
likelihood calculation. Consider a contact network model with parameters
$\theta$, and an estimated transmission tree $T$ with $n$ tips.
{\color{blue}\uline{The transmission tree is a record of an epidemic's progress
through a network. Tips of the transmission tree correspond to sampled
individuals, while internal nodes correspond to transmissions and may be
labelled with the transmitting individual. Transmission trees will be described
in more detail in the sequel, but for now they can be thought of as
representing the exact path an epidemic has taken. }} In general, we do not
know the labels of the internal nodes of $T$, only the labels of its tips. To
fit this model using likelihood-based methods, we must calculate the likelihood
of $\theta$, that is, $\Pr(T \mid \theta)$. Let $\G$ be the set of all possible
contact networks, and $\Nu$ be the set of all possible labellings of the
internal nodes of $T$. We can write the likelihood as
\begin{align}
\begin{split}
  \label{eq:netlik}
  \Pr(T \mid \theta)
    &= \sum_{\nu \in \Nu} \Pr(T, \nu \mid \theta) \\
    &= \sum_{G \in \G} \sum_{\nu \in \Nu} \Pr(T, \nu \mid G, \theta) \Pr(G \mid \theta) \\
    &= \sum_{G \in \G} \sum_{\nu \in \Nu} \Pr(T, \nu \mid G) \Pr(G \mid \theta),
\end{split}
\end{align}
the last equality following from the fact that $T$ and $\nu$ depend only on
$G$, not on $\theta$. Although $\Pr(T, \nu \mid G)$ and $\Pr(G \mid \theta)$
may individually be straightforward to calculate, the number of possible
directed graphs on $N$ nodes is $2^{N(N-1)}$~\autocite{harary2014graphical},
larger if the nodes and edges in the graph may have different labels or
attributes. Hence, the number of terms in the sum is at least exponential in
$n$, as there must be at least $n$ nodes in the network. In addition,
\cref{eq:netlik} assumes that $T$ is complete, meaning that all infected
individuals were sampled. This is rarely the case in practice - most often, we
only have access to a subset of the infected individuals. In this case, the
likelihood calculation becomes even more complex, because we must also sum over
all possible complete trees.

{\color{blue}\uline{
In a Bayesian setting, our goal is to estimate the posterior distribution on
model parameters given our data, $\Pr(\theta \mid T)$. Bayes' theorem relates
the posterior distribution to the likelihood,}
\[
  \Pr(\theta \mid T) = \frac{\Pr(T \mid \theta) \Pr(\theta)}{\Pr(T)}.
\]
\uline{For all but the simplest mathematical models, the calculation of the
normalizing constant $\Pr(T)$ is an intractable problem. What we have argued in
the previous paragraph is that calculating $\Pr(T \mid \theta)$ for a single
parameter candidate $\theta$ also seems likely to be computationally hard.
Therefore, the problem of fitting contact network models to phylogenies seems
to be of the \emph{doubly-intractible} type~\autocite{murray2012mcmc}, which
would imply that these models are not amenable to neither \gls{ML} nor Bayesian
inference techniques. Although both methods are able to cope with an
intractible normalizing constant (for example, by local search for \gls{ML} or
Bayesian \gls{MCMC}), neither can avoid the intractable likelihood
calculations. }}

Depending on the network model studied, it is possible that \cref{eq:netlik}
could be simplified into a tractable expression. However, a simpler alternative
to likelihood-based methods, which would apply to any network model, is
provided by \gls{ABC}~\autocite{rubin1984bayesianly, tavare1997inferring,
fu1997estimating, beaumont2002approximate}. All of the ingredients required to
apply \gls{ABC} to this problem are readily available. Simulating networks is
straightforward under a variety of models. Epidemics on those networks, and the
corresponding transmission trees, can also be easily simulated. As mentioned
above, contact networks can profoundly affect transmission tree shape, and
those shapes can be compared using a highly informative similarity measure
called the ``tree kernel''~\autocite{poon2013mapping}. \Gls{ABC} can be
implemented with \gls{SMC}, which has several advantages over other
algorithms~\autocite{mckinley2009inference}. A recently-developed adaptive
algorithm requiring minimal tuning on the part of the user makes \gls{SMC} an
even more attractive approach~\autocite{del2012adaptive}. In summary, our
method to infer contact network parameters will combine the following:
stochastic simulation of epidemics on networks, the tree kernel, and adaptive
\gls{ABC}-\gls{SMC}.

{\color{blue}\uline{
Our method will expand on the framework developed
by~\autocite{poon2015phylodynamic}, who combined \gls{ABC} with the tree kernel
to infer parameters of population genetic models from viral phylogenies using
\gls{MCMC}. The term ``kernel-assisted \gls{ABC}'' refers to any application of
\gls{ABC} using a kernel function as a distance metric, including that
developed by \citeauthor{poon2015phylodynamic}. However, for ease of
exposition, we will often use the term ``kernel-assisted \gls{ABC}'' to refer
to our method specifically. Note that this is distinct from the version of
\gls{ABC} based on kernel Bayes rule introduced by
\textcite{fukumizu2011kernel}.
}}

Empirical studies of sexual contact networks have found that these networks
tend to be scale-free~\autocite{colgate1989risk, liljeros2001web,
schneeberger2004scale,clemenccon2015statistical}, meaning that their degree
distributions follow a power law (although there has been some disagreement,
see \autocite{handcock2004likelihood, bansal2007individual}). Preferential
attachment has been postulated as a mechanism by which scale-free networks
could be generated~\autocite{barabasi1999emergence}. The \gls{BA}
model~\autocite{barabasi1999emergence} is one of the simplest preferential
attachment models, which makes it a natural choice to explore with our method.
The second aim of this work is \emph{to use simulations to investigate the
parameters of the \gls{BA} model, including whether they have a detectable
impact on tree shape, and whether they can be accurately recovered using
kernel-assisted \gls{ABC}.}

Due to its high global prevalence and fast mutation rate, \gls{HIV} is one of
the most commonly-studied viruses in a phylodynamic context. Consequently, a
large volume of \gls{HIV} sequence data is publicly available, more than for
any other pathogen, and including sequences sampled from diverse geographic and
demographic contexts. At the time of this writing, there were $635,400$ HIV
sequences publicly available in GenBank, annotated with 172 distinct countries
of origin. Since \gls{HIV} is almost always spread through either sexual
contact or sharing of injection drug supplies, the contact networks underlying
\gls{HIV} epidemics are driven by social dynamics and are therefore likely to
be highly structured~\autocite{clemenccon2015statistical}. Moreover, since no
cure yet exists, efforts to curtail the progression of an epidemic have relied
on preventing further transmissions through measures such as \gls{tasp} and
education leading to behaviour change. The effectiveness of this type of
intervention can vary significantly based on the underlying structure of the
network and the particular nodes to whom the intervention is
targeted~\autocite{little2014using,wang2015targeting}. Due to this combination
of data availability and potential public health impact, \gls{HIV} is an
obvious context in which our method could be applied. Therefore, the third and
final aim of this work is \emph{to apply kernel-assisted \gls{ABC} to fit the
\gls{BA} model to existing \gls{HIV} outbreaks}.

To summarize, this work has three objectives. First, we will develop a method
which uses kernel-assisted \gls{ABC} to infer parameters of contact network
models from observed transmission trees. Second, we will use simulations to
characterize the parameters of the \gls{BA} network model in terms of their
effect on tree shape and how accurately they can be recovered with
kernel-assisted \gls{ABC}. Finally, we will apply the method to fit the
\gls{BA} model to several real-world \gls{HIV} datasets.

\glsreset{ABC}
\glsreset{BA}
\glsreset{HIV}
\glsreset{ML}
\glsreset{SMC}