### 2016-06-09 doc/thesis/preamble.tex

\documentclass[11pt]{report}
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage[top=1in, bottom=1in, inner=1.25in, outer=1in]{geometry}
\usepackage{setspace}
\usepackage[backend=biber, style=numeric-comp, sorting=none]{biblatex}
\usepackage{nicefrac}
\usepackage{amsmath, amssymb}
\usepackage{algorithm}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{multicol}
\usepackage{multirow}
\usepackage[hidelinks]{hyperref}
\usepackage[acronym,nonumberlist]{glossaries}
\usepackage{glossary-mcols}
\usepackage{cleveref}
\usepackage{array}
\usepackage{textcomp}
\usepackage{float}
\usepackage{pdflscape}
\usepackage{fancyhdr}
\usepackage[dvipsnames]{xcolor}
\usepackage[titletoc]{appendix}
\usepackage{tocloft}
\usepackage[normalem]{ulem}

\newcommand{\comment}[1]{(\textbf{#1})}

\newcommand{\defn}[1]{\textit{#1}}
\newcommand{\software}[1]{\textit{#1}}
\newcommand{\etal}{\textit{et al.}}
\newcommand{\eg}{\textit{e.g.}}
\newcommand{\ie}{\textit{i.e.}}

\newcommand{\set}[1]{\left\lbrace#1\right\rbrace}
\newcommand{\sett}[1]{\{#1\}}
\renewcommand{\emptyset}{\varnothing}
\renewcommand{\vec}[1]{\mathbf{#1}}
\renewcommand{\star}[1]{#1^*}
\renewcommand{\d}{\mathrm{d}\,}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\maps}{:}

\newcommand{\N}{\mathcal{N}}
\newcommand{\I}{\mathbb{I}}
\newcommand{\F}{\mathcal{F}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\R}{\mathbb{R}}
\renewcommand{\L}{\mathcal{L}}

\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

\DeclareMathOperator{\Exponential}{Exponential}
\DeclareMathOperator{\Uniform}{Uniform}
\DeclareMathOperator{\DiscreteUniform}{DiscreteUniform}
\DeclareMathOperator{\Var}{Var}
\DeclareMathOperator{\inc}{in}
\DeclareMathOperator{\out}{out}
\DeclareMathOperator{\ESS}{ESS}
\DeclareMathOperator{\nc}{nc}
\DeclareMathOperator{\nl}{nl}

\newfloat{data}{hb}{lod}
\floatname{data}{Data}
\crefname{data}{data}{data}

\newacronym{MCMC}{MCMC}{Markov chain Monte Carlo}
\newacronym{ER}{ER}{Erd\H{o}s-R\'enyi}
\newacronym{BA}{BA}{Barab\'asi-Albert}
\newacronym{WS}{WS}{Watts-Strogatz}
\newacronym{SI}{SI}{susceptible-infected}
\newacronym{SVM}{SVM}{support vector machine}
\newacronym{SVR}{SVR}{support vector regression}
\newacronym{kSVM}{kSVM}{kernel support vector machine}
\newacronym{kSVR}{kSVR}{kernel support vector regression}
\newacronym{PCA}{PCA}{principal components analysis}
\newacronym{kPCA}{kPCA}{kernel principal components analysis}
\newacronym{SMC}{SMC}{sequential Monte Carlo}
\newacronym{ABC}{ABC}{approximate Bayesian computation}
\newacronym{nltt}{nLTT}{normalized lineages-through-time}
\newacronym{ltt}{LTT}{lineages-through-time}
\newacronym{pdf}{pdf}{probability density function}
\newacronym{rv}{rv}{random variable}
\newacronym{ML}{ML}{maximum likelihood}
\newacronym{MAP}{MAP}{maximum \textit{a posteriori}}
\newacronym{GSL}{GSL}{GNU scientific library}
\newacronym{SIR}{SIR}{susceptible-infected-recovered}
\newacronym{IS}{IS}{importance sampling}
\newacronym{SIS}{SIS}{sequential importance sampling}
\newacronym{HMM}{HMM}{hidden Markov model}
\newacronym{iid}{i.i.d.}{independent and identically distributed}
\newacronym{HPD}{HPD}{highest posterior density}
\newacronym{HDI}{HDI}{highest density interval}
\newacronym{HIV}{HIV}{human immunodeficiency virus}
\newacronym{SARS}{SARS}{severe acute respiratory syndrome}
\newacronym{ODE}{ODE}{ordinary differential equation}
\newacronym{GLM}{GLM}{generalized linear model}
\newacronym{tasp}{TasP}{treatment as prevention}
\newacronym{MH}{MH}{Metropolis-Hastings}
\newacronym{GTR}{GTR}{generalized time-reversible}
\newacronym{ESS}{ESS}{expected sample size}
\newacronym{MSM}{MSM}{men who have sex with men}
\newacronym{ERGM}{ERGM}{exponential random graph model}
\newacronym{IDU}{IDU}{injection drug users}
\newacronym{ANOVA}{ANOVA}{analysis of variance}
\newacronym{MANOVA}{MANOVA}{multivariate analysis of variance}
\newacronym{AIC}{AIC}{Akaike information criterion}
\newacronym{POSIX}{POSIX}{Portable Operating System Interface}
\newacronym{IQR}{IQR}{interquartile range}
\newacronym{MPI}{MPI}{message passing interface}
\newacronym{PA}{PA}{preferential attachment}

\newglossaryentry{alpha}{
  name=\ensuremath{\alpha},
  description=preferential attachment power parameter in Barab\'asi-Albert networks
}
\newglossaryentry{m}{
  name=\ensuremath{m},
  description=number of edges added per vertex when constructing a Barab\'asi-Albert network
}
\newglossaryentry{N}{
  name=\ensuremath{N},
  description=total number of nodes in a contact network
}
\newglossaryentry{I}{
  name=\ensuremath{I},
  description=number of infected nodes in a contact network at the time of
  transmission tree sampling
}
\newglossaryentry{gamma}{
  name=$\gamma$,
  description=exponent of power-law degree distribution in scale-free networks
}
\newglossaryentry{p}{
  name=\ensuremath{p},
  description=edge density parameter in Erd\H{o}s-R\'enyi networks
}
\newglossaryentry{lambda}{
  name=\ensuremath{\lambda},
  description=decay factor meta-parameter for tree kernel
}
\newglossaryentry{sigma}{
  name=\ensuremath{\sigma},
  description=radial basis function variance meta-parameter for tree kernel
}
\newglossaryentry{n}{
  name=\ensuremath{n},
  description=number of particles used for ABC-SMC
}
\newglossaryentry{M}{
  name=\ensuremath{M},
  description=number of simulated datasets per particle in ABC-SMC
}
\newglossaryentry{alphaess}{
  name=\ensuremath{\alpha_{\ESS}},
  description=per-iteration rate of expected sample size decay in adaptive ABC-SMC
}

\graphicspath{{../figures/}}
\newcommand{\tablepath}{../tables}
\frenchspacing
\onehalfspacing

\pagestyle{fancyplain}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\rfoot{\thepage}

\renewcommand\UrlFont{\color{Blue}\rmfamily}

\title{Phylodynamic inference of contact network parameters through approximate
Bayesian computation}
\author{Rosemary M. McCloskey and Art F.Y. Poon}

\addbibresource{papers.bib}
\makeglossaries