### 2016-06-16 doc/thesis/methods_results.tex

\add{In this chapter, we will address the three research aims
of this thesis introduced in \cref{sec:obj}. First, in \cref{sec:netabc}, we
describe \software{netabc}, a computer program that implements an
\acrlong{ABC}-based algorithm to fit contact network models to phylogenetic
data. We also provide a justification for the use of \gls{ABC} for this problem
by arguing that the likelihood functions required to fit these models by more
conventional means are likely to be computationally intractable.  Second, in
\cref{sec:ba}, we perform a simulation study to investigate the \acrlong{BA}
network model, which uses a preferential attachment mechanism to generate
networks with the power law degree distributions observed in real world social
and sexual networks. We progress through two exploratory analyses testing the
identifiability of the model's parameters, and conclude by testing
\software{netabc}'s ability to recover the parameters from simulated
transmission trees. Third, in \cref{sec:hiv}, we apply \software{netabc} to fit
the \gls{BA} model to six real world \gls{HIV} datasets, with the understanding
of the model parameters' identifiability gained through the simulation
experiments. We conclude the chapter with a unified discussion of the three
research aims, including interpretation of the results of both the simulated
and real data experiments, as well as an examination of the limitations of our
approach and opportunities for future investigation.}

\section{\software{Netabc}: a computer program for estimation of contact
network parameters with ABC}
\label{sec:netabc}

\software{Netabc} is a computer program to perform statistical inference of
contact network parameters from an estimated transmission tree using
\gls{ABC}. \add{As discussed in \cref{sec:obj}, the principal statistical
algorithm used by \software{netabc} is adaptive
\gls{ABC}-\gls{SMC}~\autocite{del2012adaptive}. In addition, there are two
supplementary components which are specific to the domain of phylogenetics and
contact networks: Gillespie simulation~\autocite{gillespie1976general}, to
simulate transmission trees on contact networks; and the tree
kernel~\autocite{poon2013mapping}, which is used as the distance function in
\gls{ABC} to compare transmission trees~\autocite{poon2015phylodynamic} (see
\cref{sec:abc}).} We give a high-level overview of the program here, before
describing these components in detail. \software{Netabc} takes as input an
estimated transmission tree, which can be derived from a viral phylogeny by
rooting and time-scaling as described in \cref{subsec:phylodynamics} or
estimated by other methods~\autocite{cottam2008integrating,
jombart2011reconstructing, ypma2012unravelling, morelli2012bayesian,
didelot2014bayesian, hall2015epidemic}. We variously refer to this estimated
transmission tree as the observed tree, input tree, or true tree.

As described in \cref{sec:smc}, \software{netabc} keeps track of a population
of particles $x^{(k)}$ indexed by an integere $k$, each of which contains
particular parameter values $\theta^{(k)}$ for the \add{contact network} model
we are trying to fit \add{to the input tree}. A small number \gls{M} of contact
networks $z^{(k,i)}$, $1 \leq i \leq \gls{M}$, are generated under the model
for each particle in accordance with that particle's parameters. An epidemic is
simulated over each of these networks using Gillespie simulation, and by
keeping track of its progress, a transmission tree is obtained. Thus, each
particle becomes associated with several simulated transmission trees. These
trees are compared to the input tree using the tree kernel. Particles are
weighted according to the similarity of their associated simulated trees with
the true tree, with more similar trees receiving higher weights. The particles
are iteratively perturbed to explore the parameter space, and particles with
simulated trees too distant from the true tree are periodically dropped and
resampled. Once a convergence criterion is attained, the final set of particles
is used as a Monte Carlo approximation to the target distribution of \gls{ABC},
which is assumed to resemble the posterior distribution on model parameters
(see \cref{sec:abc}). A graphical schematic of this algorithm is shown in
\cref{fig:abcsmc}.

\begin{figure}
    \includegraphics{abc-smc.pdf}
    \caption[
        Graphical schematic of the ABC-SMC algorithm implemented in \software{netabc}.
    ]{
        Graphical schematic of the \gls{ABC}-\gls{SMC} algorithm implemented in
        \software{netabc}. Particles are initially drawn from their prior
        distributions, making the initial population a Monte Carlo
        approximation to the prior. At each iteration, particles are perturbed,
        and a distance threshold around the true tree contracts. Particles are
        rejected, and eventually resampled, when all their associated simulated
        trees lie outside the threshold. As the algorithm progresses, the
        population smoothly approaches a Monte Carlo approximation of the
        \gls{ABC} target distribution, which is assumed to resemble the
        posterior.
    }
    \label{fig:abcsmc}
\end{figure}

\software{Netabc} is written in the \software{C} programming language. The
\software{igraph} library~\autocite{csardi2006igraph} is used to generate and
store contact networks and phylogenies. Judy arrays~\autocite{baskins2004judy}
are used for hash tables and dynamic programming matrices. The
\gls{GSL}~\autocite{gough2009gnu} is used to generate random draws from
probability distributions, and to \del{perform the bisection step} \add{solve
for the next $\varepsilon$ by bisection} in the adaptive \gls{ABC}-\gls{SMC}
algorithm. Parallelization is implemented with \gls{POSIX}
threads~\autocite{barney2009posix}. In addition to the \software{netabc} binary
to perform \gls{ABC}, we provide three additional stand-alone utilities:
\software{treekernel}, to calculate the tree kernel; \software{nettree}, to
simulate a transmission tree over a contact network; and \software{treestat},
to compute various summary statistics of phylogenies. The programs are freely
available at \url{https://github.com/rmcclosk/netabc}.

To check that our implementation of Gillespie simulation was correct, we
reproduced Figure 1A of \textcite{leventhal2012inferring} (our
\cref{fig:leventhal}), which plots the imbalance of transmission trees
simulated over four network models at various levels of pathogen
transmissibility. Our implementation of adaptive \gls{ABC}-\gls{SMC} was tested
by applying it to the same mixture of Gaussians used by
\textcite{del2012adaptive} to demonstrate their method (originally used
by~\textcite{sisson2007sequential}). We were able to obtain a close
approximation to the function (see \cref{fig:smctest}), and attained the
stopping condition used by the authors in a comparable number of steps. To
check that the algorithm would converge to a bimodal distribution, we also
applied it to a mixture of two Gaussians with means $\pm4$ and variances 1. The
algorithm was able to recover both peaks (\cref{fig:smctest2}).

\subsection{Simulation of transmission trees over contact networks}
\label{subsec:nettree}

The simulation of epidemics, and the corresponding transmission trees, over
contact networks is performed in \software{netabc} using the Gillespie
simulation algorithm~\autocite{gillespie1976general}. This method has been
independently implemented and applied by several
authors~\autocite[\textit{e.g.}][]{o2011contact, robinson2013dynamics,
leventhal2012inferring, groendyke2011bayesian, villandre2016assessment}.
\textcite{groendyke2011bayesian} published their implementation as an
\software{R} package, but since the \gls{SMC} algorithm is quite
computationally intensive, we chose to implement our own version in
\software{C} \add{as part of \software{netabc}}.

Let $G = (V, E)$ be a directed contact network. We assume the individual nodes
and edges of $G$ follow the dynamics of the \gls{SIR}
model~\autocite{kermack1927contribution}. Each directed edge $e = (u, v)$ in
the network is associated with a transmission rate $\beta_e$, which indicates
that, once $u$ becomes infected, the waiting time until $u$ infects $v$ is
distributed as $\Exponential(\gls{beta}_e)$. Note that $v$ may become infected
before this time has elapsed, if $v$ has other incoming edges. $v$ also has a
removal rate $\gls{nu}_v$, so that the waiting time until removal of $v$ from
the population is $\Exponential(\gls{nu}_v)$. Removal may correspond to death
or recovery with immunity, or a combination of both, but in our implementation
recovered nodes never re-enter the susceptible population. We define a
\defn{discordant edge} as an edge $(u, v)$ where $u$ is infected and $v$ has
never been infected. \add{In the epidemiology literature, the symbol $\gamma$
is usually used in place of \gls{nu}; we use \gls{nu} here to distinguish the
recovery rate from the power law exponent of scale free networks (see
\cref{subsec:pa}).}

To describe the algorithm, we introduce some notation and variables. Let
$\inc(v)$ be the set of incoming edges to $v$, and $\out(v)$ be the set of
outgoing edges from $v$. Let \gls{gI} be the set of infected nodes in the
network, \gls{gR} be the set of removed nodes, \gls{gS} be the set of
susceptible nodes, and \gls{gD} be the set of discordant edges in the network.
Let \gls{beta} be the total transmission rate over all discordant edges, and
\gls{nu} be the total removal rate of all infected nodes,
\[
    \gls{beta} = \sum_{e \in \gls{gD}} \gls{beta}_e, \quad
    \gls{nu} = \sum_{v \in \gls{gI}} \gls{nu}_v.
\]
The variables \gls{gS}, \gls{gI}, \gls{gR}, \gls{gD}, \gls{beta}, and \gls{nu}
are all updated as the simulation progresses. When a node $v$ becomes infected,
it is deleted from \gls{gS} and added to \gls{gI}. Any formerly discordant
edges in $\inc(v)$ are deleted from \gls{gD}, and edges in $\out(v)$ to nodes
in \gls{gS} are added to \gls{gD}. If $v$ is later removed, it is deleted from
\gls{gI} and added to \gls{gR}, and any discordant edges in $\out(v)$ are
deleted from \gls{gD}. At the time of either infection or removal, the
variables \gls{beta} and \gls{nu} are updated to reflect the changes in the
network. \del{The updates to \gls{gS}, \gls{gI}, \gls{gR}, \gls{gD},
\gls{beta}, and \gls{nu} are straightforward and are not written explicitly in
the algorithm.}

\newcommand{\tip}{\mathit{tip}}

The Gillespie simulation algorithm is given as \cref{alg:nettree}. The
transmission tree \gls{T} is simulated along with the epidemic. We keep a map
called ``$\tip$'', which maps infected nodes in \gls{gI} to the tips of
\gls{T}. The simulation continues until either there are no discordant edges
left in the network, or we reach a user-defined cutoff of time (\gls{tmax}) or
number of infections (\gls{I}). We use the notation $\Uniform(0, 1)$ to
indicate a number drawn from a uniform distribution on $(0, 1)$, and
$\Exponential(\lambda)$ to indicate a number drawn from an exponential
distribution with rate $\lambda$. The combined number of internal nodes and
tips in \gls{T} is denoted $|T|$. \add{The updates to \gls{gS}, \gls{gI},
\gls{gR}, \gls{gD}, \gls{beta}, and \gls{nu} described in the previous
paragraph are not written explicitly in \cref{alg:nettree}, as they are quite
straightforward and would only obfuscate the pseudocode.}

\begin{algorithm}
    \label{alg:nettree}
    \caption{Simulation of an epidemic and transmission tree over a contact network}
    \begin{algorithmic}
        \State infect a node $v$ at random, updating \gls{gS}, \gls{gI},
                 \gls{gD}, \gls{beta}, and \gls{gamma}
        \State $\gls{T} \gets$ a single node with label $1$
        \State $\tip[v] \gets 1$
        \State $\gls{t} \gets 0$
        \While{$\gls{gD} \neq \emptyset$ and $|\gls{gI}| + |\gls{gR}| < \gls{I}$
                and $\gls{t} < \gls{tmax}$}
            \State $s \gets \min(\gls{tmax} - \gls{t},\,
                \Exponential(\gls{beta} + \gls{nu}))$
            \For{$v \in \tip$}
                \State{extend the branch length of $\tip[v]$ by $s$}
            \EndFor
            \State $\gls{t} \gets \gls{t} + s$
            \If{$\gls{t} < \gls{tmax}$}
                \If{$\Uniform(0, \gls{beta} + \gls{nu}) < \gls{beta}$}
                    \State choose an edge $e = (u, v)$ from \gls{gD} with
                        probability $\gls{beta}_e / \gls{beta}$ and infect $v$
                    \State $\tip[v] \gets |\gls{T}|+1$
                    \Comment add new tips to tree and tip array
                    \State $\tip[u] \gets |\gls{T}|+2$
                    \Comment corresponding to $u$ and $v$ 
                    \State add tips with labels $(|\gls{T}|+1)$ and
                        $(|\gls{T}|+2)$ to \gls{T}
                    \State connect the new nodes to $\tip[v]$ in $\gls{T}$,
                        with branch lengths $0$ 
                \Else
                    \State choose a node $v$ from \gls{gI} with probability
                        $\gls{nu}_v / \gls{nu}$ and remove $v$
                    \State delete $v$ from $\tip$
                \EndIf
                \State update \gls{gS}, \gls{gI}, \gls{gR}, \gls{gD},
                    \gls{beta}, and \gls{nu}
            \EndIf
        \EndWhile
    \end{algorithmic}
\end{algorithm}

\subsection{Phylogenetic kernel}

The tree kernel developed by \textcite{poon2013mapping} provides a
comprehensive similarity score between two phylogenetic trees, via the
dot-product of the two trees' feature vectors in the space of all possible
subset trees with branch lengths (see \cref{subsec:treeshape}). \add{Because
the branch lengths are continuous, there are infinitely many possible subset
trees; hence, the feature space is infinite-dimensional.} The kernel was
implemented using the fast algorithm developed by
\textcite{moschitti2006making}. First, the production rule of each node, which
is the total number of children and the number of leaf children, is recorded.
The nodes of both trees are ordered by production rule, and a list of pairs of
nodes sharing the same production rule is created. These are the nodes for
which the value of the tree kernel must be computed - all other pairs have a
value of zero. The pairs to be compared are then re-ordered so that the child
nodes are always evaluated before their parents. Due to its recursive
definition, ordering the pairs in this way allows the tree kernel to be
computed by dynamic programming. The complexity of this implementation is
$O(|\gls{T}_1||\gls{T}_2|)$ for the two trees $\gls{T}_1$ and $\gls{T}_2$ being
compared.

The tree kernel cannot be used directly as a distance measure for \gls{ABC},
since it is maximized, not minimized, when the two trees being compared are the
same. Therefore, we defined the distance between two trees as
\[
    \gls{rho}(\gls{T}_1, \gls{T}_2) = 1 - \frac{K(\gls{T}_1, \gls{T}_2)}{\sqrt{K(\gls{T}_1, \gls{T}_1) K(\gls{T}_2, \gls{T}_2)}},
\]
which is a number between 0 and 1 that is minimized when $\gls{T}_1 = \gls{T}_2$. This is
similar to the normalization used by \textcite{collins2002new,
poon2013mapping}.

\subsection{Adaptive sequential Monte Carlo for Approximate Bayesian computation}
\label{subsec:adaptsmc}

We implemented the adaptive \gls{SMC} algorithm for \gls{ABC} developed by
\textcite{del2012adaptive}. This algorithm is similar to the reference
\gls{ABC}-\gls{SMC} algorithm described in \cref{subsec:abcalg}, except that
the sequence of tolerances $\gls{epsilon}_i$ is automatically determined rather
than specified in advance. The tolerances are chosen such that the \gls{ESS} of
the particle population, which indicates the quality of the Monte Carlo
approximation (see \cref{subsec:sis}), decays at a controlled rate. A
sudden precipitous drop in \gls{ESS} would indicate that only a small number of
particles had non-zero importance weights, which would result in a very poor
Monte Carlo approximation to the target distribution. This situation is
referred to as the collapse of the approximation \add{or particle degeneracy
(see \cref{subsec:smc})} and is mitigated by the adaptive approach. A single
parameter \del{$\alpha$ (not to be confused with the \gls{BA} model parameter)}
controls the decay rate. \add{In the original paper of
\textcite{del2012adaptive}, the parameter is called $\alpha$, but to avoid
confusion with the \gls{BA} parameter of the same name we will refer to it
here as \gls{alphaess}.} The tolerance $\gls{epsilon}_i$ is chosen to satisfy
\[
    \ESS(w_i) = \gls{alphaess} \ESS(w_{i-1}),
\]
where, $w_i$ is the vector of weights at the $i$th step. Note that, since $w_i$
depends on $\gls{epsilon}_i$, this equation solves for the updated weights and
the updated tolerance simultaneously. As pointed out by
\textcite{del2012adaptive}, the equation has no analytic solution, but can be
solved numerically by bisection. The forward kernels $K_i$ are taken to be
\gls{MCMC} kernels with stationary distributions $\pi_{\gls{epsilon}_i}$ and
proposal distributions
\[
    \gls{q}_i\left(\theta^{(k)}, \theta^{(k)\prime}\right) 
    \prod_{j=1}^{\gls{M}} \lik\left(z^{(j,k)\prime} \middle| \theta^{(k)\prime} \right),
\]
where $\theta^{(k)}$ is the vector of model parameters \add{associated with
particle $x^{(k)}$} and $z^{(j,k)\prime}$, $1 \leq j \leq \gls{M}$, are \gls{M}
datasets simulated according to $\theta^{(k)\prime}$. In our implementation,
the $\gls{q}_i$ \add{are constructed component-wise for $\theta$ out of}
Gaussian proposals for continuous parameters and Poisson proposals for discrete
parameters. For the Poisson proposals, the number of \add{discrete} steps to
move the particle is drawn from a Poisson distribution, and the direction in
which to move the particle is chosen uniformly at random. The variance of each
proposal distribution was set equal to twice the empirical variance of the
particles, following~\autocite{beaumont2009adaptive, del2012adaptive}. The
backwards kernels are
\[
    L_{i-1}(x', x) = \frac{\pi_i(x)K_i(x, x')}{\pi_i(x')}.
\]
Here we have written $x'$ for $x_i$ and $x$ for $x_{i-1}$ to emphasize that
$x_{i-1}$ is the current value of the particle and $x_i$ is the proposed value.
When substituted into \cref{eq:smcwt}, the forward kernels $K_i(x, x')$ and
densities $\pi_i(x') = \pi_{\varepsilon_i}(x')$ cancel out, and we are left
with the following weight update.
\begin{align*}
    w_i(x) 
    & \madd{\propto w_{i-1}(x) \frac{\pi_i(x') L_{i-1}(x', x)}{\pi_{i-1}(x)K_i(x, x')}}
        & \madd{\text{importance weight update from SMC}} \\
    &\madd{=w_{i-1}(x) \frac{\pi_i(x') \pi_i(x) K_i(x, x')}{\pi_{i-1}(x)K_i(x, x')\pi_i(x')}}
        & \madd{\text{substitute } L_{i-1}} \\
    &= w_{i-1}(x) \frac{\pi_i(x)}{\pi_{i-1}(x)}
        & \madd{\text{cancel } K_i(x, x') \text{ and } \pi_i(x')} \\
    &= \madd{w_{i-1}(x) \frac{\pi(\theta) \prod_{j=1}^M f(z^{(j)\prime} \mid \theta) 
                        \sum_{j=i}^M \I_{A_{\varepsilon_i, y}}(z^{(j)})}
                       {\pi(\theta) \prod_{j=1}^M f(z^{(j)\prime} \mid \theta) 
                       \sum_{j=i}^M \I_{A_{\varepsilon_{i-1}, y}}(z^{(j)})}}
        & \madd{\text{definition of } \pi_i(x)} \\
    &= w_{i-1}(x) \frac{\sum_{j=i}^M \I_{A_{\varepsilon_i, y}}(z^{(j)})}
        {\sum_{j=i}^M \I_{A_{\varepsilon_{i-1}, y}}(z^{(j)})}
        & \madd{\text{cancel prior and likelihood.}}
\end{align*}
In other words, when the distance threshold $\gls{epsilon}_{i-1}$ is contracted
to $\gls{epsilon}_i$, the particles' weights are multiplied by the proportion of
simulated datasets that are still inside the new threshold. The user may
specify a final tolerance $\gls{epsilon}$, or a final acceptance rate of the
\gls{MCMC} kernel, and the algorithm will be stopped when either of these
termination conditions is reached. The latter condition stops the algorithm
when the particles are not moving around very much, implying little change in
the estimated target.

\subsection{Justification for approach}
\label{subsec:just}

\add{We present here a non-rigorous justification for the use of
\gls{ABC} for the problem at hand, as opposed to more frequently-used
approaches for fitting mathematical models (see \cref{chp:prelim}). Consider a
contact network model with parameters $\theta$, and an estimated transmission
tree \gls{T}. Taking a Bayesian approach, our aim is to obtain a sample from
the posterior distribution on the model's parameters given our data,}
\[
    \madd{\post(\theta \mid \gls{T}) = \frac{\lik(\gls{T} \mid \theta) \prior(\theta)}
    {\int_{\Theta}\lik(\gls{T} \mid \theta) \prior(\theta) \d\theta}.}
\]
\add{For all but the simplest models, the normalizing constant in the
denominator is an intractable integral. What we shall argue here is that, in
contrast to most commonly studied mathematical models, the likelihood
$\lik(\gls{T} \mid \theta)$ is also likely to be intractible in our case.}

\begin{figure}
    \centering
    \includegraphics[width=0.6\textwidth]{transtree}
    \caption[
        Illustration of an estimated transmission tree without labels and two
        possible underlying complete transmission trees with labels.
    ]{
        Illustration of an estimated transmission tree without labels (left)
        and two possible underlying complete transmission trees with labels
        (right). In the top right scenario, the epidemic began with node $a$
        who transmitted first to $b$ and then to $c$. In the bottom right
        scenario, $b$ was the index case; $b$ infected $c$, who went on to
        infect $a$. A transmission tree estimated from a viral phylogeny would
        have the same topology and tip labels in both cases.
    }
    \label{fig:tt}
\end{figure}

\add{As discussed in \cref{subsec:tt}, the internal nodes of transmission trees
represent transmission events, and are labelled with the donor in the
associated transmission pair. However, when we estimate a transmission tree
from viral sequence data, we generally only know the labels of the tips of the
tree, not the labels of the internal nodes. In viral phylogenies, the
transmissions are at least partially preserved through the evolutionary
relationships among the viruses, but the directionality of those transmissions
is unknown. Thus, a single estimated transmission tree can correspond to many
possible pathways of the epidemic through the network. \Cref{fig:tt}
illustrates this concept for a simple transmission tree with three tips. When
calculating a likelihood given a transmission tree, we must sum over all
possible labellings of the internal nodes. Let $\L$ be the set of such
labellings. Then}
\begin{align}
    \madd{\lik(\gls{T} \mid \theta) = \sum_{l \in \L} \lik(\gls{T},\, l \mid \theta).}
    \label{eq:lik1}
\end{align}
\add{A contact network model assigns a probability density to each possible
contact network. Transmission trees are realized over particular contact
networks, not over the model itself. Therefore, we must also sum over all
contact networks which could be generated by the model. Let $\G$ be the set of
all possible contact networks. Summing \cref{eq:lik1} over $\G$ gives}
\begin{align}
    \madd{\lik(\gls{T} \mid \theta) = \sum_{G \in \G} \sum_{l \in \L} 
        \lik(\gls{T},\, l \mid G, \theta) \lik(G \mid \theta).}
    \label{eq:lik2}
\end{align}
\add{This can be simplified somewhat by noticing that, given a specific
contact network, the labelled transmission tree depends only on that network
and not on the model that generated it. That is, $\lik(\gls{T}, \nu \mid G,
\theta) = \lik(\gls{T},\, l \mid G)$, and}
\begin{align}
    \madd{\lik(\gls{T} \mid \theta) = \sum_{G \in \G} \lik(G \mid \theta) \sum_{l \in \L} 
    \lik(\gls{T}, \nu \mid G)}
    \label{eq:lik3}
\end{align}
\add{Under the assumption that both transmission and removal are Poisson
processes, calculating $\lik(\gls{T},\, l \mid G)$ can be accomplished by a
straightforward modification of the Gillespie simulation algorithm
(\cref{alg:nettree}). Rather than choosing transmission or removal events
according to their probabilities, the events would be deterministically chosen
based on the transmission tree and the probabilities of each event would be
multiplied together. Assuming efficient data structures for storing lists of
nodes and edges, the complexity of this calculation would be $O(|\gls{T}|)$.
The number of possible labellings of internal nodes of $\gls{T}$ is easily seen
to be $2^{(|\gls{T}|-1)/2}$ by noticing that each of the $(|\gls{T}|-1)/2$
internal nodes must be labelled with the same label as either its right child
or its left child. Although exponential calculations of this nature can are
often be simplified on trees using dynamic programming (\eg
\autocite{pupko2000fast}), it cannot be straightforwardly applied in this case
because the subtrees' probabilities depend on the existing epidemic progress
(their parents and siblings). Hence, calculating the inner sum over labels may
take time $O(2^{(|\gls{T}|-1)/2})$.}

\add{The outer sum, over all contact networks, is also difficult to evaluate
in general. There are $2^{\gls{N}(\gls{N}-1)}$ directed graphs on \gls{N}
nodes~\autocite{harary2014graphical}. There must be at least as many nodes in
the contact network as the number of tips in the tree, which is
$(|\gls{T}|+1)/2$. Of course, it is very likely that there are more nodes in
the network than observed tips because some individuals are never infected
and/or some infected individuals are never sampled. The complexity of
calculating $\lik(G \mid \theta)$ is obviously dependent on the particular
model being investigated. For the \gls{BA} model, we might have to sum over all
possible orders in which the nodes could be added, and all assignments of edges
to the nodes which generated them. However, even in the case that calculating
$\lik(G \mid \theta)$ can be done in constant time, the sum (\ref{eq:lik3})
still has at least $O(2^{|\gls{T}|^2})$ terms.}

\add{We have shown that both the normalizing constant $\int_\Theta \lik(\gls{T}
\mid \theta) \prior(\theta) \d\theta$ and the likelihood $\lik(T \mid \theta)$
are likely computationally prohibitive to calculate. If this is the case, the
problem of fitting contact network models to phylogenies seems to be of the
\emph{doubly-intractible} type~\autocite{murray2012mcmc}, which would imply that
these models are not amenable to neither \gls{ML} nor Bayesian inference
techniques. Although both methods are able to cope with an intractible
normalizing constant (for example, by local search for \gls{ML} or Bayesian
\gls{MCMC}), neither can avoid the intractable likelihood calculations. This
justifies the use of \gls{ABC}, which is a likelihood-free method.}

\add{We have not proven here that \cref{eq:lik3} is impossible to calculate in
polynomial time - it could be possible to algebraically simplify the sum into a
tractable expression. Furthermore, under certain models, a large proportion of
$\G$ may have zero probability, which would enable the simplification of the
outer sum on a model-specific basis. It should also be noted that extensions of
Bayesian \gls{MCMC} have been developed for doubly-intractable
problems~\autocite{liang2010double, murray2012mcmc}, which might be adaptable
to the problem at hand. These have not been as widely used as \gls{ABC}, nor
are they as easily parallelizable as \gls{SMC}.}

\section{Analysis of \acrlong{BA} model with synthetic data}
\label{sec:ba}
\glsreset{BA}

\subsection{Why study the \acrlong{BA} model?}

\add{We developed \software{netabc} with the objective of extracting useful,
quantitative information about network structures from viral phylogenies. An
important aspect of ``usefulness'' is model specification and the biological or
epidemiological interpretation of the parameters. We want the model to be
realistic, but no so complicated that it becomes difficult to interpret. At
least some of the parameters should be of interest from a theoretical or
practical perspective, or there would be no point in estimating them. Since
\software{netabc} is a phylodynamic method, intended to be used with viral
sequence data, we would also like to choose parameters which may be difficult
to estimate with more standard methods. Otherwise, our method provides no
advantage. The \gls{BA} model (\cref{subsec:pa}) satisfies these criteria,
albeit some better than others. The purported realism of the model stems from
the fat-tailed degree distributions it produces, which are similar to those
observed in real world sexual networks~\autocite{colgate1989risk,
liljeros2001web, schneeberger2004scale, clemenccon2015statistical,
rothenberg2007large}. Moreover, the ``rich get richer'' phenomenon, where
popular individuals attract new connections at an elevated rate, is intuitively
reasonable for both sexual~\autocite{de2007preferential} and
\gls{IDU}~\autocite{dombrowski2013topological} networks. However, the model is
very simple, assuming that all nodes form the same number of links when added
to the network and share the same preference for popular individuals.}

\add{In this thesis, we consider four parameters related to the \gls{BA} model,
denoted \gls{N}, \gls{m}, \gls{alpha}, and \gls{I} (see \cref{subsec:pa}). The
first three of these parameterize the network structure, while \gls{I} is
related to the simulation of transmission trees over the network. However, we
will refer to all four as \gls{BA} parameters. \gls{N} denotes the total number
of nodes in the network, or equivalently, susceptible individuals in the
population. \gls{m} is the number of new undirected edges added for each new
vertex, or equivalently one-half of the average degree. \gls{alpha} is the
power of preferential attachment -- new nodes are attached to existing nodes of
degree $d$ with probability proportional to $d^{\gls{alpha}} + 1$. Finally,
\gls{I} is the number of infected individuals at the time when sampling occurs.
The \gls{alpha} parameter is unitless, while \gls{m} has units of edges or
connections per vertex, and \gls{N} and \gls{I} both have units of nodes or
individuals.}

\add{From a public health standpoint, all four parameters are of some interest.
The prevalence \gls{I} can be used to estimate the resources required to combat
an ongoing epidemic, while total susceptible population size \gls{N} provides a
similar metric for preventative measures. The average degree of the network, in
this case $2\gls{m}$, is directly related to \gls{R0}, the basic reproductive
number~\autocite{britton2002bayesian}. \gls{R0} quantifies the number of
secondary infections ultimately caused by one infected individual; higher
\gls{R0} generally indicates faster epidemic growth and/or larger eventual
epidemic size~\autocite{anderson1992infectious}. In a homogeneously mixed
population, the proportion of the population which must be vaccinated to
control an epidemic can be expressed in terms of
\gls{R0}~\autocite{anderson1982directly}. Although optimal vaccination
strategies may differ in heterogeneous contact
structures~\autocite{ma2013importance}, it is reasonable to suppose that there
would still be a relationship between \gls{m}, \gls{R0}, and the vaccination
threshold. The preferential attachment power \gls{alpha} quantifies the degree
to which high-degree nodes, also called
superspreaders~\autocite{kemper1980identification}, characterize the network
structure. Superspreaders have been hypothesized to play a greater-than-average
role in the spread of several diseases~\autocite{stadler2013uncovering,
shen2004superspreading}. If so, network-based
interventions~\autocite{little2014using, wang2015targeting} may be worth
considering as part of an epidemic control strategy. \gls{alpha} can also offer
some insight into how the network would react to the removal of nodes.
\textcite{dombrowski2013topological} found evidence of preferential attachment
in \gls{IDU} networks, and suggested that as a consequence of this
characteristic, the removal of random nodes (such as through a police
crackdown) might inadventently make it easier for epidemics to spread. When
individuals with only one or two connections lose them, they might tend to seek
out well-known (that is, high-degree) members of the community, thus increasing
those individuals' connectivity even further.}

\add{All four \gls{BA} parameters can be estimated using more conventional
approaches, but this estimation may be challenging depending on characteristics
of the epidemic under consideration. In theory, any network parameter can be
estimated by explicitly constructing the contact network, although this is
highly resource intensive and is hampered by misreporting and other
challenges~\autocite{eames2015six}. All parameters become more difficult to
estimate when the infected population is ``hidden'' due to illegal or
stigmatized behaviour, as is sometimes the case with \gls{HIV} outbreaks among
\gls{IDU}, \gls{MSM}, or sex workers. Of course, phylodynamic methods are
equally ineffective when the population is completely hidden, since we must at
least be able to sample viral sequences from its members. If a population is
accessible, \gls{N} and \gls{I} may be estimable by surveying and/or testing
each member of the population for the disease. However, this will not tell us
if there are large compartments of the population that we simply have not
sampled. Our hope is that estimating \gls{N} and \gls{I} phylogenetically might
provide this additional information. The average degree of the network,
$2\gls{m}$, is also estimable by a survey, although individuals may be
unwilling or unable to disclose how many contacts they have had. The estimation
of \gls{alpha} is more complex, as this parameter is most strongly reflected in
the connectivity of very high degree nodes, who are rare in the population.
Locating them might require contact tracing, or respondent-driven
sampling~\autocite{heckathorn1997respondent}. Even if the full degree
distribution of a network is available, there are models other than
preferential attachment which can produce scale-free
networks~\autocite[\eg][]{kumar2000stochastic}. \textcite{de2007preferential}
were able to estimate \gls{alpha} by maximum likelihood using partner count
data from several sequential time intervals, but they admit such detailed data
are not usually available. Moreover, their dataset was constructed via a random
survey, which would likely miss the few high-degree nodes characterizing a
power law degree distribution. In summary, each of the \gls{BA} parameters may
be estimated without phylodynamics, but there are sufficient difficulties that
we believe an alternative method using sequence data is warranted.}

\subsection{Using synthetic data to investigate identifiability and sources of
estimation error and bias}

\add{We have argued that the parameters of the \gls{BA} model are interesting
and worth estimating with phylodynamic methods. However, these estimates will
only constitute ``useful'' information about the network if the parameters are
identifiable from phylogenetic data. Roughly speaking, the identifiability of a
parameter says how much information about that parameter can possibly be
obtained from the observed data. If the parameters of the \gls{BA} model do not
influence tree shape at all, then we cannot possibly estimate them -- the
posterior distribution will exactly resemble the prior, no matter how accurate
a representation of the posterior we are able to produce. Hence, before
proceeding with a full validation of \software{netabc} on simulated data, we
undertook two experiments designed to assess the identifiability of the
\gls{BA} parameters. These experiments only investigated one parameter of the
\gls{BA} model at a time while holding all others fixed, a strategy commonly
used when performing sensitivity analyses of mathematical models. This allowed
us to perform a fast preliminary analysis without dealing with the ``curse of
dimensionality'' of the full parameter space. The experiments are motivated and
described on a high level here, with more detail provided in the next section.}

\add{First, we simulated trees under three different values of each parameter,
and asked how well we could tell the different trees apart. The better we are
able to distinguish the trees, the more identifiability we might expect for the
corresponding parameter when we attempt to estimate it with \gls{ABC}. This
experiment also had the secondary purpose of validating our choice of the tree
kernel as a distance measure in \gls{ABC}. To tell the trees apart, we used a
classifier based on the tree kernel, but we also tested two other tree shape
statistics: one which considers only the topology, and another which considers
only branching times. Since the tree kernel incorporates both of these sources
of information, we expected it to outperform the other two statistics. Finally,
the tree kernel can be ``tuned'' by adjusting the values of the meta-parameters
\gls{lambda} and \gls{sigma}. The results of this experiment were used to
select values for these meta-parameters to carry forward to the rest of the
thesis, based on their accuracy in distinguishing the different trees. }

\add{A second experiment was designed to test whether we could actually
estimate the parameters numerically, rather than just telling three different
values apart, and also to assess how identifiability varied in different
regions of the parameter space. An individual tree was compared to simulated
trees on a one-dimensional grid of values of one \gls{BA} parameter, to obtain
a ``distribution'' of kernel score values. From this distribution, estimates
and credible intervals of the parameter could be calculated. Repeating this
experiment with trees located throughout the parameter space allowed us to
beter quantify the identifiability. Furthermore, doing marginal estimation
(that is, estimating one parameter with all others fixed) can provide insight
into any biases observed when doing joint estimation with \gls{ABC}. If grid
search is inaccurate, it indicates a lack of parameter identifiability.
However, if the marginal grid search estimates are accurate but the estimates
obtained with \gls{ABC} are biased, this points to confounding between the
parameters which could only be observed when they are all estimated jointly.}

\add{After these preliminary experiments, the strategy we used for testing
\software{netabc} was a standard simulation-based validation. Transmission
trees were simulated under several combinations of parameter values, and we
tried to recover these values with \software{netabc}. We then used a
multivariable analysis to investigate how accuracy of these estimates was
influenced by the true parameter values.}

\add{In the previous section, we argued that the \gls{BA} model parameters were
worth investigating, and here we have presented several computational
experiments designed to assess their identifiability. There is a final, more
technical aspect of our method's ``usefulness'' to consider, which is the
accuracy of the \gls{ABC} approximation to the posterior. As discussed in
\cref{sec:abc}, \gls{ABC} does not target the posterior distribution directly,
but rather an approximate posterior derived from simulated data and a distance
function. \gls{ABC} \emph{assumes} that this approximate posterior resembles
the true posterior, and it is critical for our estimates' relevance that this
assumption holds. There are two potential causes of an inaccurate \gls{ABC}
approximation~\autocite{fearnhead2012constructing}. First, the Monte Carlo
approximation to the \gls{ABC} target distribution may be poor, due to the
settings used for \gls{ABC}-\gls{SMC}. Second, the \gls{ABC} target
distribution may not resemble the posterior, due to a poor choice of distance
function. We designed two experiments to investigate the impact of these
sources of error.}

\add{The Monte Carlo approximation error is fairly easily quantified by simply
increasing the computing power used for \gls{SMC}. We ran one simulation using
a larger number of pakticles, more simulated datasets per particle, and a
higher value for \gls{alphaess}. A substantial improvement in accuracy
resulting from these changes would likely indicate a high Monte Carlo error
with the lower settings. The second issue, the resemblance of the \gls{ABC}
target distribution to the true posterior, is somewhat more difficult to
investigate. We do not have access to the true posterior, even for simulations
where the true parameter values are known. To address this source of error, we
performed marginal parameter estimation with \gls{ABC} by informing
\software{netabc} of some of the true parameter values. Any inaccuracy or bias
observed only in the joint estimation results, but not the marginal estimates, 
is most easily explained by interdependence between parameters in the true
posterior. However, errors observed in both marginal and joint estimates could
be due to either the shape of the true posterior or an inaccurate \gls{ABC}
approximation, and we have no way to distinguish one from the other. In other
words, this experiment provided only an upper bound on the error due to an
inaccurate \gls{ABC} approximation.}

\subsection{Methods}

\del{We investigated four parameters related to the \gls{BA} contact network
model, denoted \gls{N}, \gls{m}, \gls{alpha}, \gls{I} (see \cref{subsec:pa}).
The first three of these are parameters of the model itself, while \gls{I} is
related to the simulation of transmission trees over the network. However, we
will refer to all four as \gls{BA} parameters. \gls{N} denotes the total number
of nodes in the network, or equivalently, susceptible individuals in the
population. When a node is added to the network, \gls{m} new undirected edges
are added incident to it, and are attached to existing nodes of degree $d$ with
probability proportional to $d^{\gls{alpha}} + 1$ (\cref{subsec:pa}). To
simulate transmission trees over a \gls{BA} network, we allowed an epidemic to
spread until \gls{I} nodes were infected, and sampled a transmission tree at
that time.}

\add{For all simulations,} we assumed that all contacts had symmetric
transmission risk, which was implemented by replacing each undirected edge in
the network with two directed edges (one in each direction). Nodes in our
networks followed simple \gls{SI} dynamics, meaning that they became infected
at a rate proportional to their number of infected neighbours, and never
recovered. We did not consider the time scale of the transmission trees in
these simulations, only their shape. Therefore, the transmission rate along
each edge in the network was set to 1, the removal rate of each node was set to
0, and all transmission trees' branch lengths were scaled by their mean. The
\textit{igraph} library's implementation of the BA
model~\autocite{csardi2006igraph} was used to generate the graphs. The analyses
were run on Westgrid (\url{https://www.westgrid.ca/}) and a local computer
cluster. With the exception of our own C programs, all analyses were done in
\software{R}, and all packages listed below are \software{R} packages.
\add{Code to run all experiments is freely available at
    \url{https://github.com/rmcclosk/thesis}.}

\subsubsection*{Classifiers for BA model parameters based on tree shape}
\label{subsec:kernel}

\glsreset{nltt}

\add{Our first computational experiment was designed as an exploratory analysis
of the four \gls{BA} model parameters defined above: \gls{alpha}, \gls{I},
\gls{m}, and \gls{N}. The objective of this experiment was to determine whether
any of the four parameters were identifiable from the shape of the transmission
tree, as quantified by the tree kernel. Each of the \gls{BA} model parameters
was varied one at a time, while holding the other parameters at fixed, known
values. Contact networks were generated according to each set of parameter
values, and transmission trees were simulated over the networks. We then
evaluated how well a classifier based on the tree kernel could differentiate
the trees simulated under distinct parameter values. If the classifier's
cross-validation accuracy was high, this could be taken as an indication that
the parameter in question was identifiable in the range of values considered. 
A caveat of this preliminary analysis is that, since all parameters but one
were held at known values, nothing could be said about the identifiability of
\emph{combinations} of parameters; this issue will be explored later by jointly
estimating all parameters with \gls{ABC}}.

\add{In addition to testing for identifiability, a secondary objective of
this analysis was to validate the use of the tree kernel as a distance measure
for \gls{ABC} in our context. As discussed in \cref{sec:abc}, the choice of
distance function is extremely important for the accuracy of the \gls{ABC}
approximation to the posterior. Therefore, we evaluated two additional tree
statistics in the same manner as we evaluated the tree kernel (that is, by
constructing and testing a classifier). First, we considered Sackin's
index~\autocite{shao1990tree}, which measures the degree of imbalance or
asymmetry in a phylogeny (see \cref{subsec:treeshape}). Sackin's index is
widely used for characterizing phylogenies~\autocite{frost2013modelling} and
has been demonstrated to vary between transmission trees simulated under
different contact network types~\autocite{leventhal2012inferring}. Sackin's
index does not take branch lengths into account, considering only the tree's
topology. The other statistic we considered was the
\gls{nltt}~\autocite{janzen2015approximate}, which compares two trees based on
normalized distributions of their branching times (see
\cref{subsec:treeshape}). In contrast with Sackin's index, the \gls{nltt} does
not explicity consider the trees' topologies, but it does use their normalized
branch lengths. While the \gls{nltt} is a newly developed statistic not yet in
widespread use, the unnormalized \gls{ltt}~\autocite{nee1992tempo} was the
basis of seminal early work extracting epidemiological information from
phylogenies~\autocite{holmes1995revealing}. We expected the tree kernel to
classify the \gls{BA} parameters more accurately than either Sackin's index or
the \gls{nltt}, since the tree kernel takes both topology and branch lengths
into account.}

This experiment involved a large number of variables that were varied
combinatorially. For ease of exposition, we will describe a single experiment
first, then enumerate the values of all variables for which the experiment was
repeated. The parameters of the tree kernel, $\lambda$ and $\sigma$
(\cref{subsec:treeshape}), will be referred to as \defn{meta-parameters} to
distinguish them from the parameters of the \gls{BA} model.

The attachment power parameter \gls{alpha} was varied among three values: 0.5,
1.0, and 1.5. For each value, the \software{sample\_pa} function in the
\software{igraph} package was used to simulate 100 networks, with the other
parameters set to \gls{N} = 5000 and \gls{m} = 2. This step yielded a total of
300 networks. An epidemic was simulated on each network using our
\software{nettree} binary until \gls{I} = 1000 nodes were infected, at which
point 500 of them were sampled to form a transmission tree. A total of 300
transmission trees were thus obtained, comprised of 100 trees for each of the
three values of \gls{alpha}. The trees were ``ladderized'' so that the subtree
descending from the left child of each node was not smaller than that
descending from the right child. Summary statistics, such as Sackin's index and
the ratio of internal to terminal branch lengths, were computed for each
simulated tree using our \software{treestat} binary. The trees were visualized
using the \software{ape} package~\autocite{paradis2004ape}. {\color{blue}\uline{
Both the tree kernel and the \gls{nltt} are pairwise statistics, and the
\glspl{SVM} classifiers we used to investigate them operate on pairwise
distance matrices.}} Our \software{treekernel} binary was used to calculate the
value of the kernel for each pair of trees, with the meta-parameters set to
$\lambda = 0.3$ and $\sigma = 4$. These values were stored in a symmetric 300
$\times$ 300 kernel matrix.  Similarly, we computed the \gls{nltt} statistic
between each pair of trees using our \software{treestat} binary, and stored
them in a second $300 \times 300$ matrix.

To investigate the identifiability of \gls{alpha} from tree shape, we
constructed classifiers for \gls{alpha} based on the three tree shape
statistics discussed above. First, we used the \software{kernlab}
package~\autocite{zeileis2004kernlab} to create a \gls{kSVR} classifier using
the computed kernel matrix. Second, we used the \software{e1071}
package~\autocite{meyer2015e1071} to create an ordinary \gls{SVR} classifier
using the pairwise \gls{nltt} matrix. Finally, we performed an ordinary linear
regression of \gls{alpha} against Sackin's index. Each of these classifiers was
evaluated with 1000 two-fold cross-validations. We also performed a \gls{kPCA}
projection of the kernel matrix, and used it to visualize the separation of the
different \gls{alpha} values in the tree kernel's feature space. A schematic of
this experiment is presented in \cref{fig:kernelexpt}.

\begin{figure}[ht]
    \centering
    \includegraphics{kernel-expt.pdf}
    \caption[
        Schematic of classifier experiments investigating identifiability of BA
        model parameters from tree shapes.
    ]{
        Schematic of classifier experiments investigating identifiability of BA
        model parameters from tree shapes. The parameters of the BA model were
        varied one at a time \add{while holding all others fixed}. Transmission
        trees were simulated under three different values of each
        parameter\del{, then compared pairwise using the tree kernel.}
        Classifiers were constructed for each parameter \add{based on three
        tree shape statistics}, and their accuracy was evaluated by
        cross-validation. Kernel-\gls{PCA} projections were used to visually
        examine the separation of the trees in the feature space defined by the
        tree kernel.
    }
    \label{fig:kernelexpt}
\end{figure}

Similar experiments were performed with the values shown in
\cref{tab:kernelexpt}. The other three \gls{BA} parameters, \gls{N}, \gls{m},
and \gls{I}, were each varied while holding the others fixed. The experiments
for \gls{alpha}, \gls{m}, and \gls{N} were repeated with three different values
of \gls{I}. All experiments were repeated with trees having three different
numbers of tips. Kernel matrices were computed for all pairs of the
meta-parameters \gls{lambda} = \sett{0.2, 0.3, 0.4} and \gls{sigma} =
\sett{\nicefrac18, \nicefrac14, \nicefrac12, 1, 2, 4, 8}.

\begin{landscape}
\begin{table}[ht]
    \centering
    \input{\tablepath/kernel-expt}
    \caption[
        Values of parameters and meta-parameters used in classifier experiments
        to investigate identifiability of BA model parameters from tree shapes.
    ]{
        Values of parameters and meta-parameters used in classifier experiments
        to investigate identifiability of BA model parameters from tree shapes.
        Each row corresponds to one of the BA model parameters. One kernel
        matrix was created for every combination of values except the one
        indicated in the ``varied parameter'' column, which was varied when
        producing simulated trees.
    }
    \label{tab:kernelexpt}
\end{table}

\begin{table}[ht]
    \centering
    \input{\tablepath/gridsearch-expt}
    \caption[
        Values of parameters and meta-parameters used in grid search
        experiments to further investigate identifiability of BA model
        parameters.
    ]{
        Values of parameters and meta-parameters used in grid search
        experiments to further investigate identifiability of BA model
        parameters. Trees were simulated under the test values, and compared to
        a grid of trees simulated under the grid values. Kernel scores were
        used to calculate point estimates and credible intervals for each
        parameter, which were compared to the test values.
    }
  \label{tab:gridexpt}
\end{table}
\end{landscape}

\subsubsection*{Grid search}

\add{The previous experiment was an exploratory analysis intended to determine
which of the \gls{BA} parameters were identifiable, and whether the tree kernel
could potentially be used to distinguish different parameter values when all
others were held fixed. In this experiment, which was still of an exploratory
nature, we continued to consider one parameter at a time while fixing the other
three. However, rather than checking for identifiability, we were now
interested in quantifying the accuracy and precision of kernel score-based
estimates. This was done by examining the distribution of kernel scores on a
grid of parameter values, when trees simulated according to those values were
compared with a single simulated test tree.}

As in the previous section, we will begin by describing a single experiment,
and then list the variables for which similar experiments were performed. We
varied \gls{alpha} along a narrowly spaced grid of values: 0, 0.01, \ldots, 2.
For each value, fifteen networks were generated with \software{igraph}, and
transmission trees were simulated over each using \software{nettree}. These
trees will be referred to as ``grid trees''. Next, one further test tree was
simulated with the test value \gls{alpha} = 0. Both the grid trees and the test
tree had 500 tips, and were simulated with the other \gls{BA} parameters set to
\add{the known values} $\gls{N} = 5000$, $\gls{m} = 2$, and $\gls{I} = 1000$.
The test tree was compared to each of the grid trees using the tree kernel,
with the meta-parameters set to $\gls{lambda} = 0.3$ and $\gls{sigma} = 4$,
using the \software{treekernel} binary. The median kernel score was calculated
for each grid value, and the scores were normalized such that the area under
the curve was equal to 1. \del{The grid value with the highest median kernel
score was taken as the point estimate for the test value.} \add{For all
parameters except \gls{m}}, 50\% and 95\% highest denstity intervals wer
obtained using the \software{hpd} function in the \software{TeachingDemos}
package~\autocite{snow2013teachingdemos}. \add{Since the \software{hpd}
function assumes a continuous distribution, we implemented our own version for
discrete distributions to use for \gls{m}.}

Each experiment of the type just described was repeated ten times with the same
test value. Similar experiments were performed for each of the four \gls{BA}
parameters, with several test values and trees of varying sizes. The variables
are listed in \cref{tab:gridexpt}. A graphical schematic of the grid search
experiments is shown in \cref{fig:gridexpt}.

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{gridsearch-expt}
    \caption[
        Schematic of grid search experiment to further investigate
        identifiability of BA model parameters from tree shapes.
    ]{
        Schematic of grid search experiment to further investigate
        identifiability of BA model parameters from tree shapes. Trees were
        simulated along a narrowly spaced grid of values of one parameter
        (``grid trees'') \add{with all other parameters fixed to known values}.
        Separate trees were simulated for a small subset of the grid values
        (``test trees''), \add{also holding the other parameters fixed}. Each
        test tree was compared to every grid tree using the tree kernel, and
        the resulting kernel scores were normalized to resemble a probability
        density from which the mode and 95\% highest density interval were
        calculated.
    }
    \label{fig:gridexpt}
\end{figure}

\subsubsection*{Approximate Bayesian computation}

\add{Our final synthetic data experiment was designed to test the full
\gls{ABC}-\gls{SMC} algorithm by jointly estimating the four parameters of the
\gls{BA} model. We used the standard validation approach of simulating
transmission trees under the model with known parameter values and attempting
to recover those values with \software{netabc}. The algorithm was not informed
of any of the true parameter values for the main set of simulations. Despite
the fact that the parameter values used to generate the simulated transmission
trees were known, the true posterior distributions on the \gls{BA} parameters
were unknown. Therefore, any apparent errors or biases in the estimates could
be due to either poor performance of our method, or to real features of the
posterior distribution. The latter type of error reflects on the suitability of
the model, but does not invalidate the use of our method in cases where the
parameters are more identifiable. Two retrospective experiments were peformed
to disambiguate some of the observed errors: one where we ran a simulation with
increased computational power to test for an increase in accuracy, and a second
where we estimated parameters marginally to remove confounding from the other
parameters in the joint posterior.}

\del{We simulated three trees each under a variety of parameter values and ran
the \software{netabc} program to estimate posterior distributions for the
parameters.} The parameter values and priors used are listed in
\cref{tab:abcexpt}. The tree kernel meta-parameters were set to $\gls{lambda} =
0.3$ and $\gls{sigma} = 4$. The \gls{SMC} algorithm was run with 1000
particles, five sampled datasets per particle, and \gls{alphaess} \del{(not to
be confused with the \gls{BA} preferential attachment parameter, see
\cref{subsec:adaptsmc})} set to 0.95. The algorithm was stopped when the
acceptance rate of the \gls{MH} kernel dropped below 1.5\%, the same criterion
used by \textcite{del2012adaptive}. \add{For visualization,} approximate
marginal posterior densities for each parameter were calculated using the
\software{density} function in \software{R} applied to the final weighted
population of particles. \del{Credible intervals were obtained for each
parameter using the \software{HPDinterval} function in the \software{coda}
package~\autocite{plummer2006coda}.} \add{Posterior means obtained for each
parameter using the \software{wtd.mean} function in the \software{Hmisc}
package~\autocite{harrell2016hmisc}. Credible intervals were obtained using the
\software{hpd} function in the \software{TeachingDemos}
package~\autocite{snow2013teachingdemos} for \gls{alpha} \gls{I}, and \gls{N}, 
and using our own implementation for discrete distributions for \gls{m}.}

\begin{table}[ht]
    \centering
    \input{\tablepath/abc-expt}
    \caption[
        Parameter values used in simulation experiments to test accuracy of
        \gls{BA} model fitting with \software{netabc}.
    ]{
        Parameter values used in simulation experiments to test accuracy of
        \gls{BA} model fitting with \software{netabc}. Trees were simulated
        under the test values, and \software{netabc} was used to estimate
        posterior distributions on the \gls{BA} parameters for each simulated
        tree. \software{Netabc} was na\"ive to the true parameter values.
    }
    \label{tab:abcexpt}
\end{table}

\add{To evaluate the effects of the true parameter values on the accuracy of
the posterior mean estimates, we analysed the \gls{alpha} and \gls{I}
parameters individually using \glspl{GLM}. The response variable was the error
of the point estimate, and the predictor variables were the true values of
\gls{alpha}, \gls{I}, and \gls{m}. We did not test for differences across true
values of \gls{N}, because \gls{N} was not varied in these simulations. The
distribution family and link function for the \glspl{GLM} were chosen as
Gaussian and inverse, respectively, by examination of residual plots and
\gls{AIC}. The $p$-values of the estimated \glspl{GLM} coefficients were
corrected using Holm-Bonferroni correction~\autocite{holm1979simple} with $n =
6$ (two \glspl{GLM} with three predictors each). Because there was clearly
little to no identifiability of \gls{N} and \gls{m} with \gls{ABC} (see results
in next section), we did not construct \glspl{GLM} for those parameters.}

Two additional simulations were performed to address \del{potential sources of error}
\add{the possible impact of two types of model misspecification}. To consider
the effect of heterogeneity among nodes, we generated a network where half the
nodes were attached with power $\gls{alpha} = 0.5$ and the other half with power
$\gls{alpha} = 1.5$. The other parameters for this network were $\gls{N} = 5000$,
$\gls{I} = 1000$, and $\gls{m} = 2$. To investigate the effects of potential
sampling bias~\autocite{karcher2016quantifying}, we simulated a transmission
tree where the tips were sampled in a peer-driven fashion, rather than at
random. That is, the probability to sample a node was twice as high if any of
that node's network peers had already been sampled. The parameters of this
network were $\gls{N} = 5000$, $\gls{I} = 2000$, $\gls{m} = 2$, and
$\gls{alpha} = 0.5$.

\end{document}

\subsection{Results}

\subsubsection*{Classifiers for BA model parameters based on tree shape}

Trees simulated under different values of \gls{alpha} were visibly quite
distinct (\cref{fig:alphatrees}). In particular, higher values of \gls{alpha}
produce networks with a small number of highly connected nodes, which, once
infected, are likely to transmit to many other nodes. This results in a more
unbalanced, ladder-like structure in the phylogeny, compared to networks with
lower \gls{alpha} values. None of the other three parameters produced trees
that were as easily distinguished from each other
(\cref{fig:Itrees,fig:mtrees,fig:Ntrees,fig:Itrees}).  Sackin's index, which
measures tree imbalance, was significantly correlated with all four parameters
    (for $\alpha$, $I$, $m$, and $N$ respectively: Spearman's rho =
     0.85,
     \ensuremath{-0.12},
     \ensuremath{-0.13},
     0.09;
     $p$-values
     ${<}10^{-5}$,
     $0.003$,
     ${<}10^{-5}$,
     ${<}10^{-5}$).
The ratio of internal to terminal branch lengths was negatively correlated with
\gls{alpha} and \gls{I}, and positively correlated with \gls{m} and \gls{N}
  (Spearman's rho
    \ensuremath{-0.84},
    \ensuremath{-0.69},
    0.1,
    0.18;
  all $p < 10^{-5}$).

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{kernel-alpha-tree.pdf}
  \caption[Simulated transmission trees under three different values of BA parameter $\alpha$]{
    Simulated transmission trees under three different values of BA parameter
    $\alpha$. Epidemics were simulated on \gls{BA} networks of 5000 nodes, with
    \gls{alpha} equal to 0.5, 1.0, or 1.5, until 1000 individuals were
    infected. Transmission trees were created by sampling 500 infected nodes.
    Higher \gls{alpha} values produced networks with a small number of
    highly-connected nodes, resulting in highly unbalanced, ladder-like trees.
  }
  \label{fig:alphatrees}
\end{figure}

\Cref{fig:kpca} shows \gls{kPCA} projections of the simulated trees onto the
first two principal components of the kernel matrix. The figure shows only the
simulations with 500-tip trees and 1000 infected nodes. The three \gls{alpha}
and \gls{I} values considered are well separated from each other in feature
space. On the other hand, the three \gls{N} values overlap significantly, and
the three \gls{m} values are virtually indistinguishable. Similar observations
can be made for other values of \gls{I} and the number of tips
(\cref{fig:alphakpca,fig:Nkpca,fig:Ikpca,fig:mkpca}). The values of \gls{I} and
\gls{N} separated more clearly with larger numbers of tips, and in the case of
\gls{N}, with larger epidemic sizes {\color{blue}(\cref{fig:Ikpca,fig:Nkpca})}.

\begin{figure}[ht]
  \centering
  \includegraphics{kernel-kpca.pdf}
  \caption[Kernel-PCA projections of simulated trees under varying BA
           parameter values.]{
    Each parameter of the \gls{BA} model was individually varied to produce 300
    simulated trees. Kernel matrices were formed from all pairwise kernel
    scores among each set of 300 trees. The trees were projected onto the first
    two principal components of the kernel matrix calculated using \gls{kPCA}.
    All trees had 500 tips. The parameters not being varied were set to
    \gls{alpha} = 1, \gls{I} = 1000, \gls{m} = 2, and \gls{N} = 5000. The tree
    kernel meta-parameters were $\lambda = 0.3$ and $\sigma = 4$.
  }
  \label{fig:kpca}
\end{figure}

Accuracy of the \gls{kSVR} classifiers varied based on the parameter being
tested (\cref{fig:rsquared}, left). Classifiers based on two other tree
statistics, the \gls{nltt} and Sackin's index, generally exhibited worse
performance than the tree kernel, although the magnitude of the disparity
varied between the parameters (\cref{fig:rsquared}, centre and right). The
results were largely robust to variations in the tree kernel meta-parameters
$\lambda$ and $\sigma$, although accuracy varied between different epidemic and
sampling scenarios
(\cref{fig:alphacrossv,fig:mcrossv,fig:Icrossv,fig:Ncrossv}).

When classifying $\alpha$, the \gls{kSVR} classifier had an average $R^2$ of 
    0.92,
compared to 
    0.56
for the \gls{nltt}-based SVR, and
    0.75
for the linear regression against Sackin's index. There was little variation
about the mean for different tree and epidemic sizes. No classifier could
accurately identify the $m$ parameter in any epidemic scenario, with average
$R^2$ values of 
  0.12 for \gls{kSVR},
  0.01 for the \gls{nltt}, and
  0.06
for Sackin's index. Again, there was little variation in accuracy between
epidemic scenarios, although the accuracy of the \gls{kSVR} was slightly higher
on 1000-tip trees 
    (average $R^2$ 
     0.01,
     0.11,
     0.32
     for 100, 500, and 1000 tips respectively).

The accuracy of classifiers for $I$ varied significantly with the number of
tips in the tree. For 100-tip trees, the average $R^2$ values were
  0.7,
  0.55, and
  0.02
for the tree kernel, \gls{nltt}, and Sackin's index respectively. For 500-tip
trees, the values increased to
  0.93,
  0.83, and
  0.07.
Finally, the performance of classifiers for $N$ depended heavily on the
epidemic scenario. The $R^2$ of the \gls{kSVR} classifier ranged from
  0.08
for the smallest epidemic and smallest sample size, to
  0.82
for the largest. Likewise, $R^2$ for the \gls{nltt}-based SVR ranged from 
  0.01
to
  0.54.
Sackin's index did not accurately classify $N$ in any scenario, with an average
$R^2$ of
  0.03
and little variation between scenarios.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{kernel-rsquared.pdf}
  \caption[Cross-validation accuracy of kernel-SVR, nLTT-based SVR, and
  Sackin's index regression classifiers for BA model parameters.]{
      Cross-validation accuracy of kernel-SVR classifier (left), SVR classifier
      using \gls{nltt} (centre), and linear regression using Sackin's index
      (right) for \gls{BA} model parameters. Kernel meta-parameters were set to
      $\lambda = 0.3$ and $\sigma = 4$. Each point was calculated based on 300
      simulated transmission trees over networks with three different values of
      the parameter being tested, assuming perfect knowledge of the other
      parameters. Vertical lines are empirical 95\% confidence intervals based
      on 1000 two-fold cross-validations. {\color{blue}\uline{ The classifiers
      for \gls{I} were not evaluated with 1000-tip trees, because one of the
      tested \gls{I} values was 500, and it is not possible to sample a tree of
      size 1000 from 500 infected individuals.}}
  }
  \label{fig:rsquared}
\end{figure}

\subsubsection*{Marginal parameter estimates with grid search}

{\color{red}\sout{The accuracy of grid search estimates largely paralleled that
of the \gls{kSVR} classifiers.}}\Cref{fig:gridest} shows point estimates and
{\color{blue}\uline{50\%}} and 95\% highest density intervals for each of the
\gls{BA} parameters, for one replicate experiment with 500-tip trees. Plots
showing the point estimates for all replicates can be found in
\cref{fig:gridptalpha,fig:gridptI,fig:gridptm,fig:gridptN}. For all parameters
except $m$, the error of point estimates was negatively correlated with the
number of sampled tips in the tree (for \gls{alpha}, \gls{I}, and \gls{N}
respectively: Spearman's $\rho$ = \ensuremath{-0.22},
\ensuremath{-0.51}, \ensuremath{-0.16}; $p$-values
    $4\!\times\!10^{-4}$,
    ${<}10^{-5}$,
    $0.01$).
The 95\% highest density intervals obtained for all parameters were extremely
wide, occupying $>$75\% of the grid in all cases (\cref{fig:gridest}).

{\color{blue}\uline{Across all replicates, $R^2$ values for the correlations
between the estimated and true values were
    $0.91$,
    $0.91$,
    $0.25$, and
    $0.54$
for \gls{alpha}, \gls{I}, \gls{m}, and \gls{N} respectively. The mean absolute
errors of the point estimates were
    $0.14$,
    $310$,
    $1.31$, and
    $2419$,
representing
    $7$\%,
    $8$\%,
    $26$\%, and
    $17$\%
of the respective grids.}

\uline{Qualitatively, each of the parameters exhibited weak identifiability
within particular sections of the grid} (\cref{fig:gridest}). \uline{The 50\%
\glspl{HDI} for \gls{alpha} were nearly identical for the values $\gls{alpha}
\leq 0.5$ (on average [X-X]) and for $\gls{alpha} \geq 1.5$ ([X-X]).
For \gls{I}, similar 50\% \gls{HPD} were observed for $\gls{I} \leq 1500$
(average [X-X]) and for $\gls{I} > 3000$ ([X-X]). For \gls{m}, both the 50\%
and 95\% \glspl{HDI} were generally identical for $\gls{m} > 2$ (average [X-X]
for 50\% and [X-X] for 95\% \gls{HDI}). We did not observe a similar pattern
for \gls{N}, however the boundaries of the \gls{HDI} did not always
monotonically increase with the actual value.}

\Cref{fig:gridalpha,fig:gridI,fig:gridm,fig:gridN} \uline{show kernel score
distributions of kernel scores along the grid for each parameter. The
distributions for some values, such as $\gls{alpha} = 1.25$, $\gls{I} = 500$
and $4500$, $\gls{m} = 1$, and $\gls{N} = 1000$, exhibited distinct peaks
around the true value. This indicates that these values produce distinctively
shaped trees that can be identified with the tree kernel, when the other
parameter values are fixed and known. However, for the majority of values of
each parameter, the score distributions were fairly flat around the true value. 
This means there is a range of values which produce similarly shaped trees, and
the parameter is less identifiable within that range. The exception was
\gls{I}, whose score distributions exhibited a more or less rounded shape with
the highest point near the true value. }}

{\color{red}\sout{The \gls{alpha} parameter was the most accurately estimated,
with point estimates having an average deviation of 
    0.14
from the true value, on a grid from 0 to 2. The error of point estimates varied
significantly between true values of \gls{alpha}
    (one-way \gls{ANOVA}, $p$ ${<}10^{-5}$). In
particular, errors were lower for the values \gls{alpha} = 1.0 and 1.25 than
for the other values
    (average errors 
    0.03
    for \gls{alpha} = 1.0 or 1.5 vs.
    0.17
    for \gls{alpha} $\neq$ 1.0 or 1.5),
and this difference was significant
    (Wilcoxon rank-sum test, $p {<}10^{-5}$,
}\cref{fig:gridptalpha}).
\sout{These two values exhibited different qualitative behaviour than the other
values in terms of the distribution of kernel scores along the grid
}(\cref{fig:gridalpha}). \sout{In particular, there was a pronounced peak in scores
around the true value, in contrast to the other values where the scores were
flat around the true value. The effect was most obvious for the value
\gls{alpha} = 1.25.}}

%The average absolute error of the point estimates for \gls{I} was 
%    d["I", round(mean(error))] individuals,
%on a grid of 500 to 5000, and these errors differed between true values of
%\gls{I}
%    (one-way \gls{ANOVA}, $p pp(av["I", p], eq=TRUE)$).
%The errors for $2000 \leq I \leq 3000$ were higher than those for the other
%values
%    (average errors
%     d["I", .SD[true_value >= 2000 & true_value <= 3000, round(mean(error))]]
%     for $2000 \leq I \leq 3000$ vs.
%     d["I", .SD[true_value < 2000 | true_value > 3000, round(mean(error))]]
%     for $I < 2000$ or $I > 3000$),
%and this difference was significant
%    (Wilcoxon rank-sum test, $p pp(I.test$p.value, eq=TRUE)$,
%     \cref{fig:gridptI}).
%Kernel score distributions for all test values exhibited a similar rounded
%shape (\cref{fig:gridI}). 
%
%The average error for \gls{m} was
%    d["m", round(mean(error), 2)] edges per vertex,
%on a grid from 1 to 6. The error varied significantly between the true values
%of \gls{m} 
%    (one-way \gls{ANOVA}, $p pp(av["m", p], eq=TRUE)$).
%Errors for the value \gls{m} = 1 were lower than the other values
%    (average errors
%    d["m", .SD[true_value == 1, round(mean(error), 2)]]
%    for $m = 1$ vs.
%    d["m", .SD[true_value > 1, round(mean(error), 2)]]
%    for $m > 1$),
%and this difference was significant
%    (Wilcoxon rank-sum test, $p pp(m.test$p.value, eq=TRUE)$,
%     \cref{fig:gridptm}).
%The value $m = 1$ causes the network to take on a distinct shape relative to
%higher \gls{m} values, namely a tree (\ie there are no cycles, see
%\cref{subsec:treeshape}). The kernel score distribution had a peak at $m = 1$
%when this was the true value, and a valley at $m = 1$ when the true value of
%$m$ was greater that 1 (\cref{fig:gridm}).
%
%The average error for \gls{N} was 
%    d[parameter == "N", round(mean(error))] individuals,
%on a grid from 1000 to 15000, and was varied significantly with the true value
%of \gls{N}
%    (one-way \gls{ANOVA}, $p pp(av["N", p], eq=TRUE)$).
%The errors were lower for $N \leq 3000$
%    (average errors
%     d["N", .SD[true_value <= 3000, round(mean(error))]]
%     for $N \leq 3000$ vs.
%     d["N", .SD[true_value > 3000, round(mean(error))]]
%     for $N > 3000$),
%and this difference was significant
%    (Wilcoxon rank-sum test, $p pp(N.test$p.value, eq=TRUE)$,
%     \cref{fig:gridptN}).
%The kernel score distribution had a peak at $N = 1000$ when this was the true
%value, and a valley there otherwise (\cref{fig:gridN}). Except for this valley,
%the distributions were flat for $N > 3000$.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{gridsearch-example}
  \caption[
    Grid search estimates of \gls{BA} model parameters for one replicate
    experiment with trees of size 500.
  ]{
    Grid search estimates of \gls{BA} model parameters for one replicate
    experiment with trees of size 500. Point estimates (dots), 95\% highest
    density intervals (lines), and 50\% highest density intervals (notches) for
    each \gls{BA} model parameter, obtained using grid search. Networks and
    transmission trees were simulated over a grid of values for each parameter
    while holding the others fixed. For a subset of the grid values ($x$-axis),
    test networks and trees were created and compared to each tree on the grid
    using the tree kernel. The kernel scores along the grid were normalized to
    resemble a probability distribution, from which the mode and highest
    density interval were calculated. 
  } 
  \label{fig:gridest}
\end{figure}

\subsubsection*{Joint parameter estimates with \software{netabc}}

\Cref{fig:abcptm2} shows {\color{red}\sout{\gls{MAP}}}
{\color{blue}\uline{posterior mean}} point estimates of the BA model parameters
obtained with ABC on simulated data. The estimates shown
correspond only to the simulations where the $m$ parameter was set to 2,
however the results for $m = 3$ and $m = 4$ were similar
(\cref{fig:abcptm3,fig:abcptm4}). Average boundaries of 95\% \gls{HPD}
intervals are given in \cref{tab:abchpd}.

{\color{blue}
\uline{Across all simulations, the median [\gls{IQR}] absolute errors of the parameter 
estimates obtained with \software{netabc} were
    0.11 
    [0.03 - 
    0.25]
for \gls{alpha},
    492 
    [294 - 
    782]
for \gls{I},
    1 
    [0 - 
    1]
for \gls{m}, and
    4153 
    [3660 - 
    4489]
for \gls{N}. These errors comprised, respectively,
    6\%,
    11\%,
    17\%, and
    29\%
of the regions of nonzero prior density. Average \gls{HPD} interval widths were
    0.68,
    2454,
    3.01, and
    12046,
representing 
    34\%,
    55\%,
    50\%, and
    83\%
of the nonzero prior density regions. Point estimates of \gls{I} were upwardly
biased: \gls{I} was overestimated in
    69
out of 
    72
simulations
    (96\%).
The estimates for \gls{m} and \gls{N} were similar across all simulations
(median [IQR] point estimates
    3
    [3 - 
    3]
and 
    9153
    [8660 - 
    9489])
regardless of the true values of any of the \gls{BA} parameters.}

\uline{To analyse the effects of the true parameter values on the
accuracy our estimates of \gls{alpha} and \gls{I}, we fitted one \glspl{GLM}
for each of these two parameters, with error rate as the dependent variable and
the true parameter values as independent variables. Since the estimates of
\gls{m} and \gls{N} were roughly equal across all simulations}
(\cref{fig:abcptm2,fig:abcptm3,fig:abcptm4}), \uline{\glspl{GLM} were not
fitted for these parameters. The estimated coefficients are shown in
}\cref{tab:glmalpha,tab:glmI}. \uline{For the parameters \gls{alpha}, \gls{I},
and \gls{N}, the \gls{GLM} were fitted using the inverse link function. That
is, if $p$ is the true value of the parameter and $\hat{p}$ is a random
variable representing our estimate of the parameter, the \gls{GLM} posits a
relationship of the form }
\[
    \E(|p-\hat{p}|) = (\beta_0 + \beta_\alpha \alpha + \beta_I I + \beta_m m)^{-1},
\]
\uline{where the $\beta$'s are coefficients to be fitted. If the true value
\gls{alpha}, say, is increased by one, the \emph{inverse} of the expected
absolute error will increase by $\beta_\alpha$. If $\beta_\alpha$ is positive,
it means that the absolute error decreases as the true value of \gls{alpha}
increases.}}

% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Thu Jun 16 16:24:45 2016
\begin{table}[ht]
\centering
\begin{tabular}{llll}
  \hline
Parameter & Estimate & Standard error & $p$-value \\ 
  \hline
(Intercept) & $2$ & $0.6$ & $0.01$ \\ 
  $\alpha$ & $10$ & $2$ & ${<}10^{-5}$ \\ 
  I & $-3\!\times\!10^{-4}$ & $2\!\times\!10^{-4}$ & $0.7$ \\ 
  m & $0.5$ & $0.2$ & $0.01$ \\ 
   \hline
\end{tabular}
\caption[Parameters of a fitted GLM relating error in estimated $\alpha$ to true values of BA parameters.]{Parameters of a fitted GLM relating error in estimated $\alpha$ to true values of BA parameters. GLM was fitted with a Gaussian distribution and inverse link function. Coefficients are interpretable as additive effects on the inverse of the mean error.} 
\label{tab:glmalpha}
\end{table}
% latex table generated in R 3.2.3 by xtable 1.8-2 package
% Thu Jun 16 16:24:45 2016
\begin{table}[ht]
\centering
\begin{tabular}{llll}
  \hline
Parameter & Estimate & Standard error & $p$-value \\ 
  \hline
(Intercept) & $0.004$ & $5\!\times\!10^{-4}$ & ${<}10^{-5}$ \\ 
  $\alpha$ & $-0.001$ & $2\!\times\!10^{-4}$ & ${<}10^{-5}$ \\ 
  I & $-4\!\times\!10^{-7}$ & $2\!\times\!10^{-7}$ & $0.05$ \\ 
  m & $-7\!\times\!10^{-5}$ & $8\!\times\!10^{-5}$ & $1$ \\ 
   \hline
\end{tabular}
\caption[Parameters of a fitted GLM relating error in estimated $I$ to true values of BA parameters.]{Parameters of a fitted GLM relating error in estimated $I$ to true values of BA parameters. GLM was fitted with a gaussian distribution and inverse link function. Coefficients are interpretable as additive effects on the inverse of the mean error.} 
\label{tab:glmI}
\end{table}

{\color{blue}\uline{The \gls{GLM} analysis indicated that the error in estimates of
\gls{alpha} decreased with larger true values of \gls{alpha} 
    ($p {<}10^{-5}$)
and \gls{m}
    ($p =0.01$)
but was not significatly affected by \gls{I}} (\cref{tab:glmalpha}). 
\uline{Qualitatively, \gls{alpha} seemed to be only weakly identifiable between
the values of 0 and 0.5} (\cref{fig:abcptm2}). \uline{The error in the
estimated prevalence \gls{I} was slightly lower for smaller values of
\gls{alpha}
    ($p {<}10^{-5}$)
and \gls{I}
    ($p =0.05$),
but was not significantly affected by the true value of \gls{m}} 
(\cref{tab:glmI}).}

{\color{red}\sout{
The accuracy of the parameter estimates obtained with ABC
paralleled the results from the \gls{kSVR} classifier. Of the four parameters,
$\alpha$ was the most accurately estimated, with point estimates having a
median [IQR] absolute error of 
    0.11 
    [0.03 - 
    0.25].
The errors when the true value of $\alpha$ was zero were significantly greater
than those for the other values 
    (Wilcoxon rank-sum test, $p$ = $0$).
Errors in estimating $\alpha$ 
%also varied with the true value of $m$ just at
%the threshold of statistical significance
%    (one-way ANOVA, $p 
%    pp(anova(lm(alpha_error~true_m, d))[1, "Pr(>F)"], eq=TRUE)$),
did not vary across the true values of $m$ or $I$ (both
one-way ANOVA).}

\sout{Estimates for $I$ were relatively accurate, with point estimate errors of
    492 
    [294 - 
    782] individuals.
These errors were significantly higher when the true value of $\alpha$ was
at least 1
    (Wilcoxon rank-sum test, $p$ = $0$)
and when the true value of $I$ was 2000 ($p < 10^{-5}$). The true value of $m$
did not affect the estimates of $I$ (one-way ANOVA).}

\sout{The $m$ parameter was estimated correctly in
    37 \%
of simulations barely better than random guessing. The true values of the other
parameters did not significantly affect the
estimates of $m$ (both one-way ANOVA).}

\sout{Finally, the total number of nodes $N$ was consistently over-estimated by about
a factor of two (error 4153
[3660 - 4489] individuals). No parameters influenced the accuracy of the $N$
estimates (all one-way ANOVA).}}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{abc-point-estimate-m2}
  \vspace{6pt}
  \caption[
    Posterior mean point estimates for \gls{BA} model parameters obtained by
    running \software{netabc} on simulated data, for simulations with \gls{m} =
    2.
  ]{
    Posterior mean point estimates for \gls{BA} model parameters obtained by
    running \software{netabc} on simulated data, for simulations with \gls{m} =
    2. Dashed lines indicate true values. (A) Estimates of \gls{alpha} and
    \gls{I} which were varied in these simulations against known values. (B)
    Estimates of \gls{m} and \gls{N} which were held fixed in these simulations
    at the values \gls{m} = 2 and \gls{N} = 5000.
  }
  \label{fig:abcptm2}
\end{figure}

\begin{table*}[ht]
  \centering
  \input{\tablepath/abc-hpd.tex}
  \caption[
      Average posterior mean point estimates and 95\% highest posterior density
      (HPD) interval widths for BA model parameter estimates obtained with
      ABC.
  ]{
      Average posterior mean point estimates and 95\% highest posterior density
      (HPD) interval widths for BA model parameter estimates obtained with
      ABC. Three transmission trees were simulated under each
      combination of the listed parameter values, and the parameters were
      estimated with ABC without training.
  }
  \label{tab:abchpd}
\end{table*}

The dispersion of the ABC approximation to the posterior also varied between
the parameters, with narrower HPD intervals for the parameters with more
accurate point estimates (\cref{tab:abchpd}). \Cref{fig:abcex} shows
the distributions for one simulation. Equivalent plots for one replicate
simulation with each studied parameter combination can be found in
\cref{fig:0.0-1000-2-5000-0,fig:0.5-1000-2-5000-0,fig:1.0-1000-2-5000-0,fig:1.5-1000-2-5000-0,fig:0.0-2000-2-5000-0,fig:0.5-2000-2-5000-0,fig:1.0-2000-2-5000-0,fig:1.5-2000-2-5000-0,fig:0.0-1000-3-5000-0,fig:0.5-1000-3-5000-0,fig:1.0-1000-3-5000-0,fig:1.5-1000-3-5000-0,fig:0.0-2000-3-5000-0,fig:0.5-2000-3-5000-0,fig:1.0-2000-3-5000-0,fig:1.5-2000-3-5000-0,fig:0.0-1000-4-5000-0,fig:0.5-1000-4-5000-0,fig:1.0-1000-4-5000-0,fig:1.5-1000-4-5000-0,fig:0.0-2000-4-5000-0,fig:0.5-2000-4-5000-0,fig:1.0-2000-4-5000-0,fig:1.5-2000-4-5000-0}. HPD intervals around $\alpha$ and $I$ were often narrow
relative to the region of nonzero prior density, whereas the intervals for $m$
and $N$ were more widely dispersed.

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{abc-posterior-example}
  \vspace{6pt}
  \caption[
    Marginal posterior distributions of BA model parameters estimated
    with ABC for a single simulated transmission tree.
  ]{
    Marginal posterior distributions of BA model parameters estimated
    with ABC for a single simulated transmission tree. Dotted
    lines indicate true values, solid lines indicate posterior mean point
    estimates, and shaded areas show 95\% highest posterior density intervals.
  }
  \label{fig:abcex}
\end{figure}

To test the effect of model misspecification, we simulated one network where
the nodes exhibited heterogeneous preferential attachment power (half 0.5, the
other half 1.5), with $m$ = 2, $N$ = 5000, and $I$ = 1000. The posterior mean
[95\% HPD] estimates for each parameter were: 
$\alpha$, 
  1.03 
  [0.67 -
   1.18];
$I$,
  1474 
  [511 -
   2990];
$m$,
  3 
  [1 -
   5];
$N$,
  9861 
  [3710-
   14977].
The approximate posterior distributions for this simulation are shown in
\cref{fig:mixed}. To test the effect of sampling bias, we sampled one
transmission tree in a peer-driven fashion, where the probability to sample a
node was twice as high if one of its peers had already been sampled. The
parameters for this experiment were $N$ = 5000, $m$ = 2, $\alpha$ = 0.5, and
$I$ = 2000. The estimated values were
$\alpha$, 
  0.3 
  [0 -
   0.63];
$I$,
  2449 
  [1417 -
   3811];
$m$,
  3 
  [2 -
   5];
$N$,
  9132 
  [2852 -
   14780].
The approximate posterior distributions are shown in \cref{fig:peerdriven}. Both
of these results were in line with estimates obtained on other simulated
datasets (\cref{tab:abchpd}), although the estimate of peer-driven sampling for
$\alpha$ was somewhat lower than typical.

\section{Application to real world HIV data}
\label{sec:hiv}

\subsection{Methods}

{\color{blue}\uline{
The synthetic data experiments presented in the previous section demonstrated 
that the \gls{alpha} and \gls{I} parameters, and to a lesser extent the \gls{N}
parameter, had measurable impacts on tree shape and were identifiable with
\gls{ABC}. Therefore, our next step was to use
\software{netabc} to estimate the parameters of the \gls{BA} model for real
world \gls{HIV} epidemics. }} Because the \gls{BA} model assumes a single
connected contact network, it is most appropriate to apply to groups of
individuals who are epidemiologically related. Therefore, we searched for
published \gls{HIV} datasets which originated from existing clusters, either
phylogenetically or geographically defined. In addition, we analysed an
in-house dataset sampled from \gls{HIV}-positive individuals in British
Columbia, Canada (the ``BC data''). The datasets are summarized in
\cref{tab:data}.

\begin{table}[ht]
  \centering
  \input{\tablepath/realdata-bc}
  \caption[Characteristics of published HIV datasets analyzed with \software{netabc}.]
  {
    Characteristics of published HIV datasets analyzed with \software{netabc}. 
    Abbreviations: MSM, men who have sex with men; HET, heterosexual; IDU,
    injection drug users. The \textcite{novitsky2013phylogenetic,novitsky2014impact} data
    were sampled from a primarily heterosexual risk environment but did not
    explicitly exclude other risk factors.
  }
  \label{tab:data}
\end{table}

We downloaded all sequences associated with each published study from GenBank.
{\color{blue}\uline{For the \textcite{niculescu2015recent} data, only sequences
from \gls{IDU} (whose sequence identifiers included the letters ``DU'') were
included in the analysis. \textcite{kao2011surveillance} found a very strong
association in their study population between subtype and risk group - subtype
B was almost always associated with \gls{MSM}, whereas \gls{IDU} were usually
infected with a circulating recombinant form. Since there were many more
subtype B sequences in their data than sequences of other subtypes, we
restricted our analysis to the subtype B sequences and labelled this dataset as
\gls{MSM}. Three datasets (\textcite{zetterberg2004two},
\textcite{grabowski2014role}, and \textcite{mccormack2002early}) included both
\textit{env} and \textit{gag} sequences. Each gene was analyzed separately to
assess the robustness of \software{netabc} to the particular \gls{HIV} gene
sequence used to estimate a transmission tree.}}

{\color{red}\sout{For the \textcite{novitsky2014impact} data,}} Each
\textit{env} sequence was aligned pairwise to the HXB2 reference sequence
(GenBank accession number K03455), and the hypervariable regions were clipped
out with \software{BioPython} version 1.66+~\autocite{cock2009biopython}.
Sequences were multiply aligned using \software{MUSCLE} version 3.8.31
\autocite{edgar2004muscle}, and alignments were manually inspected with
\software{Seaview} version 4.4.2 \autocite{gouy2010seaview}. Phylogenies were
constructed from the nucleotide alignments by approximate maximum likelihood
using \software{FastTree2} version 2.1.7 \autocite{price2010fasttree} with the
\gls{GTR} model~\autocite{tavare1986some}. Transmission trees were estimated by
rooting and time-scaling the phylogenies by root-to-tip regression, using a
modified version of Path-O-Gen (distributed as part of
BEAST~\autocite{drummond2007beast}) as described
previously~\autocite{poon2015phylodynamic}.

{\color{red}\sout{Three}} 
{\color{blue}\uline{Six}} of the datasets 
\autocite[][and the BC data]{li2015hiv,novitsky2014impact,grabowski2014role,little2014using}
were initially larger than the others, containing 1265, 1299, 1026/915
(\textit{env}/\textit{gag}), 648, and 7923 sequences respectively. To ensure
that the analyses were comparable, we reduced these to a number of sequences
similar to the smaller datasets. For the \citeauthor{li2015hiv} and BC
datasets, we detected clusters of size 280 and 399 respectively using a
patristic distance cutoff of 0.02 as described
previously~\autocite{poon2015impact}. Only sequences within these clusters were
carried forward. For the
\textcite{grabowski2014role,novitsky2014impact,little2014using} datasets, no
large clusters were detected using the same cutoff, so we analysed subsets of
size 225, 180, and 180 respectively. {\color{blue}\uline{The subset of the
\citeauthor{grabowski2014role} data was chosen by eye such that the individuals
were monophyletic in both the \textit{gag} and \textit{env} trees. The other
subsets were arbitrarily chosen subtrees from phylogenies of the complete
datasets.}}

For all datasets, we used the priors $\alpha$ $\sim$ Uniform(0, 2) and $N$ and
$I$ jointly uniform on the region \{$n \leq N \leq 10000$, $n \leq I \leq
10000$, $I \leq N$\}, where $n$ is the number of tips in the tree (see
\cref{tab:data}). Since the value $m = 1$ produces networks with no cycles,
which we considered fairly implausible, we ran one analysis with the prior $m
\sim$ DiscreteUniform(1, 5), and one with the prior $m \sim$ DiscreteUniform(2,
5). The other parameters to the SMC algorithm were the same as used for the
simulation experiments, except that we used 10000 particles instead of 1000 to
increase the accuracy of the estimated posterior. This was computationally
feasible due to the small number of runs required for this analysis.

Empirical studies of contact networks often report the exponent $\gamma$ of the
power law degree distribution. To compare our results to the literature, we
simulated 100 networks each according to the posterior mean parameter estimates
obtained for each investigated dataset. The power-law exponent $\gamma$ was
calculated for each network using the \software{fit\_power\_law} function in
\software{igraph}, with the `R.mle' implementation. The median of the 100
$\gamma$ values was taken as a point estimate for the associated dataset.

\subsection{Results}

We applied \software{netabc} to {\color{red}\sout{five}}
{\color{blue}\uline{ten}} published HIV datasets, {\color{blue}\uline{plus
one in-house, unpublished dataset}} (\cref{tab:data}).
{\color{red}\sout{and found substantial heterogeneity among the parameter
estimates}} {\color{blue}\uline{Posterior mean point estimates and 50\% and
95\% \glspl{HPD} for each parameter are shown in} \cref{fig:abchpd}, \uline{and
marginal posterior distributions in}
\cref{fig:cuevas,fig:li,fig:niculescu,fig:novitsky,fig:wang}.
\Cref{fig:abchpdm2} \uline{shows point estimates and \glspl{HPD} obtained when
the value $\gls{m} = 1$ was disallowed by the prior. Since the results
indicated that $\gls{m} = 1$ was the most credible value for several datasets, 
all results discussed henceforth are for the prior $\gls{m} \sim
\DiscreteUniform(1, 5)$ unless otherwise stated. For clarity, datasets will be
referred to as their risk group and location. For example, \gls{IDU}/Estonia
refers to the} \textcite{zetterberg2004two} \uline{dataset.}}

{\color{red}\sout{
Two of the datasets (\textcite{niculescu2015recent, wang2015targeting}) had
estimated $\alpha$ values near unity for the prior allowing $m = 1$ (posterior
mean [95\% \gls{HPD}] 
  0.73 
  [0.05 - 
   1.18]
and
  0.55 
  [0.01 -
   0.99] respectively).
The estimates did not change appreciably when $m = 1$ was disallowed by the
prior, although the credible interval of the \textcite{niculescu2015recent}
data was narrower
  (0.05 - 
   1.18).
When $m = 1$ was permitted, the \textcite{li2015hiv, cuevas2009hiv} both had
low estimated $\alpha$ values
  (0.33 
  [0 - 
  0.76]
and
  0.27 
  [0 -
   0.59]). 
However, the estimates increased when $m = 1$ was not permitted, although the
HPD intervals remained roughly the same
  (0.58 
  [0.06 - 
  0.99]
and
  0.48 
  [0.02 -
   0.87]).
The \textcite{novitsky2014impact} data had a fairly low estimated $\alpha$
for both priors on $m$
  (0.55 for $m \geq 1$;
   0.53 for $m \geq 2$).
However, the confidence interval was much wider when $m = 1$ was allowed
  ([0 -
    1.75] for $m \geq 1$ vs.
   [0 -
    1.75] for $m \geq 2$). }}

{\color{blue}\uline{Posterior mean point estimates for the \acrlong{PA} power
\gls{alpha} ranged from 
    0.9
for the \gls{IDU}/Estonia data to
    0.27
for the mixed/Spain data. When aggregated by risk group, the average estimates
were
    0.73
for \gls{IDU},
    0.43
for primarily heterosexual risk, and
    0.39
for \gls{MSM}. These values were obtained with \textit{gag} for the datasets
where both \textit{gag} and \textit{env} were sequenced, but the estimates for
\gls{alpha} did not change appreciably between the two genes} (TODO figure).
\uline{There was a large amount of uncertainty associated with these
estimates. 95\% \gls{HPD} were very wide for most datasets, often encompassing
almost the entire range from 0 to 1} (\cref{fig:abchpd}). \uline{For all the
datasets except HET/Malawi and MSM/Beijing, the posterior mean was either
outside, or very close to the border of, the 50\% \gls{HPD}. This indicates
that the posterior distributions were diffuse with heavy tails, rather than
having most of their mass around the mode.}}

{\color{red}\sout{For all the datasets except \citeauthor{novitsky2014impact},
estimated values of $I$ were below 2000 when $m = 1$ was allowed, with
relatively narrow HPD intervals compared to the nonzero prior density region
  (\citeauthor{cuevas2009hiv}, 701 
  [289 -
   1279];
   \citeauthor{niculescu2015recent}, 747
  [136 - 
   2378];
  \citeauthor{li2015hiv}, 1390 
  [310 -
   2821];
   \citeauthor{wang2015targeting}, 675
  [175 - 
   1400]).
The \citeauthor{novitsky2014impact} data was the outlier, with a very high
estimated $I$, and HPD interval spanning almost the entire prior region
  (5431 
  [183 -
   8739]).
The $I$ estimates and HPD intervals were generally robust to the choice of
prior on $m$, with slightly narrower HPD intervals (compare
\cref{fig:abchpd,fig:abchpdm2}).}}

The posterior mean of $m$ was equal to 1 for all but the
\citeauthor{novitsky2014impact} data, when this value was allowed. However, the
upper bound of the HPD interval was different for each dataset
  (\citeauthor{niculescu2015recent}, 4;
   \citeauthor{wang2015targeting}, 1;
   \citeauthor{li2015hiv}, 1;
   \citeauthor{cuevas2009hiv}, 1).
When $m = 1$ was disallowed, the estimate for all datasets was either 2 or 3,
with HPD intervals spanning the entire prior region. The estimates for the
total number of nodes $N$ were largely uninformative for all samples, with
almost all posterior mean estimates greater than 7500 and HPD intervals
spanning almost the entire nonzero prior density region. The only exception was
the \citeauthor{li2015hiv} data, for which the estimate was lower 
  (5916)
when $m = 1$ was allowed.

\begin{figure*}[ht]
  \centering
  \includegraphics{realdata-hpd-bc}
  \caption[
      Maximum \textit{a posteriori} point estimates and 95\% HPD intervals for
      parameters of the BA network model, fitted to six HIV datasets
      with \software{netabc}.]
  {
      Maximum \textit{a posteriori} point estimates and 95\% HPD intervals for
      parameters of the BA network model, fitted to six HIV datasets with
      \software{netabc}. Legend labels indicate risk group and country of
      origin. Abbreviations: IDU, injection drug users; MSM, men who have sex
      with men; HET, heterosexual.
  }
  \label{fig:abchpd}
\end{figure*}

To make our analyses comparable to the existing network literature, we
estimated values of the power law exponent $\gamma$ for each of the datasets 
investigated, using the posterior median estimates with the prior allowing
\gls{m} = 1. The results are summarized in \cref{tab:gamma}. All of the
estimated exponents were in the range $2 \leq \gamma \leq 2.5$, which is on the
lower end of the range $2 \leq \gamma \leq 4$ reported in the literature.

\begin{table}
    \centering
    \input{\tablepath/realdata-gamma.tex}
    \caption[
        Estimated power law exponents for six HIV datasets based on maximum
        \textit{a priori} estimates of BA model parameters.
    ]{
        Estimated power law exponents for six HIV datasets based on maximum
        \textit{a priori} estimates of BA model parameters. 100 networks were
        simulated using posterior mean parameter estimates obtained with
        \software{netabc}. The power law exponent $\gamma$ was estimated for
        each, and the median of those estimates was used as a point estimate
        for the corresponding dataset.
    }
    \label{tab:gamma}
\end{table}