### 2016-06-23 doc/thesis/revision_1_summary.tex

\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{mathptmx}
\usepackage{amsmath}
\begin{document}

\section*{Major changes}

\subsection*{Structure}

\begin{itemize}
  \item Took likelihood calculations out of objective section and moved them
    into research chapter (sec. 2.1.4 "Justification for approach").
  \item Instead of displaying many posterior distributions from synthetic data 
    in the appendices, only show two in the main text: one "good" (ie. low
    error rate) and one "bad".
  \item Structure of chapter 2 has been rearranged: one section about method,
    one about synthetic data experiments, one about real data, and then the
    discussion. 
\end{itemize}

\subsection*{Analysis}

\begin{itemize}
  \item It turns out that only networks with alpha = 1 have a power law
    distribution (p. 12). However power laws still capture the salient features
    of the degree distributions (fig. A26) so I continue to report them.
  \item Throughout, maximum a posteriori estimates were replaced with posterior
    means, which are more conventional and give better results.
  \item Removed ad-hoc statistical tests from results on grid search and ABC.
    For grid search, only error distributions were reported numerically.
    Qualitative statements were made about kernel score distributions. For ABC,
    effects of each parameter on error rates was analysed with a p-value
    corrected GLM (p. 52-53).
  \item Two extra analyses to address possible sources of ABC error (a reviewer
    asked for this) described on p. 46, results on p. 62-63.
  \item Show some 2-dimensional marginals (p. 55, 57).
  \item The real data analysis has been significantly expanded to include six
    more datasets.
  \item Analysed three datasets with both env and gag genes and compare results
    (fig. 2.17 on p. 65).
  \item Found several published estimates of power law exponent gamma from
    literature (table 2.9 on p. 74). 
\end{itemize}

\subsection*{Writing}

\begin{itemize}
  \item Added some introduction to SMC section to clarify what the relationship
    to our problem is, ie. we are trying to make a particle approximation to
    the posterior.
  \item Same thing for ABC section (p. 21-23).
  \item Added summary of chapter 1 (p. 26-27).
  \item Added introductory paragraph to chapter 2 (p. 28).
  \item Expanded on why ABC is needed in this setting (p. 33-36), why we should
    study the BA model (p. 36-37).
  \item Wrote high-level summary and motivation of all synthetic data
    experiments (p. 38-39, 40, 44).
  \item Results section for real data experiments has been entirely rewritten.
  \item Discussion of real data results is also entirely new. I relate our
    results to a couple of papers looking at preferential attachment (PA)
    networks, one which found sub-linear PA for real world data (which we also
    found), and one which claimed IDU networks were more PA-ish than sexual
    networks (which we also found too).
\end{itemize}

\section*{Minor changes}

\begin{itemize}
  \item Added text to abstract (UBC limit is 350 words).
  \item Better definitions for "production rule", "decay factor", and
    "ladderize" (p. 8-9).
  \item More conventional definition of importance sampling (p. 16).\
  \item Added subsection under SMC for actual SMC (SIS with resampling), plus
    discussion of particle degeneracy (p. 18).
  \item Throughout, multi-step mathematical derivations have had justifications
    added beside each step (eg. p. 20).
  \item Added a paragraph about the importance of the distance function for ABC
    (p. 23-24).
  \item Changed "$\alpha$" from Del Moral paper to "$\alpha_{\text{ESS}}$" to avoid
    confusion (p. 32).
  \item Fig. 2.7 is now in color.
  \item Added notches for 50\% highest density intervals (similar to IQR) in
    fig. 2.8.
  \item Reported relative errors for I and N in grid search.
  \item Refer to datasets as eg. "IDU/Romania data" instead of "Niculescu et
    al. data" for easier interpretation.
  \item Proper scaling and legends for fig. 2.18 (p. 69).
\end{itemize}

\end{document}