### 2016-06-25 doc/paper/coverletter.tex

\documentclass[12pt]{letter}
\usepackage[T1]{fontenc}                                                         
\usepackage{garamondx}                                                           
\usepackage[garamondx,cmbraces]{newtxmath}

\signature{Rosemary M. McCloskey}
\address{BC Centre for Excellence in HIV/AIDS \\ 608-1081 Burrard Street \\ Vancouver, BC, Canada V6Z 1Y6}
\begin{document}
\begin{letter}{ }

\opening{Dear editors of \textit{Virus Evolution},}

On behalf of my coauthors and myself, I am submitting a manuscript entitled
``\textit{Reconstructing contact network parameters from viral phylogenies}''
for your consideration.

In this work, we have developed a method to fit contact network models to viral
phylogenetic data. These models capture the contact heterogeneity and
population structure often observed in real host communities. It has previously
been demonstrated analytically and through simulation studies that network
structure can affect the speed and pattern of epidemic spread. Consequently,
the often-made assumption of a homogeneously mixed population may introduce
errors into epidemiological predictions. Furthermore, the network itself may be
of interest from a public health perspective, such as for investigating the
presence of highly-connected ``superspreader'' nodes.

Despite their potential usefulness, most network parameters are extraordinarily
difficult to estimate with existing methods. Direct epidemiological
investigation of contact networks requires extensive contact tracing and may be
hampered by misreporting. Viral genetic data are much easier to collect, but as
far as we are aware, no other phylodynamic methods have been developed to make
use of this data to fit network models. Therefore, our work opens new avenues
for phylodynamic investigation, and we believe it will be of broad interest to
the molecular evolution and epidemiology community.

\closing{Thank you for your time,}
\end{letter}
\end{document}