### 2016-06-25 doc/paper/paper.Rnw

\documentclass[12pt]{article}
\usepackage[T1]{fontenc}
\usepackage{mathptmx}
\usepackage[backend=biber, style=numeric-comp, sorting=none]{biblatex}
\usepackage{fullpage}
\usepackage{array}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{cleveref}
\usepackage[normalem]{ulem}
\usepackage{color}
\usepackage{amsmath}

\addbibresource{papers.bib}

% beginning of diff stuff
\usepackage{twoopt}

\let\mtextcite\textcite
\let\mautocite\autocite
\let\mciteauthor\citeauthor
\renewcommandtwoopt{\textcite}[3][][]{\mbox{\mtextcite[#1][#2]{#3}}}
\renewcommandtwoopt{\autocite}[3][][]{\mbox{\mautocite[#1][#2]{#3}}}
\renewcommand{\citeauthor}[1]{\mbox{\mciteauthor{#1}}}

\newcommand{\add}[1]{\color{blue} \uline{#1} \color{black}}
\newcommand{\del}[1]{\color{red} \sout{#1} \color{black}}
% end of diff stuff

\newcommand{\tablepath}{../tables}
\graphicspath{{../figures/}}

\begin{document}
\onehalfspacing

\title{Reconstructing contact network parameters from viral phylogenies}
\author{Rosemary M. McCloskey$^1$, Richard H. Liang$^1$, and Art F.Y. Poon$^{1,2}$ \\
\small $^1$BC Centre for Excellence in HIV/AIDS, Vancouver, Canada \\ \small $^2$ Department of Medicine, University of British Columbia, Vancouver, Canada}

<<setup, include=FALSE>>=
    library(netabc)
    library(Hmisc)

    pp <- function (p) {
        if (p < 1e-5) {
            "{<}10^{-5}"
        } else {
            paste0("=", latexSN(round(p, -floor(log10(p)))))
        }
    }
@

\maketitle
\abstract{\input{abstract}}

\section*{Introduction}

When an infectious disease spreads through a population, transmissions are
generally more likely to occur between certain pairs of individuals. Such pairs
must have a particular mode of contact with one another, which varies with the
mode of transmission of the disease. For airborne pathogens, physical proximity
may be sufficient, while for sexually transmitted diseases, sexual or in some
cases blood-to-blood contact is required. The population together with the set
of links between individuals along which transmission can occur is called the
contact network~\autocite{klovdahl1985social, morris1993epidemiology}. The
structure of the contact network underlying an epidemic can profoundly impact
the speed and pattern of the epidemic's expansion. Network structure can
influence the prevalence trajectory~\autocite{o2011contact, ma2013importance}
and epidemic threshold~\autocite{barthelemy2005dynamical}, in turn affecting
the estimates of quantities such as effective population
size~\autocite{goodreau2006assessing}.  From a public health perspective,
contact networks have been explored as tools for curtailing epidemic spread, by
way of interventions targeted to well-connected
nodes~\autocite{wang2015targeting}. True contact networks are a challenging
type of data to collect, requiring extensive epidemiological
investigation~\autocite{welch2011statistical, eames2015six}.

Viral sequence data, on the other hand, has become relatively inexpensive and
straightforward to collect on a population level. Due to the high mutation rate
of RNA viruses, epidemiological processes impact the course of viral evolution,
thereby shaping the inter-host viral
phylogeny~\autocite{drummond2003measurably}. The term ``phylodynamics'' was
coined to describe this interaction, as well as the growing family of inference
methods to estimate epidemiological parameters from viral
phylogenies~\autocite{grenfell2004unifying}. These methods have revealed
diverse properties of local viral outbreaks, from basic reproductive
number~\autocite{stadler2012estimating}, to the degree of
clustering~\autocite{hughes2009molecular}, to the elevated transmission risk
during acute infection~\autocite{volz2012simple}. On the other hand, although
sophisticated methods have been developed for fitting complex population
genetic models to phylogenies~\autocite{volz2012complex,
rasmussen2014phylodynamic}, inference of structural network parameters has to
date been limited. However, it has been shown that network structure has a
tangible impact on phylogeny shape~\autocite{leventhal2012inferring,
colijn2014phylogenetic, goodreau2006assessing, robinson2013dynamics,
villandre2016assessment}, suggesting that such statistical inference might be
possible~\autocite{welch2011statistical}.

Survey-based studies of sexual networks~\autocite{colgate1989risk,
liljeros2001web, schneeberger2004scale, latora2006network, rothenberg2007large,
clemenccon2015statistical} have found that these networks tend to have a degree
distribution which follows a power law \autocite[although there has been some
disagreement, see][]{handcock2004likelihood}. That is, the number of nodes of
degree $k$ is proportional to $k^{-\gamma}$ for some constant $\gamma$. These
networks are also referred to as
``scale-free''~\autocite{barabasi1999emergence}. One process by which
scale-free networks can be generated is preferential attachment, where nodes
with a high number of contacts attract new connections at an elevated rate. The
first contact network model incorporating preferential attachment was
introduced by \textcite{barabasi1999emergence}, and is now referred to as the
Barab\'asi-Albert (BA) model. Under this model, networks are formed by
iteratively adding nodes with $m$ new edges each. In the most commonly studied
formulation, these new edges are joined to existing nodes of degree $k$ with
probability proportional to $k$, so that nodes of high degree tend to attract
more connections. \citeauthor{barabasi1999emergence} suggested an extension
where the probability of attaching to a node of degree $k$ is $k^\alpha$ for
some non-negative constant $\alpha$, and we use this extension in this work.
\add{When $\alpha \neq 1$, the degree distribution is no longer a power law:
for $\alpha < 1$, the distribution is a stretched exponential, while for
$\alpha > 1$, it is a ``gelation'' type distribution where one or a few hub
nodes are connected to nearly every other node in the graph.}

Previous work offers precedent for the possibility of statistical inference of
structural network parameters. \textcite{britton2002bayesian} develop a Bayesian
approach to estimate the edge density in an Erd\H{o}s-R\'enyi
network~\autocite{erdos1960evolution} given observed infection dates, and
optionally recovery dates. Their approach was later extended by
\textcite{groendyke2011bayesian} and applied to a much larger data set of 188
individuals. \textcite{volz2007susceptible, volz2008sir} developed differential
equations describing the spread of a susceptible-infected (SI) epidemic on
static and dynamic contact networks with several degree distributions, which
could in principle be used for inference if observed incidence trajectories
were available. \textcite{brown2011transmission} analysed the degree distribution
of an approximate transmission network, estimated based on genetic similarity
and estimated times of infection, relating 60\% of HIV-infected men who have
sex with men (MSM) in the United Kingdom. The transmission network is a
subgraph of the contact network which includes only those edges which have
already led to a new infection. The authors found that a Waring distribution,
which is produced by a more sophisticated preferential attachment model, was a
good fit to their estimated network.

Standard methods of model fitting involve calculation of the likelihood of
observed data under the model. In maximum likelihood estimation, a quantity
proportional to the likelihood is optimized, often through a standard
multi-dimensional numerical optimization procedure. Bayesian methods integrate
prior information by optimizing the posterior probability instead. To avoid
calculation of a normalizing constant, Bayesian inference is often performed
using Markov chain Monte Carlo (MCMC), which uses likelihood \emph{ratios} in
which the normalizing constants cancel out. Unfortunately, it is generally
difficult to explicitly calculate the likelihood of an observed transmission
tree under a contact network model, even up to a normalizing constant. To do
so, it would be necessary to integrate over all possible networks, and also
over all possible labellings of the internal nodes of the transmission tree.
While it is not known (to us) whether such integration is tractable, a simpler
alternative is offered by likelihood-free methods, namely approximate Bayesian
computation (ABC)~\autocite{tavare1997inferring, beaumont2002approximate}. ABC
leverages the fact that, although calculating the likelihood may be
impractical, generating simulated datasets according to a model is often
straightforward. If our model fits the data well, the simulated data it
produces should be similar to the observed data. More formally, if $D$ is the
observed data, the posterior distribution $f(\theta \mid D)$ on model
parameters $\theta$ is replaced as the target of statistical inference by
$f(\theta \mid \rho(\hat{D}, D) < \varepsilon)$, where $\rho$ is a distance
function, $\hat{D}$ is a simulated dataset according to $\theta$, and
$\varepsilon$ is a small tolerance~\autocite{sunnaaker2013approximate}. \del{In
the specific case when $\rho$ is a kernel function, the approach is known as
ABC~\autocite{nakagome2013kernel, poon2015phylodynamic}.} \add{Our
group~\autocite{poon2015phylodynamic} and others~\autocite{park2015k2} have
demonstrated that taking $\rho$ to be kernel function TODO}

Here, we develop a method using ABC to estimate the parameters of contact
network models from observed phylogenetic data. The distance function we use is
the tree kernel developed by \textcite{poon2013mapping}, which computes a
weighted dot product of the trees' representations in the space of all possible
subset trees. We apply the method to investigate the parameters of the BA
network model on a variety of simulated and real datasets. Our results show
that some network parameters can be inferred with reasonable accuracy, while
others \del{have a minimal detectable impact on tree shape and therefore cannot
be estimated accurately}\add{are weakly- or non-identifiable with ABC}. We also
find that these parameters can vary considerably between real epidemics from
different settings. \add{TODO: something about sub-linear PA and IDU}

\section*{Methods}

\subsection*{\textit{Netabc}: phylogenetic inference of contact network
parameters with ABC}

We have developed an ABC-based method to perform statistical inference of
contact network parameters from a transmission tree estimated from an observed
viral phylogeny. We implemented the adaptive sequential Monte Carlo (SMC)
algorithm for ABC developed by \textcite{del2012adaptive}. The SMC algorithm
keeps track of a population of parameter ``particles'', which are initially
sampled from the parameters' joint prior distribution. Several datasets are
simulated under the model of interest for each of the particles. In this case,
the datasets are transmission trees, which are generated by a two-step process.
First, a contact network is simulated according to the network model being fit.
Second, a transmission tree is simulated over that network with a Gillespie
simulation algorithm~\autocite{gillespie1976general}, in the same fashion as
several previous studies \autocite[\textit{e.g.}][]{robinson2013dynamics,
leventhal2012inferring}. The particles are weighted according to the similarity
between their associated simulated trees and the observed tree. To quantify
this similarity, we used the tree kernel developed by
\textcite{poon2013mapping}. Particles are iteratively perturbed by applying a
Metropolis-Hastings kernel and, if the move is accepted, simulating new
datasets under the new parameters. When a particle's weight drops to zero,
because its simulated trees are too dissimilar to the observed tree, the
particle is dropped from the population, and eventually replaced by a resampled
particle with a higher weight. As the algorithm progresses, the population
converges to a Monte Carlo approximation of the ABC target distribution, which
is assumed to approximate the desired posterior~\autocite{del2012adaptive,
sunnaaker2013approximate}.

\add{In the original formulation of ABC-SMC~\autocite{sisson2007sequential,
beaumont2009adaptive}, the user is required to specify a decreasing sequence of
tolerances $\lbrace\varepsilon_i\rbrace$. At iteration $i$, particles with no
associated simulated datasets within distance $\varepsilon_i$ of the observed
data are removed from the population. In the adaptive version of
\textcite{del2012adaptive}, the sequence of tolerances is determined
automatically by fixing the decay rate of the population's expected sample size
(ESS) to a user-defined value. \citeauthor{del2012adaptive} call this value
$\alpha$, but we will refer to it here as $\alpha_{\text{ESS}}$ to avoid
confusion with the preferential attachment power parameter of the BA model. A
computer program implementing our method is freely available at
\url{https://github.com/rmcclosk/netabc} (last accessed \today).}

To check that our implementation of Gillespie simulation was correct, we
reproduced Figure 1A of \textcite{leventhal2012inferring} (our
\cref{fig:leventhal}), which plots the unbalancedness of transmission trees
simulated over four network models at various levels of pathogen
transmissibility. Our implementation of adaptive ABC-SMC was tested by applying
it to the same mixture of Gaussians used by \citeauthor{del2012adaptive} to
demonstrate their method (originally used by~\textcite{sisson2007sequential}).
We were able to obtain a close approximation to the function (see
\cref{fig:smctest}), and attained the stopping condition used by the authors in
a comparable number of steps.

Nodes in our networks followed simple SI dynamics, meaning that they became
infected at a rate proportional to their number of infected neighbours, and
never recovered. For all analyses, the transmission trees' branch lengths were
scaled by dividing by their mean. We used the \textit{igraph} library's
implementation of the BA model~\autocite{csardi2006igraph} to generate the
graphs. The analyses were run on Westgrid (\url{https://www.westgrid.ca/}) and
a local computer cluster.

\subsection*{\del{Kernel classifiers}\add{Classifiers for BA model parameters from tree shapes}}

\add{We considered four parameters related to the BA model, denoted $N$, $m$,
$alpha$, and $I$. The first three of these parameterize the network structure,
while $I$ is related to the simulation of transmission trees over the network.
However, we will refer to all four as BA parameters. $N$ denotes the total
number of nodes in the network, or equivalently, susceptible individuals in the
population. $m$ is the number of new undirected edges added for each new
vertex, or equivalently one-half of the average degree. $\alpha$ is the power
of preferential attachment -- new nodes are attached to existing nodes of
degree $d$ with probability proportional to $d^{\alpha} + 1$. Finally, $I$ is
the number of infected individuals at the time when sampling occurs. The
$\alpha$ parameter is unitless, while $m$ has units of edges or connections per
vertex, and $N$ and $I$ both have units of nodes or individuals.}

\add{Before proceeding with a full validation of \textit{netabc} on simulated
data, we undertook two experiments designed to assess the identifiability of
the BA parameters. These experiments only investigated one parameter of the
BA model at a time while holding all others fixed, a strategy commonly used
when performing sensitivity analyses of mathematical models. This allowed us to
perform a fast preliminary analysis without dealing with the ``curse of
dimensionality'' of the full parameter space. We simulated trees under three
different values of each parameter, and asked how well we could tell the
different trees apart. The better we are able to distinguish the trees, the
more identifiability we might expect for the corresponding parameter when we
attempt to estimate it with ABC.}

\add{This experiment also had the secondary purpose of validating our choice of
the tree kernel as a distance measure in ABC. To tell the trees apart, we used
a classifier based on the tree kernel, but we also tested two other tree shape
statistics. Sackin's index~\autocite{shao1990tree} is a measure of tree
imbalance which not take branch lengths into account, considering only the
topology. The normalized lineages-through-time
\autocite[nLTT,][]{janzen2015approximate} compares two trees based on normalized
distributions of their branching times, and does not explicitly consider the
topology. In addition, the tree kernel can be tuned by adjusting the values of
the meta-parameters $\lambda$ and $\sigma$ (the ``decay factor'' and ``radial
basis function variance'', see \textcite{poon2013mapping}). The results of this
experiment were used to select values for these meta-parameters to carry
forward based on their accuracy in distinguishing the different trees.}

\del{We used the phylogenetic kernel developed by \textcite{poon2013mapping}}
To test whether the parameters of the BA model had a \add{measurable} effect on
tree shape, 100 networks were simulated under each of three different values of
$\alpha$: 0.5, 1.0, and 1.5 (300 networks total). The other parameters were
fixed to the following values: $N = 5000$, $I = 1000$, and $m = 2$. A
transmission tree with 500 tips was simulated over each network (300
transmission trees total). The 300 trees were compared pairwise with the tree
kernel to form a $300 \times 300$ kernel matrix. The kernel meta-parameters
$\lambda$ \del{(the ``decay factor''),} and $\sigma$ \del{(the ``radial basis
function variance'')~\autocite[see][]{poon2013mapping},} were set to 0.3 and 4
respectively. \add{We also computed a $300 \times 300$ matrix of pairwise nLTT
values, and a $1 \times 300$ vector of Sackin's index values.} We constructed 
\add{three classifiers for $\alpha$:} a kernel support vector regression (kSVR)
\add{from the kernel matrix} with the \textit{kernlab}
package~\autocite{zeileis2004kernlab}, an ordinary SVR \add{from the nLTT
matrix} with the e1071 package~\autocite{meyer2015e1071}, and a linear
regression from the Sackin's index values. The accuracy of each classifier was 
evaluated with 1000 two-fold cross validations.

Three similar experiments were performed for the other BA model parameters (one
experiment per parameter). $m$ was varied between 2, 3, and 4; $I$ between 500,
1000, and 2000; and $N$ between 3000, 5000, and 8000. The parameters not being
tested were fixed at the values $N$ = 5000, $I$ = 1000, $m$ = 2, and $\alpha$ =
1. Thus, we performed a total of four cross-validations \add{for each
classifier}, one for each of the BA model parameters $\alpha$, $I$, $m$, and
$N$. We repeated these four cross-validations with different values of
$\lambda$ (0.2, 0.3, and 0.4) and $\sigma$ ($2^{-3}$, $2^{-2}$, \ldots, $2^3$),
as well as on trees with differing numbers of tips (100, 500, and 1000).
\add{For the structural parameters $\alpha$, $m$, and $N$, the experiments were
repeated with three different values of $I$ (500, 1000, and 2000).} \del{and in
epidemics of differing size (500, 1000, and 2000).} The combination of the
number of sampled individuals (\textit{i.e.} the number of tips) and the
epidemic size (\textit{i.e.} $I$) will be referred to as an ``epidemic
scenario''. When evaluating the classifier for $I$, we did not consider trees
with 1000 tips, because one of the tested $I$ values was 500, and the number of
tips cannot be larger than $I$.

\del{For each of the four parameters, we also tested a linear regression against
Sackin's index~\autocite{shao1990tree} and an ordinary SVR based on the normalized
lineages-through-time (nLTT) statistic~\autocite{janzen2015approximate}.}

\subsection*{ABC simulations}

\add{We tested \textit{netabc} by jointly estimating the four parameters of the
BA model. We used the standard validation approach of simulating transmission
trees under the model with known parameter values and attempting to recover
those values with \textit{netabc}. The algorithm was not informed of any of
the true parameter values for the main set of simulations.} We simulated three
transmission trees, each with 500 tips, under every element of the Cartesian
product of these parameter values: $N$ = 5000, $I$ = \{1000, 2000\}, $m$ = \{2,
3, 4\}, and $\alpha$ = \{0.0, 0.5, 1, 1.5\}. This produced a total of 24
parameter combinations $\times$ three trees per combination = 72 trees total.
The adaptive ABC algorithm was applied to each tree with these priors: $m \sim$
DiscreteUniform(1, 5), $\alpha \sim$ Uniform(0, 2), and $(N, I)$ jointly
uniform on the region \{$500 \leq N \leq 15000$, $500 \leq I \leq 5000$, $I
\leq N$\}. Proposals for $\alpha$, $N$, and $I$ were Gaussian, while proposals
for $m$ were Poisson. Following \textcite{del2012adaptive} and
\textcite{beaumont2009adaptive}, the variance of all proposals was equal to the
empirical variance of the particles.

The SMC settings used were 1000 particles, 5 simulated datasets per particle,
and \del{the ``quality'' parameter controlling the decay rate of the tolerance
$\varepsilon$ set to}\add{$\alpha_{\text{ESS}} =  0.95$}. We used the same
stopping criterion as \citeauthor{del2012adaptive}, namely when the MCMC
acceptance rate dropped below 1.5\%. \del{Point estimates for the parameters
were obtained by taking the highest point of an estimated kernel density on the
final set of particles, calculated using the \textit{density} function with the
default parameters in \textit{R}.} \add{Approximate posterior means for the
parameters were obtained by taking the weighted average of the final set of
particles.} Highest posterior density (HPD) intervals were calculated with the
\textit{HPDinterval} function from the \textit{R} package
\textit{coda}~\autocite{plummer2006coda}.

\add{To evaluate the effects of the true parameter values on the accuracy of
the posterior mean estimates, we analyzed the $\alpha$ and $I$ parameters
individually using generalized linear models (GLMs) The response variable was
the error of the point estimate, and the predictor variables were the true
values of $\alpha$, $I$, and $m$. We did not test for differences across true
values of $N$, because $N$ was not varied in these simulations. The
distribution family and link function for the GLMs were Gaussian and inverse,
respectively, chosen by examination of residual plots and Akaike information
criteria (AIC). The $p$-values of the estimated GLM coefficients were corrected
using Holm-Bonferroni correction~\autocite{holm1979simple} with $n =
6$ (two GLMs with three predictors each). Because there was clearly little to
no identifiability of $N$ and $m$ with ABC (see results in next section), we
did not construct GLMs for those parameters.}

Two further simulations were performed to address \del{potential sources of
error} \add{the possible impact of two types of model misspecification}. To
evaluate the effect of model misspecification in the case of heterogeneity
among nodes, we generated a network where half the nodes were attached with
power $\alpha$ = 0.5, and the other half with power $\alpha$ = 1.5. The other
parameters for this network were $N$ = 5000, $I$ = 1000, and $m$ = 2. To
investigate the effects of potential sampling bias, we simulated a transmission
tree where the tips were sampled in a peer-driven fashion, rather than at
random. That is, the probability to sample a node was twice as high if any of
that node's network peers had already been sampled. The parameters of this
network were $N= 5000$, $I = 2000$, $m = 2$, and $\alpha = 0.5$.

\add{Despite the fact that the parameter values used to generate the simulated
transmission trees were known, the true posterior distributions of the BA
parameters were unknown. Therefore, any apparent errors or biases in the
estimates could be due to either poor performance of our method, or to real
features of the posterior distribution. Two retrospective experiments were
performed to disambiguate some of the observed errors. To assess the impact of
the SMC settings on \textit{netabc}'s accuracy, we ran \textit{netabc}
twice on the same simulated transmission tree. For the first run, the SMC
settings were the same as in the other simulations: 1000 particles, 5 simulated
transmission trees per particle, and $\alpha_{\text{ESS}} = 0.95$. The second
run was performed with 2000 particles, 10 simulated transmission trees per
particle, and $\alpha_{\text{ESS}} = 0.99$. To investigate the extent to which
errors in the estimated BA parameters were due to true features of the
posterior, rather than an inaccurate ABC approximation, we performed marginal
estimation for one set of parameter values. Each combination of 1, 2, or 3
model parameters (14 combinations total) was fixed to their known values, and
the remaining parameters were estimated with \textit{netabc}. The parameter
values were $\alpha = 0.0$, $m = 2$, $I = 2000$, and $N = 5000$.}

\subsection*{Investigation of published data}

We applied our ABC method to ten published HIV datasets. Because the BA model
generates networks with a single connected component, we specifically searched
for datasets which originated from existing clusters, either phylogenetically
or geographically defined. Characteristics of the datasets we investigated are
given in \cref{tab:data}. \add{For clarity, we will refer to each dataset by
its risk group and location of origin in the text. For example, the
\textcite{zetterberg2004two} data will be referred to as IDU/Estonia.}

\begin{table*}[!t]
  \centering
  \input{\tablepath/realdata}
  \caption{
    Characteristics of published datasets investigated with ABC.
    Acronyms: MSM, men who have sex with men; IDU, injection drug users; HET,
    heterosexual. The HET data were sampled from a primarily heterosexual risk
    environment, but did not explicitly exclude other risk factors. The number
    of sequences column indicates how many sequences were included in our
    analysis; there may have been additional sequences linked to the study
    which we excluded for various reasons (see methods).
  }
  \label{tab:data}
\end{table*}

We downloaded all sequences associated with each published study from GenBank.
\add{For the IDU/Romania data, only sequences from injection drug users (IDU,
whose sequence identifiers included the letters ``DU'') were included in the
analysis. \textcite{kao2011surveillance} (MSM/Taiwan) found a strong
association in their study population between subtype and risk group - subtype
B was most often associated with men who have sex with men (MSM), whereas IDU
were usually infected with a circulating recombinant form. Since there were
many more subtype B sequences in their data than sequences of other subtypes,
we restricted our analysis to the subtype B sequences and labelled this dataset
as MSM. Three datasets (IDU/Estonia, HET/Uganda, and HET/Malawi) included both
\textit{env} and \textit{gag} sequences. Each gene was analyzed separately to
assess the robustness of \textit{netabc} to the particular HIV gene sequence
used to estimate a transmission tree.}

\del{For the \textcite{novitsky2014impact} data,} Each \textit{env} sequence
was aligned pairwise to the HXB2 reference sequence (GenBank accession number
K03455), and the hypervariable regions were clipped out with
\textit{BioPython} version 1.66+~\autocite{cock2009biopython}. Sequences were
multiply aligned using \textit{MUSCLE} version 3.8.31
\autocite{edgar2004muscle}, and alignments were manually inspected with
\textit{Seaview} version 4.4.2 \autocite{gouy2010seaview}. Phylogenies were
constructed from the nucleotide alignments by approximate maximum likelihood
using \textit{FastTree2} version 2.1.7 \autocite{price2010fasttree} with the
generalized time-reversible (GTR) model~\autocite{tavare1986some}. Transmission
trees were estimated by rooting and time-scaling the phylogenies by root-to-tip
regression, using a modified version of Path-O-Gen (distributed as part of
BEAST~\autocite{drummond2007beast}) as described
previously~\autocite{poon2015phylodynamic}.

\del{Two} \add{Four} of the datasets \del{\autocite{li2015hiv,novitsky2014impact}}
(MSM/Shanghai, HET/Botswana, HET/Uganda, and MSM/USA)
were initially much larger than the others, containing 1265, 1299,
\add{1026/915 (\textit{env}/\textit{gag}), and 648} sequences respectively. To
ensure that the analyses were comparable, we reduced these to a number of
sequences similar to the smaller datasets. For the MSM/Shanghai data, we
detected a cluster of size 280 using a patristic distance cutoff of 0.02 as
described previously~\autocite{poon2015impact}. Only sequences within this
cluster were carried forward. For the HET/Uganda, HET/Botswana, and MSM/USA
data, no large clusters were detected using the same cutoff, so we analysed
\del{a subtree} subsets of sizes 255, 180, and 180 respectively. \add{The
subset of the HET/Uganda data was chosen by eye such that the individuals were
monopolistic in both the \textit{gag} and \textit{env} trees. The other subsets
were arbitrarily chosen subtrees from phylogenies of the complete datasets.}

For all datasets, we used the priors $\alpha$ $\sim$ Uniform(0, 2) and $N$ and
$I$ jointly uniform on the region \{$n \leq N \leq 10000$, $n \leq I \leq
10000$, $I \leq N$\}, where $n$ is the number of tips in the tree. Since the
value $m = 1$ produces networks with no cycles, which we considered fairly
implausible, we ran one analysis with the prior $m \sim$ DiscreteUniform(1, 5),
and one with the prior $m \sim$ DiscreteUniform(2, 5). The other parameters to
the SMC algorithm were the same as used for the simulation experiments,
\add{except that we used 10000 particles instead of 1000 to increase the
accuracy of the estimated posterior. This was computationally feasible due to
the small number of runs required for this analysis.}

\section*{Results}

\subsection*{\del{Kernel classifiers}\add{Classifiers for BA model parameters from tree shapes}}

<<classifiers, include=FALSE>>=
    get.kernel.data <- function (param, step) {
        md <- collect.metadata(paste0('../../simulations/kernel-', param, '/', step, '/*'))
        if ('rbf_variance' %in% colnames(md)) {
          md <- subset(md, rbf_variance == 4 & decay_factor == 0.3)
        }
        if (param == 'alpha') {
            md <- subset(md, m == 2)
        }
        d <- setDT(collect.data(rownames(md)))
        if (param %in% c('alpha', 'm')) {
            d[,list(rsquared=mean(rsquared)), by=c('nsimnode', 'ntip')]
        } else if (param == 'I') {
            d[,list(rsquared=mean(rsquared)), by=ntip]
        } else {
            d[,list(rsquared=mean(rsquared)), by=c('I', 'ntip')]
        }
    }
    kernel.alpha <- get.kernel.data('alpha' ,'classifier')
    kernel.m <- get.kernel.data('m', 'classifier')
    kernel.I <- get.kernel.data('I', 'classifier')
    kernel.N <- get.kernel.data('N', 'classifier')

    sackin.alpha <- get.kernel.data('alpha', 'stats-classifier')
    sackin.m <- get.kernel.data('m', 'stats-classifier')
    sackin.I <- get.kernel.data('I', 'stats-classifier')
    sackin.N <- get.kernel.data('N', 'stats-classifier')

    nltt.alpha <- get.kernel.data('alpha', 'nltt-classifier')
    nltt.m <- get.kernel.data('m', 'nltt-classifier')
    nltt.I <- get.kernel.data('I', 'nltt-classifier')
    nltt.N <- get.kernel.data('N', 'nltt-classifier')
@

We investigated the \add{identifiability of four} parameters of the BA network
model~\autocite{barabasi1999emergence}: the number of nodes $N$, the
preferential attachment power $\alpha$, the number of edges added per vertex
$m$, and the number of infected nodes $I$. \del{In addition to $m$ and $\alpha$
(see Introduction), we considered $N$, which denotes the total number of nodes
in the network, and $I$, which is the number of infected nodes at which to stop
the simulation and sample the transmission tree.} To examine the effect of
these parameters on tree shape, we simulated transmission trees under different
parameter values, calculated pairwise tree kernel scores between them, and
attempted to classify the trees using a kernel support vector machine (kSVR).
We also tested classifiers based on Sackin's index~\autocite{shao1990tree} and
the normalized lineages-through-time (nLTT)
statistic~\autocite{janzen2015approximate}. The accuracy of each classifier
\add{on all BA parameters is shown in} \del{varied based on the parameter
being tested} \cref{fig:rsquared}. Classifiers based on \del{two other tree
statistics,} the nLTT and Sackin's index generally exhibited worse performance
than the tree kernel, although the magnitude of the disparity varied between
the parameters (\cref{fig:rsquared}, centre and right). The results were
largely robust to variations in the tree kernel meta-parameters $\lambda$ and
$\sigma$, \add{although accuracy varied between different epidemic and sampling
scenarios} (\cref{fig:alphacrossv,fig:mcrossv,fig:Ncrossv,fig:Icrossv}).
\add{For all parameters except $m$, the absolute number of tips in the tree had
a much greater impact on accuracy than the proportion of infected individuals
these tips represented. However, for $m$, both the number and proportion of
sampled tips had a strong impact or the accuracy of the kSVR
(\cref{fig:mcrossv}).}

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{kernel-rsquared.pdf}
    \caption[
        Cross-validation accuracy of kernel-SVR, nLTT-based SVR, and Sackin's
        index regression classifiers for BA model parameters.
    ]{
        Cross-validation accuracy of kernel-SVR classifier (left), SVR
        classifier using nLTT (centre), and linear regression using
        Sackin's index (right) for BA model parameters. Kernel meta-parameters
        were set to $\lambda = 0.3$ and $\sigma = 4$. Each point was calculated
        based on 300 simulated transmission trees over networks with three
        different values of the parameter being tested, assuming perfect
        knowledge of the other parameters. Vertical lines are empirical 95\%
        confidence intervals based on 1000 two-fold cross-validations. \add{The
        classifiers for $I$ were not evaluated with 1000-tip trees, because one
        of the tested $I$ values was 500, and it is not possible to sample a
        tree of size 1000 from 500 infected individuals.}
    }
    \label{fig:rsquared}
\end{figure}

\add{The accuracy of each classifier on all BA parameters is shown in
\cref{fig:rsquared}. Classifiers based on the nLTT and Sackin's index generally
exhibited worse performance than the tree kernel, although the magnitude of the
disparity varied between the parameters (\cref{fig:rsquared}, centre and
right). The results were largely robust to variations in the tree kernel
meta-parameters $\lambda$ and $\sigma$, although accuracy varied between
different epidemic and sampling scenarios
(\cref{fig:alphacrossv,fig:mcrossv,fig:Icrossv,fig:Ncrossv}). For all
parameters except $m$, the absolute number of tips in the tree had a much
greater impact on accuracy than the proportion of infected individuals these
tips represented. However, for $m$, both the number and proportion of sampled
tips had a strong impact or the accuracy of the kSVR (\cref{fig:mcrossv}).}

\end{document}

The \gls{kSVR} classifier for \gls{alpha} had an average $R^2$ of 
    \Sexpr{kernel.alpha[,round(mean(rsquared), 2)]},
compared to 
    %\S%expr{nltt.alpha[,round(mean(rsquared), 2)]}
for the nLTT-based SVR, and
    %\S%expr{sackin.alpha[,round(mean(rsquared), 2)]}
for the linear regression against Sackin's index. There was little variation
about the mean for different tree and epidemic sizes. No classifier could
accurately identify the $m$ parameter in any epidemic scenario, with average
$R^2$ values of 
  %\S%expr{kernel.m[,round(mean(rsquared), 2)]} for kSVR,
  %\S%expr{nltt.m[,round(mean(rsquared), 2)]} for the nLTT, and
  %\S%expr{sackin.m[,round(mean(rsquared), 2)]}
for Sackin's index. Again, there was little variation in accuracy between
epidemic scenarios, although the accuracy of the kSVR was slightly higher
on 1000-tip trees (\cref{fig:rsquared}, left).

The accuracy of classifiers for $I$ varied significantly with the number of
tips in the tree. For 100-tip trees, the average $R^2$ values were
  %\S%expr{kernel.I[ntip == 100, round(mean(rsquared), 2)]},
  %\S%expr{nltt.I[ntip == 100, round(mean(rsquared), 2)]}, and
  %\S%expr{sackin.I[ntip == 100, round(mean(rsquared), 2)]}
for the tree kernel, nLTT, and Sackin's index respectively. For 500-tip
trees, the values increased to
  %\S%expr{kernel.I[ntip == 500, round(mean(rsquared), 2)]},
  %\S%expr{nltt.I[ntip == 500, round(mean(rsquared), 2)]}, and
  %\S%expr{sackin.I[ntip == 500, round(mean(rsquared), 2)]}.
Finally, the performance of classifiers for $N$ depended heavily on the
epidemic scenario. The $R^2$ of the kSVR classifier ranged from
  %\S%expr{kernel.N[,round(min(rsquared), 2)]}
for the smallest epidemic and smallest sample size, to
  %\S%expr{kernel.N[,round(max(rsquared), 2)]}
for the largest. Likewise, $R^2$ for the nLTT-based SVR ranged from 
  %\S%expr{nltt.N[,round(min(rsquared), 2)]}
to
  %\S%expr{nltt.N[,round(max(rsquared), 2)]}.
Sackin's index did not accurately classify $N$ in any scenario, with an average
$R^2$ of
  %\S%expr{sackin.N[,round(mean(rsquared), 2)]}
and little variation between scenarios.

\subsection*{ABC simulations}

<<point_est, include=FALSE, eval=FALSE>>=
    f <- Sys.glob("../../simulations/abc-pa-free-m/point-estimate/*")
    d <- fread(f)
    d[,m := floor(m)]
    d[,alpha_error := abs(true_alpha - alpha)]
    d[,N_error := abs(true_N - N)]
    d[,I_error := abs(true_I - I)]
    d[,m_error := abs(true_m - m)]

    alpha.av <- anova(lm(alpha_error ~ factor(true_alpha) + factor(true_m) + factor(true_I), d))
    m.av <- anova(lm(m_error ~ factor(true_alpha) + factor(true_m) + factor(true_I), d))
    N.av <- anova(lm(N_error ~ factor(true_alpha) + factor(true_m) + factor(true_I), d))
    I.av <- anova(lm(I_error ~ factor(true_alpha) + factor(true_m) + factor(true_I), d))

    m.tbl <- d[,prop.table(table(m_error))]
    zero.alpha.test <- wilcox.test(d[true_alpha == 0, alpha_error], 
                                   d[true_alpha != 0, alpha_error])
    alpha.I.test <- wilcox.test(d[true_alpha < 1, I_error],
                                d[true_alpha >= 1, I_error])
    I.I.test <- wilcox.test(d[true_I == 1000, I_error],
                            d[true_I == 2000, I_error])
    alpha.m.test <- wilcox.test(d[true_alpha == 0 | true_alpha == 1, m_error],
                                d[true_alpha == 0.5 | true_alpha == 1.5, m_error])
    stopifnot(alpha.av["factor(true_m)", "Pr(>F)"] <= 0.05)
    stopifnot(alpha.av["factor(true_I)", "Pr(>F)"] > 0.05)
    stopifnot(I.I.test$p.value < 1e-5)
    stopifnot(I.av["factor(true_m)", "Pr(>F)"] > 0.05)
    stopifnot(m.av["factor(true_m)", "Pr(>F)"] > 0.05)
    stopifnot(m.av["factor(true_I)", "Pr(>F)"] > 0.05)
    stopifnot(m.av["factor(true_alpha)", "Pr(>F)"] > 0.05)
    stopifnot(min(N.av[1:3,"Pr(>F)"]) > 0.05)

    alpha.m.cor <- cor.test(d[,alpha_error], d[,true_m], method="spearman")
@

\Cref{fig:abcpt} shows maximum \textit{a posteriori} (MAP) point
estimates of the BA model parameters obtained with ABC on simulated
data. The estimates shown correspond only to the simulations where the $m$
parameter was set to 2, however the results for $m = 3$ and $m = 4$ were
similar (\cref{fig:abcptm3,fig:abcptm4}). Average boundaries of 95\% HPD
intervals are given in \cref{tab:abchpd}.

The accuracy of the parameter estimates obtained with ABC
paralleled the results from the kSVR classifier. Of the four parameters,
$\alpha$ was the most accurately estimated, with point estimates having a
median [IQR] absolute error of 
    %\S%expr{d[,round(median(alpha_error), 2)]} 
    [%\S%expr{d[,round(quantile(alpha_error, 0.25), 2)]} - 
    %\S%expr{d[,round(quantile(alpha_error, 0.75), 2)]}].
The errors when the true value of $\alpha$ was zero were significantly greater
than those for the other values 
    (Wilcoxon rank-sum test, $p$ = %\S%expr{latexSN(round(zero.alpha.test$p.value, 4))}$).
Errors in estimating $\alpha$ also varied with the true value of $m$ just at
the threshold of statistical significance
    ($p$ %\S%expr{pp(alpha.av["factor(true_m)", "Pr(>F)"])}$),
but did not vary across the true values of $N$ or $I$ (both one-way ANOVA).
Estimates for $I$ were relatively accurate, with point estimate errors of
    %\S%expr{d[,round(median(I_error))]} 
    [%\S%expr{d[,round(quantile(I_error, 0.25))]} - 
    %\S%expr{d[,round(quantile(I_error, 0.75))]}] individuals.
These errors were significantly higher when the true value of $\alpha$ was
at least 1
    (Wilcoxon rank-sum test, $p$ = %\S%expr{latexSN(round(alpha.I.test$p.value, 4))}$)
and when the true value of $I$ was 2000 ($p < 10^{-5}$). The true value of $m$
did not affect the estimates of $I$ (one-way ANOVA).

The $m$ parameter was estimated correctly in only
    %\S%expr{as.integer(m.tbl[1] * 100)} \%
of simulations, barely better than random guessing. The true values of the
other parameters did not significantly affect the estimates of $m$ (both
one-way ANOVA). Finally, the total number of nodes $N$ was consistently
over-estimated by about a factor of two
    (error %\S%expr{d[,round(median(N_error))]} 
    [%\S%expr{d[,round(quantile(N_error, 0.25))]} - 
     %\S%expr{d[,round(quantile(N_error, 0.75))]}] individuals).
No parameters influenced the accuracy of the $N$ estimates (all one-way ANOVA).

\begin{figure}[ht]
  \centering
  \includegraphics[width=\textwidth]{abc-point-estimate-m2}
  \vspace{6pt}
  \caption{
    Point estimates of BA model parameters obtained by running ABC
    on simulated phylogenies without training, for simulations with $m = 2$.
    Dotted lines indicate true values, and limits of the $y$-axes are regions
    of uniform prior density. (A) Estimates for $\alpha$ and $I$ against their
    true values in simulations. (B) Estimates for $m$ and $N$, which were held
    fixed in these simulations, against true values of $\alpha$.
  }
  \label{fig:abcpt}
\end{figure}

\begin{table*}[ht]
  \centering
  \input{\tablepath/abc-hpd.tex}
  \caption{
      Average maximum \textit{a posteriori} point estimates and 95\% highest
      posterior density (HPD) interval widths for BA model parameter estimates
      obtained with ABC. Three transmission trees were simulated under
      each combination of the listed parameter values, and the parameters were
      estimated with ABC without training.
  }
  \label{tab:abchpd}
\end{table*}

<<supp_labels, echo=FALSE, results="asis", eval=FALSE>>=
    N <- 5000
    replicate <- 0
    lab <- NULL
    for (m in c(2, 3, 4)) {
    for (I in c(1000, 2000)) {
    for (alpha in c(0, 0.5, 1, 1.5)) {
        lab <- c(lab, sprintf("fig:%.1f-%d-%d-%d-%d", alpha, I, m, N, replicate))
    }
    }
    }
    lab <- paste(lab, collapse = ",")
@

The dispersion of the ABC approximation to the posterior also varied between
the parameters, with narrower HPD intervals for the parameters with more
accurate point estimates (\cref{tab:abchpd}). \Cref{fig:abcex} shows
the distributions for for one simulation. Equivalent plots for one replicate
simulation with each studied parameter combination can be found in
\cref{%\S%expr{lab}}. HPD intervals around $\alpha$ and $I$ were often narrow
relative to the region of nonzero prior density, whereas the intervals for $m$
and $N$ were more widely dispersed.

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{abc-posterior-example}
  \vspace{6pt}
  \caption{
    Marginal posterior distributions of BA model parameters estimated
    with ABC for a single simulated transmission tree. Dotted
    lines and shaded polygon indicate true values.
  }
  \label{fig:abcex}
\end{figure}

<<mixed_peerdriven, include=FALSE, eval=FALSE>>=
  library(coda)
  d <- fread("bzcat ../../simulations/abc-pa-mixed-alpha/abc/*")
  d <- d[iter == max(iter)]
  d <- d[sample(1:nrow(d), prob=weight, replace=TRUE)]
  params <- c("alpha", "m", "N", "I")
  d <- melt(d, measure.vars=params, variable.name="parameter")
  f <- function (x) {
    dens <- density(x)
    hpd <- HPDinterval(mcmc(x))
    list(map=dens$x[which.max(dens$y)], hpd.min=hpd[,"lower"], 
         hpd.max=hpd[,"upper"])
  }
  d <- d[,f(value), by=parameter]
  setkey(d, parameter)

  m <- collect.metadata("../../simulations/abc-pa-peerdriven/abc/*")
  p <- fread(paste("bzcat", rownames(m)[m$sample_peer == 1]))
  p <- p[iter == max(iter)]
  p <- p[sample(1:nrow(p), prob=weight, replace=TRUE)]
  params <- c("alpha", "m", "N", "I")
  p <- melt(p, measure.vars=params, variable.name="parameter")
  f <- function (x) {
    dens <- density(x)
    hpd <- HPDinterval(mcmc(x))
    list(map=dens$x[which.max(dens$y)], hpd.min=hpd[,"lower"], 
         hpd.max=hpd[,"upper"])
  }
  p <- p[,f(value), by=parameter]
  setkey(p, parameter)
  options(scipen=5)
@

To test the effect of model misspecification, we simulated one network where
the nodes exhibited heterogeneous preferential attachment power (half 0.5, the
other half 1.5), with $m$ = 2, $N$ = 5000, and $I$ = 1000. The MAP [95\%
HPD] estimates for each parameter were: 
$\alpha$, 
  %\S%expr{round(d["alpha", map], 2)} 
  [%\S%expr{round(d["alpha", hpd.min], 2)} -
   %\S%expr{round(d["alpha", hpd.max], 2)}];
$I$,
  %\S%expr{d["I", round(map)]} 
  [%\S%expr{d["I", round(hpd.min)]} -
   %\S%expr{d["I", round(hpd.max)]}];
$m$,
  %\S%expr{d["m", floor(map)]} 
  [%\S%expr{d["m", floor(hpd.min)]} -
   %\S%expr{d["m", floor(hpd.max)]}];
$N$,
  %\S%expr{d["N", round(map)]} 
  [%\S%expr{d["N", round(hpd.min)]}-
   %\S%expr{d["N", round(hpd.max)]}].
The approximate posterior distributions for this simulation are shown in
\cref{fig:mixed}. To test the effect of sampling bias, we sampled one
transmission tree in a peer-driven fashion, where the probability to sample a
node was twice as high if one of its peers had already been sampled. The
parameters for this experiment were $N$ = 5000, $m$ = 2, $\alpha$ = 0.5, and
$I$ = 2000. The estimated values were
$\alpha$, 
  %\S%expr{p["alpha", round(map, 2)]} 
  [%\S%expr{p["alpha", round(hpd.min, 2)]} -
   %\S%expr{p["alpha", round(hpd.max, 2)]}];
$I$,
  %\S%expr{p["I", floor(map)]} 
  [%\S%expr{p["I", floor(hpd.min)]} -
   %\S%expr{p["I", floor(hpd.max)]}];
$m$,
  %\S%expr{p["m", floor(map)]} 
  [%\S%expr{p["m", floor(hpd.min)]} -
   %\S%expr{p["m", floor(hpd.max)]}];
$N$,
  %\S%expr{p["N", floor(map)]} 
  [%\S%expr{p["N", floor(hpd.min)]} -
   %\S%expr{p["N", floor(hpd.max)]}].
The approximate posterior distributions are shown in \cref{fig:peerdriven}. Both
of these results were in line with estimates obtained on other simulated
datasets (\cref{tab:abchpd}), although the estimate of peer-driven sampling for
$\alpha$ was somewhat lower than typical.

\subsection*{Real data}

<<realdata, include=FALSE, eval=FALSE>>=
  m <- collect.metadata("../../simulations/aggregates/hpd/*")
  d1 <- fread(rownames(m)[m$m_min == 1])
  d2 <- fread(rownames(m)[m$m_min == 2])

  map1 <- dcast.data.table(d1, dataset~parameter, value.var="map")
  up1 <- dcast.data.table(d1, dataset~parameter, value.var="hpd.max")
  low1 <- dcast.data.table(d1, dataset~parameter, value.var="hpd.min")
  setkey(map1, dataset)
  setkey(up1, dataset)
  setkey(low1, dataset)

  map2 <- dcast.data.table(d2, dataset~parameter, value.var="map")
  up2 <- dcast.data.table(d2, dataset~parameter, value.var="hpd.max")
  low2 <- dcast.data.table(d2, dataset~parameter, value.var="hpd.min")
  setkey(map2, dataset)
  setkey(up2, dataset)
  setkey(low2, dataset)
@

We applied ABC to five published HIV datasets (\cref{tab:data}),
and found substantial heterogeneity among the parameter estimates
(\cref{fig:abchpd,fig:abchpdm2}). Plots of the marginal posterior distributions
for each dataset are shown in
\cref{fig:cuevas,fig:li,fig:niculescu,fig:novitsky,fig:wang}.
Two of the datasets~\autocite{niculescu2015recent, wang2015targeting} had
estimated $\alpha$ values near unity for the prior allowing $m = 1$ (MAP
estimates [95\% HPD] 
  %\S%expr{map1["niculescu2015", round(alpha, 2)]} 
  [%\S%expr{low1["niculescu2015", round(alpha, 2)]} - 
   %\S%expr{up1["niculescu2015", round(alpha, 2)]}]
and
  %\S%expr{map1["wang2015", round(alpha, 2)]} 
  [%\S%expr{low1["wang2015", round(alpha, 2)]} -
   %\S%expr{up1["wang2015", round(alpha, 2)]}] respectively).
The MAP estimates did not change appreciably when $m = 1$ was disallowed by the
prior, although the credible interval of the \textcite{niculescu2015recent}
data was narrower
  (%\S%expr{low1["niculescu2015", round(alpha, 2)]} - 
   %\S%expr{up1["niculescu2015", round(alpha, 2)]}).
When $m = 1$ was permitted, the \textcite{li2015hiv, cuevas2009hiv} both had
low estimated $\alpha$ values
  (%\S%expr{map1["li2015", round(alpha, 2)]} 
  [%\S%expr{low1["li2015", round(alpha, 2)]} - 
  %\S%expr{up1["li2015", round(alpha, 2)]}]
and
  %\S%expr{map1["cuevas2009", round(alpha, 2)]} 
  [%\S%expr{low1["cuevas2009", round(alpha, 2)]} -
   %\S%expr{up1["cuevas2009", round(alpha, 2)]}]). 
However, the MAP estimates increased when $m = 1$ was not permitted, although
the HPD intervals remained roughly the same
  (%\S%expr{map2["li2015", round(alpha, 2)]} 
  [%\S%expr{low2["li2015", round(alpha, 2)]} - 
  %\S%expr{up2["li2015", round(alpha, 2)]}]
and
  %\S%expr{map2["cuevas2009", round(alpha, 2)]} 
  [%\S%expr{low2["cuevas2009", round(alpha, 2)]} -
   %\S%expr{up2["cuevas2009", round(alpha, 2)]}]).
The \textcite{novitsky2014impact} data had a fairly low estimated $\alpha$
for both priors on $m$
  (%\S%expr{map1["novitsky2014", round(alpha, 2)]} for $m \geq 1$;
   %\S%expr{map2["novitsky2014", round(alpha, 2)]} for $m \geq 2$).
However, the confidence interval was much wider when $m = 1$ was allowed
  ([%\S%expr{low1["novitsky2014", round(alpha, 2)]} -
    %\S%expr{up1["novitsky2014", round(alpha, 2)]}] for $m \geq 1$ vs.
    %\S%expr{low2["novitsky2014", round(alpha, 2)]} -
    %\S%expr{up2["novitsky2014", round(alpha, 2)]} for $m \geq 2$).

For all the datasets except \citeauthor{novitsky2014impact}, estimated values
of $I$ were below 2000 when $m = 1$ was allowed, with relatively narrow HPD
intervals compared to the nonzero prior density region
  (\citeauthor{cuevas2009hiv}, %\S%expr{map1["cuevas2009", floor(I)]} 
  [%\S%expr{low1["cuevas2009", floor(I)]} -
   %\S%expr{up1["cuevas2009", floor(I)]}];
   \citeauthor{niculescu2015recent}, %\S%expr{map1["niculescu2015", floor(I)]}
  [%\S%expr{low1["niculescu2015", floor(I)]} - 
   %\S%expr{up1["niculescu2015", floor(I)]}];
  \citeauthor{li2015hiv}, %\S%expr{map1["li2015", floor(I)]} 
  [%\S%expr{low1["li2015", floor(I)]} -
   %\S%expr{up1["li2015", floor(I)]}];
   \citeauthor{wang2015targeting}, %\S%expr{map1["wang2015", floor(I)]}
  [%\S%expr{low1["wang2015", floor(I)]} - 
   %\S%expr{up1["wang2015", floor(I)]}]).
The \citeauthor{novitsky2014impact} data was the outlier, with a very high
estimated $I$, and HPD interval spanning almost the entire prior region
  (%\S%expr{map1["novitsky2014", floor(I)]} 
  [%\S%expr{low1["novitsky2014", floor(I)]} -
   %\S%expr{up1["novitsky2014", floor(I)]}]).
The $I$ estimates and HPD intervals were generally robust to the choice of
prior on $m$, with slightly narrower HPD intervals (compare
\cref{fig:abchpd,fig:abchpdm2}).

The MAP estimate of $m$ was equal to 1 for all but the
\citeauthor{novitsky2014impact} data, when this value was allowed. However, the
upper bound of the HPD interval was different for each dataset
  (\citeauthor{niculescu2015recent}, %\S%expr{up1["niculescu2015", round(m)]};
   \citeauthor{wang2015targeting}, %\S%expr{up1["wang2015", round(m)]};
   \citeauthor{li2015hiv}, %\S%expr{up1["li2015", round(m)]};
   \citeauthor{cuevas2009hiv}, %\S%expr{up1["cuevas2009", round(m)]}).
When $m = 1$ was disallowed, the MAP for all datasets was either 2 or 3, with
HPD intervals spanning the entire prior region. The estimates for the total
number of nodes $N$ were largely uninformative for all samples, with almost all
MAP estimates greater than 7500 and HPD intervals spanning almost the entire
nonzero prior density region. The only exception was the \citeauthor{li2015hiv}
data, for which the MAP estimate was lower 
  (%\S%expr{map1["li2015", floor(N)]})
when $m = 1$ was allowed.

\begin{figure}[ht]
  \centering
  \includegraphics{realdata-hpd}
  \vspace{8pt}
  \caption{
      Maximum \textit{a posteriori} point estimates and 95\% HPD intervals for
      parameters of the BA network model, fitted to five published HIV datasets
      with ABC. $x$-axes indicate regions of nonzero prior density. In
      particular, the prior on $m$ was DiscreteUniform(1, 5).
  }
  \label{fig:abchpd}
\end{figure}

\clearpage

\section*{Discussion}

Contact networks can have a strong influence on epidemic progression, and are
potentially useful as a public health tool~\autocite{wang2015targeting,
little2014using}. Despite this, few methods exist for investigating contact
network parameters in a phylodynamic framework \autocite[although see][for
related work]{groendyke2011bayesian, volz2008sir, brown2011transmission,
leventhal2012inferring}. Kernel-ABC is a model-agnostic method which can be
used to investigate any quantity that affects tree
shape~\autocite{poon2015phylodynamic}. In this work, we developed a
ABC-based method to infer the parameters of a contact network model. The
method is general, and could be applied to any model from which contact
networks can be simulated. We demonstrated the method on the BA model, which is
a simple preferential attachment model giving rise to the power law degree
distributions commonly observed in real-world networks.

By training a kernel-SVR classifier, we found that the $\alpha$ and $I$
parameters, representing preferential attachment power and number of infected
nodes, had a strong influence on tree shape. This was reflected in the relative
accuracy of the ABC estimates of these parameters. The total number of
nodes $N$ had a weak influence on tree shape, which was most prominent when the
epidemic size $I$ and number of sampled tips were both large. The $m$
parameter, representing the number of edges created in the network per vertex,
did not produce much variation in tree shape, resulting in in both poorly
performing classifiers and uninformative ABC estimates.

$N$ was almost always significantly over-estimated using ABC. Since the
prior on $N$ and $I$ is jointly uniform on a non-rectangular region ($I \leq
N$), there is more prior mass on high $N$ values. In retrospect, it is
unreasonable to expect good estimation of $N$, because adding more nodes to a
BA network does not change the edge density or overall shape. This can be
illustrated by imagining that we add a small number of nodes to a network after
the epidemic simulation has already been completed. It is possible that none of
these new nodes attains a connection to any infected node. Thus, running the
simulation again on the new, larger network could produce the exact same
transmission tree as before. We note also that our accurate estimates of $I$
may have been influenced by this prior, which places more mass on low $I$
values. However, the MAP estimate of $I$ was very high for the
\textcite{novitsky2013phylogenetic,novitsky2014impact} data, suggesting that a
strong enough signal in the data can overcome the prior.

As noted by \textcite{lintusaari2016identifiability}, uniform priors on model
parameters may translate to highly informative priors on quantities of
interest. We observed a non-linear relationship between the preferential
attachment power $\alpha$ and the power law exponent $\gamma$
(\cref{fig:gamma}). Therefore, placing a uniform prior on $\alpha$ between 0
and 2 is equivalent to placing an informative prior that $\gamma$ is close to
2. Therefore, if we were primarily interested in $\gamma$ rather than
$\alpha$, a more sensible choice of prior might have a shape informed by
\cref{fig:gamma} and be bounded above by approximately $\alpha$ = 1.5. This
would uniformly bound $\gamma$ in the region $2 \leq \gamma \leq 4$ commonly
reported in the network literature~\autocite{liljeros2001web,
schneeberger2004scale, colgate1989risk, brown2011transmission}. We note however
that \textcite{jones2003assessment} estimated $\gamma$ values greater than
four for some datasets, in one case as high as 17, indicating that a wider
range of permitted $\gamma$ values may be warranted.

Our investigation of published HIV datasets indicated heterogeneity in the
contact network structures underlying several distinct local epidemics. When
interpreting these results, we caution that the BA model is quite simple and
most likely misspecified for these data. In particular, the average degree of a
node in the network is equal to $2m$, and therefore is constrained to be a
multiple of 2. Furthermore, we considered the case $m = 1$, where the network
has no cycles, to be implausible and therefore assigned it zero prior
probability in one set of analyses. This forced the average degree to be at
least four, which may be unrealistically high for sexual networks. The fact
that the estimated values of $\alpha$ differed substantially for three datasets
depending on whether or not $m = 1$ was allowed by the prior is further evidence
of this potential misspecification. However, we note that for two of the
datasets, the estimated values of $\alpha$ did not change much between priors,
and the estimates of $I$ were robust to the choice of prior for all datasets
studied. More sophisticated models, for example models incorporating
heterogeneity in node behaviour, are likely to provide a better fit to these
data.

With respect to the preferential attachment power $\alpha$, the five datasets
analysed fell into three categories (\cref{fig:abchpd}). First, we
estimated a preferential attachment power close to 1, indicating linear
preferential attachment, for the outbreaks studied by
\textcite{niculescu2015recent} and \textcite{wang2015targeting}. These values
were robust to specifying different priors for $m$. Both studies were of
populations in which we would expect a high degree of epidemiological
relatedness: \textcite{niculescu2015recent} studied a recent outbreak among
Romanian injection drug users (IDU), while \citeauthor{wang2015targeting}
sampled acutely infected MSM in Beijing, China. Both these are contexts in
which we would expect some of the assumptions of the BA model, such as a
connected network, relatively high mean degree, and preferential attachment
dynamics, to hold.

The remaining three datasets (\textcite{cuevas2009hiv, novitsky2014impact,
li2015hiv}) had estimated values of $\alpha$ below 0.5 when $m = 1$ was
included in the prior, but these were not robust to changing the prior to
exclude $m = 1$. For the \citeauthor{cuevas2009hiv} data, model
misspecification is likely partially responsible. While the authors found that
a large proportion of the samples were epidemiologically linked, these were
mainly in small local clusters rather than the single large component
postulated by the BA model. In addition, the mixed risk groups in the dataset
would be unlikely to significantly interact, further weakening any global
preferential attachment dynamics. The dataset studied by
\textcite{novitsky2014impact} originated from a densely sampled population
where the predominant risk factor was believed to be heterosexual exposure.
Although the MAP estimate of $\alpha$ was almost unchanged when the value $m =
1$ was excluded from the prior, the confidence interval shrank significantly.
For both priors, the estimated prevalence was extremely high, in fact higher
than the estimated HIV prevalence in the sampled region. The authors indicated
that the source of the samples was a town in close proximity to the country's
capital city, and suggested that there may have been a high degree of migration
and partner interchange between the two locations. It is possible that the
contact network underlying the subtree we investigated includes a much larger
group based in the capital city, which would explain the high estimate of $I$.
There is no clear explanation for the discrepancy between the two priors for
the \textcite{li2015hiv} data, as the subset we analyzed formed a phylogenetic
cluster and therefore was a good candidate for the BA model. However, nearly
all the posterior density was assigned to $m = 1$ when this value was allowed,
indicating that the network was more likely to have an acyclic tree structure.

In addition to the aforementioned possibility of misspecification, additional
modelling assumptions include the network being connected and static, all
transmission rates being equal, no removal after infection, identical behaviour
of all nodes, and random sampling. The last two were addressed with small-scale
experiments. We simulated a network where some nodes exhibited a higher
attachment power than others, and found that the estimated attachment power was
simply the average of the two values. This indicated that, although we could
characterize the network in aggregate, the estimated parameters could not be
said to apply to any individual node. The effect of biased sampling was
investigated by analysing a transmission tree which had been sampled in a
peer-driven fashion. The results were roughly in line with those for random
sampling, however the estimated value of $\alpha$ was lower than the average
for randomly-sampled trees. Further experiments would be necessary to fully
explore the impact of these assumptions on the method's accuracy. However,
despite these issues, we felt it was best to demonstrate the method first on a
simple model. It is possible to use this framework to fit more complex models
which address some of these issues, such as one incorporating heterogeneous
node behaviour, which may prove a fruitful avenue
for future investigations.

Our method has a number of caveats, perhaps the most significant being that it
takes a transmission tree as input. In reality, true transmission trees are not
available and must be approximated, often by way of a viral phylogeny. Although
this has been demonstrated to be a fair
approximation~\autocite[e.g.][]{leitner1996accurate}, and is frequently used in
practice~\autocite[e.g.][]{stadler2013uncovering}, the topologies of a viral
phylogeny and transmission tree can differ
significantly~\autocite{ypma2013relating} due to within-host evolution and the
sampling process. In addition, the ABC-SMC algorithm is
computationally intensive, taking about a day when run on 20 cores in parallel
with the settings we described in the methods. Nevertheless, our method is
potentially useful to epidemiological researchers interested in the general
characteristics of the network structure underlying disease outbreaks. This
work, and previous work by our group~\autocite{poon2015phylodynamic}, has
demonstrated that ABC is a broadly applicable and effective framework in
which to perform phylodynamic inference.

\section*{Acknowledgements}

This work was supported by grants from the Canadian Institutes of Health
Research (CIHR, operating grant HOP-111406), and the Bill \& Melinda Gates
Foundation (award number OPP1110049). R.M.M. was supported by a scholarship
from the CIHR Strategic Training Program in Bioinformatics. A.F.Y.P. was
supported by a CIHR New Investigator Award (Canadian HIV Vaccine Initiative,
Vaccine Discovery and Social Research) and by a Career Investigator Scholar
Award from the Michael Smith Foundation for Health Research, in partnership
with the Providence Health Care Research Institute and St. Paul's Hospital
Foundation.

\printbibliography

\clearpage
\newpage

\section*{Supplementary Figures}

\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}
\setcounter{page}{1}

\begin{figure}[ht]
  \centering
  \includegraphics{leventhal2012fig1.pdf}
  \caption{
    Reproduction of Figure 1A from Leventhal \textit{et al.} (2012) used to
    check the accuracy of our implementation of Gillespie simulation.
    Transmission trees were simulated over three types of network, with
    pathogen transmissibility varying from 0 to 1. Sackin's index was
    calculated for each simulated transmission tree.
  }
  \label{fig:leventhal}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics{smc-test.pdf}
  \caption{Approximation of mixture of Gaussians used by
    Del Moral \textit{et al.} (2012) and Sisson \textit{et al.} (2009) to test
    SMC. Solid black line indicates true distribution. Grey shaded area shows
    SMC approximation obtained with our implementation, using 10000
    particles with one simulated data point per particle.
  }
  \label{fig:smctest}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics{kernel-alpha-crossv.pdf}
  \caption{
    Cross-validation accuracy of kernel-SVM classifiers for $\alpha$ parameter
    of BA network model, for various tree kernel meta-parameters and
    epidemic scenarios. Each point was calculated based on 300 simulated
    transmission trees over networks with $\alpha$ = 0.5, 1.0, or 1.5. Dotted
    and and dashed lines indicate, respectively, performance of SVM using the
    nLTT statistic, and linear regression using Sackin's index. Facets
    are number of infected nodes before the simulation was stopped ($I$) and
    number of tips in the sampled transmission tree.
  }
  \label{fig:alphacrossv}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics{kernel-m-crossv.pdf}
  \caption{Cross-validation accuracy of kernel-SVM classifiers for $m$
      parameter of BA network model, for various tree kernel
      meta-parameters and epidemic scenarios. Each point was calculated based
      on 300 simulated transmission trees over networks with $m$ = 2, 3, or 4.
      Dotted and and dashed lines indicate, respectively, performance of SVM
      using the nLTT statistic, and linear regression using Sackin's
      index. Facets are number of infected nodes before the simulation was
      stopped ($I$) and number of tips in the sampled transmission tree.
  }
  \label{fig:mcrossv}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics{kernel-I-crossv.pdf}
  \caption{Cross-validation accuracy of kernel-SVM classifiers for number of
      infected nodes ($I$) under BA network model, for various tree
      kernel meta-parameters and two tree sizes. Each point was calculated
      based on 300 simulated transmission trees over networks with $I$ = 500,
      1000, or 2000. Dotted and and dashed lines indicate, respectively,
      performance of SVM using the nLTT statistic, and linear regression
      using Sackin's index. Facets are the number of tips in the sampled
      transmission tree.
  }
  \label{fig:Icrossv}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics{kernel-N-crossv.pdf}
  \caption{Cross-validation accuracy of kernel-SVM classifiers for total number
      of nodes ($N$) under BA network model, for various tree kernel
      meta-parameters and epidemic scenarios sizes. Each point was calculated
      based on 300 simulated transmission trees over networks with $N$ = 3000,
      5000, or 8000. Dotted and and dashed lines indicate, respectively,
      performance of SVM using the nLTT statistic, and linear regression
      using Sackin's index. Facets are the number of tips in the sampled
      transmission tree.
  }
  \label{fig:Ncrossv}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics{abc-point-estimate-m3.pdf}
  \caption{
    Point estimates of BA model parameters obtained by running ABC
    on simulated phylogenies without training, for simulations with $m = 3$.
    Dotted lines indicate true values, and limits of the $y$-axes are regions
    of uniform prior density. (A) Estimates for $\alpha$ and $I$ against their
    true values in simulations. (B) Estimates for $m$ and $N$, which were held
    fixed in these simulations, against true values of $\alpha$.
  }
  \label{fig:abcptm3}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics{abc-point-estimate-m4.pdf}
  \caption{
    Point estimates of BA model parameters obtained by running ABC
    on simulated phylogenies without training, for simulations with $m = 4$.
    Dotted lines indicate true values, and limits of the $y$-axes are regions
    of uniform prior density. (A) Estimates for $\alpha$ and $I$ against their
    true values in simulations. (B) Estimates for $m$ and $N$, which were held
    fixed in these simulations, against true values of $\alpha$.
  }
  \label{fig:abcptm4}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics{alpha-gamma.pdf}
  \caption{
      Relationship between preferential attachment power parameter $\alpha$
      and power law exponent $\gamma$ for networks simulated under the BA
      network model with $N$ = 5000 and $m$ = 2.
  }
  \label{fig:gamma}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics{mixed-posterior}
    \caption{
        Approximate marginal posterior distributions of Barab\'asi-Albert
        model parameters obtained using ABC for a network with
        heterogeneous node behaviour. Half of the nodes were attached with
        $\alpha = 0.5$, and the other half with $\alpha = 1.5$ (vertical
        dashed lines, top right). Other parameter values were $m = 2$, $I =
        1000$, and $N = 5000$ (vertical dashed lines, other than top right).
        Shaded areas indicate 95\% highest posterior density intervals.
    }
    \label{fig:mixed}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics{peerdriven-posterior}
    \caption{Approximate marginal posterior distributions of Barab\'asi-Albert
        model parameters obtained using ABC for a network with
        peer-driven sampling. An epidemic was simulated in the usual fashion,
        but rather than being sampled at random, infected nodes were sampled
        with a probability two times higher if they had any sampled neighbours
        in the contact network. Vertical dashed lines indicate true parameter
        values, and shaded areas indicate 95\% highest posterior density
        intervals.
    }
    \label{fig:peerdriven}
\end{figure}

\clearpage

\begin{figure}[ht]
  \includegraphics{realdata-hpd-m2}
  \vspace{8pt}
  \caption{
      Maximum \textit{a posteriori} point estimates and 95\% HPD intervals for
      parameters of the BA network model, fitted to five published HIV datasets
      with ABC. $x$-axes indicate regions of nonzero prior density. In
      particular, the prior on $m$ was DiscreteUniform(2, 5).
  }
  \label{fig:abchpdm2}
\end{figure}

\begin{figure}[ht]
  \includegraphics{cuevas2009-posterior}
  \caption{
      Posterior distribution of BA model parameters for
      \textcite{cuevas2009hiv} data. Vertical lines indicate maximum \textit{a
      posteriori} estimates, and shaded areas are 95\% highest posterior
      density intervals. $x$-axis indicates regions of nonzero prior density.
  }
  \label{fig:cuevas}
\end{figure}

\begin{figure}[ht]
  \includegraphics{li2015-posterior}
  \caption{
      Posterior distribution of BA model parameters for \textcite{li2015hiv}
      data. Vertical lines indicate maximum \textit{a posteriori} estimates,
      and shaded areas are 95\% highest posterior density intervals. $x$-axis
      indicates regions of nonzero prior density.
  }
  \label{fig:li}
\end{figure}

\begin{figure}[ht]
  \includegraphics{niculescu2015-posterior}
  \caption{
      Posterior distribution of BA model parameters for
      \textcite{niculescu2015recent} data. Vertical lines indicate maximum
      \textit{a posteriori} estimates, and shaded areas are 95\% highest
      posterior density intervals. $x$-axis indicates regions of nonzero prior
      density.
  }
  \label{fig:niculescu}
\end{figure}

\begin{figure}[ht]
  \includegraphics{novitsky2014-posterior}
  \caption{
      Posterior distribution of BA model parameters for
      \textcite{novitsky2013phylogenetic, novitsky2014impact} data. Vertical
      lines indicate maximum \textit{a posteriori} estimates, and shaded areas
      are 95\% highest posterior density intervals. $x$-axis indicates regions
      of nonzero prior density.
  }
  \label{fig:novitsky}
\end{figure}

\begin{figure}[ht]
  \includegraphics{wang2015-posterior}
  \caption{
      Posterior distribution of BA model parameters for
      \textcite{wang2015targeting} data. Vertical lines indicate maximum
      \textit{a posteriori} estimates, and shaded areas are 95\% highest
      posterior density intervals. $x$-axis indicates regions of nonzero prior
      density.
  }
  \label{fig:wang}
\end{figure}

<<abc_posterior, echo=FALSE, results="asis", eval=FALSE>>=
    N <- 5000
    replicate <- 0
    for (m in c(2, 3, 4)) {
    for (I in c(1000, 2000)) {
    for (alpha in c(0, 0.5, 1, 1.5)) {
        fn <- sprintf("{abc-posterior/%.1f_%d_%d_%d_%d}.pdf", alpha, I, m, N, replicate)
        lab <- sprintf("fig:%.1f-%d-%d-%d-%d", alpha, I, m, N, replicate)
        cat("\\begin{figure}[ht]\n")
        cat("\\centering\n")
        cat(paste0("\\includegraphics{", fn, "}\n"))
        cat("\\caption{\n")
        cat(sprintf("Approximate marginal posterior distributions of Barab\\'asi-Albert model parameters obtained using ABC for one simulated transmission tree with parameter values $\\alpha = %.1f$, $I = %d$, $m = %d$, and $N = %d$. Dashed lines indicate true values, and shaded areas show 95\\%% highest posterior density intervals.\n", alpha, I, m, N))
        cat("}\n")
        cat(paste0("\\label{", lab, "}\n"))
        cat("\\end{figure}\n")
        cat("\\clearpage\n\n")
    }
    }
    }
@

\end{document}