### 2016-06-27 doc/figures/transtree.tex

\documentclass[landscape]{article}
\usepackage{tikz}
\usepackage[T1]{fontenc}
\usepackage{mathptmx}
\usepackage{fullpage}

\usetikzlibrary{positioning}
\usetikzlibrary{calc}
\usetikzlibrary{arrows}
\usetikzlibrary{automata}

\graphicspath{{stock/}}

\begin{document}
\pagestyle{empty}

\begin{tikzpicture}
  [every node/.style = {inner sep=0pt},
   every path/.style = {->, >=stealth, very thick},
   label/.style = {anchor=south, color=white, inner sep=6pt},
   time/.style = {anchor=north west, inner sep=2pt}]
  \def\pht{1.5cm}
  \def\pd{0.5cm}
  \def\nd{1cm}

  \node (abc) {\includegraphics[height=\pht]{person}};
  \node at (abc.south) [label] {\huge a};

  \node (ac) [below left=\pd of abc] {\includegraphics[height=\pht]{person}};
  \node at (ac.south) [label] {\huge a};

  \node (a) [below left=\pd of ac] {\includegraphics[height=\pht]{person}};
  \node at (a.south) [label] {\huge a};
  \node (b) [below right=\pd of abc] {\includegraphics[height=\pht]{person}};
  \node at (b.south) [label] {\huge b};
  \node (c) [below right=\pd of ac] {\includegraphics[height=\pht]{person}};
  \node at (c.south) [label] {\huge c};

  \draw [red] (abc) -- (b);
  \draw [-, dashed] (abc) -- (ac);
  \draw [blue] (ac) -- (c);
  \draw [-, dashed] (ac) -- (a);

  %\node (a) [left=7cm of abc] {\includegraphics[height=\pht]{person}};
  %\node at (a.south) [label] {\huge a};
  %\node (b) [below left=\nd of a] {\includegraphics[height=\pht]{person}};
  %\node at (b.south) [label] {\huge b};
  %\node (c) [below right=\nd of a] {\includegraphics[height=\pht]{person}};
  %\node at (c.south) [label] {\huge c};

  %\draw [red] (a) -- (b);
  %\draw [blue] (a) -- (c);

  \node [below=6cm of abc] (abc) {\includegraphics[height=\pht]{person}};
  \node at (abc.south) [label] {\huge b};

  \node (ac) [below left=\pd of abc] {\includegraphics[height=\pht]{person}};
  \node at (ac.south) [label] {\huge c};

  \node (a) [below left=\pd of ac] {\includegraphics[height=\pht]{person}};
  \node at (a.south) [label] {\huge a};
  \node (b) [below right=\pd of abc] {\includegraphics[height=\pht]{person}};
  \node at (b.south) [label] {\huge b};
  \node (c) [below right=\pd of ac] {\includegraphics[height=\pht]{person}};
  \node at (c.south) [label] {\huge c};

  \draw [-, dashed] (abc) -- (b);
  \draw [green!60!black] (abc) -- (ac);
  \draw [-, dashed] (ac) -- (c);
  \draw [cyan] (ac) -- (a);

  %\node (a) [left=7cm of abc] {\includegraphics[height=\pht]{person}};
  %\node at (a.south) [label] {\huge a};
  %\node (b) [below left=\nd of a] {\includegraphics[height=\pht]{person}};
  %\node at (b.south) [label] {\huge b};
  %\node (c) [below right=\nd of a] {\includegraphics[height=\pht]{person}};
  %\node at (c.south) [label] {\huge c};

  %\draw [green!60!black] (b) -- (c);
  %\draw [cyan] (c) -- (a);

  \node [above left=2cm and 8cm of abc] (abc) {\includegraphics[height=\pht]{person}};
  \node (ac) [below left=\pd of abc] {\includegraphics[height=\pht]{person}};

  \node (a) [below left=\pd of ac] {\includegraphics[height=\pht]{person}};
  \node at (a.south) [label] {\huge a};
  \node (b) [below right=\pd of abc] {\includegraphics[height=\pht]{person}};
  \node at (b.south) [label] {\huge b};
  \node (c) [below right=\pd of ac] {\includegraphics[height=\pht]{person}};
  \node at (c.south) [label] {\huge c};

  \draw [-] (abc) -- (b);
  \draw [-] (abc) -- (ac);
  \draw [-] (ac) -- (c);
  \draw [-] (ac) -- (a);

  %\node (a) [left=7cm of abc] {\includegraphics[height=\pht]{person}};
  %\node at (a.south) [label] {\huge a};
  %\node (b) [below left=\nd of a] {\includegraphics[height=\pht]{person}};
  %\node at (b.south) [label] {\huge b};
  %\node (c) [below right=\nd of a] {\includegraphics[height=\pht]{person}};
  %\node at (c.south) [label] {\huge c};

\end{tikzpicture}
\end{document}