### 2016-06-27 doc/tables/realdata-bc-short.tex

\begin{tabular}{ccc}
  Dataset & Sequences ($n$) & Gene \\
  \hline
  IDU/Estonia~\autocite{zetterberg2004two} & 171/188 & \textit{env}/\textit{gag} \\
  IDU/Romania~\autocite{niculescu2015recent} & 136 & \textit{pol} \\
  IDU/Canada & 399 & \textit{pol} \\
  HET/Botswana~\autocite{novitsky2013phylogenetic,novitsky2014impact} & 180 & \textit{env} \\
  HET/Malawi~\autocite{mccormack2002early} & 141/154 & \textit{env}/\textit{gag} \\
  HET/Uganda~\autocite{grabowski2014role} & 225 & \textit{env}/\textit{gag} \\
  MSM/Beijing~\autocite{wang2015targeting} & 173 & \textit{pol} \\
  MSM/Taiwan~\autocite{kao2011surveillance} & 275 & \textit{pol} \\
  MSM/USA~\autocite{little2014using} & 180 & \textit{pol} \\
  MSM/Shanghai~\autocite{li2015hiv} & 280 & \textit{pol} \\
  mixed/Spain~\autocite{cuevas2009hiv} & 287 & \textit{pol} \\
  \hline
\end{tabular}