### 2016-06-27 doc/tables/realdata.tex

\begin{tabular}{ccccc}
  Reference & Sequences ($n$) & Location & Risk group & Gene \\
  \hline
  \textcite{zetterberg2004two} & 171/188 & Estonia & IDU & \textit{env}/\textit{gag} \\
  \textcite{niculescu2015recent} & 136 & Romania & IDU & \textit{pol} \\
  \textcite{novitsky2013phylogenetic} & \multirow{2}{*}{180} &
  \multirow{2}{*}{Mochudi, Botswana} & \multirow{2}{*}{HET} &
  \multirow{2}{*}{\textit{env}} \\ \textcite{novitsky2014impact} \\
  \textcite{mccormack2002early} & 141/154 & Karonga District, Malawi & HET & \textit{env}/\textit{gag} \\
  \textcite{grabowski2014role} & 225 & Rakai District, Uganda & HET & \textit{env}/\textit{gag} \\
  \textcite{wang2015targeting} & 173 & Beijing, China & MSM & \textit{pol} \\
  \textcite{kao2011surveillance} & 275 & Taiwan & MSM & \textit{pol} \\
  \textcite{little2014using} & 180 & San Fransisco, USA & MSM & \textit{pol} \\
  \textcite{li2015hiv} & 280 & Shanghai, China & MSM & \textit{pol} \\
  \textcite{cuevas2009hiv} & 287 & Basque Country, Spain & mixed & \textit{pol} \\
  \hline
\end{tabular}