### 2016-06-29 doc/defence/contactnet.tex
working on defence talk

\begin{frame}{Contact networks represent possible transmission routes}
  \begin{center}
  \begin{tikzpicture}
    \node (c) {\includegraphics[width=4cm]{contactnet-empty}};

    \uncover<2->{
    \node (a) [right=of c.north east, anchor=west] {nodes = individuals};
    \draw [->, thick, >=stealth] (c.north east) ++(-1.5cm, -0.5cm) -- (a.west);
    }

    \uncover<3->{
    \node (b) [below=0.5 of a] {edges = contacts};
    \draw [->, thick, >=stealth] (c.north east) ++(-0.75cm, -1.25cm) -- (b.west);
    }
  \end{tikzpicture}
  \end{center}
\end{frame}