### 2016-06-29 doc/defence/inference.tex

\begin{frame}{Can we fit network models from viral sequence data?}
    \begin{tikzpicture}
      [
        every path/.style={very thick, ->, >=stealth}
      ]
      \uncover<2->{
        \node at (0, 0) [anchor=center] {\includegraphics[width=\textwidth]{objects}};
        \draw (-3.5, 2) to [bend left=75] (-1.5, 2);
      }
      \only<2>{ 
        \fill [white] (0, 2) rectangle (6, -2); 
      }
      \only<3>{
        \fill [white] (2.8, 2) rectangle (6, -2); 
      }
      \uncover<3->{
        \draw (-1, 2) to [bend left=75] (1, 2);
      }

      \uncover<4->{
        \draw (1.5, 2) to [bend left=75] (3.5, 2);
      }

      \uncover<5->{
        \draw [green] (3.5, -2) to [bend left=75] node [auto, green] {\Large \checkmark} (1.5, -2);
      }

      \uncover<6->{
        \draw [green] (1, -2) to [bend left=75] node [auto, green] {\Large ``\checkmark''} (-1, -2);
      }
      \uncover<7->{
        \draw [orange] (-1.5, -2) to [bend left=75] node [auto] {\Large \alert{?}} (-3.5, -2);
      }
    \end{tikzpicture}
\end{frame}