### 2016-06-29 doc/defence/virusphylo.tex

\begin{frame}{Transmission trees shape viral phylogenies}
  \begin{center}
    \begin{tikzpicture}[
        every node/.style = {inner sep=0pt},
        every path/.style = {->, >=stealth, very thick},
        label/.style = {anchor=south, color=white, inner sep=4pt}
      ]
      \def\pht{1cm}
      \node (abcd) {\includegraphics[height=\pht]{stock/person}};
      \node at (abcd.south) [label] {\Large a};

      \uncover<2>{
      \node (b) at (abcd.east) [anchor=west] {\includegraphics[height=\pht]{stock/person}};
      \node at (b.south) [label] {\Large b};
      \draw [red] (abcd.south) to [bend right=75] (b.south);
      }

      \uncover<3->{
      \node (b) [below right=1 and 0.75 of abcd] {\includegraphics[height=\pht]{stock/person}};
      \node at (b.south) [label] {\Large b};
      \node (acd) [below left=1 and 0.5 of abcd] {\includegraphics[height=\pht]{stock/person}};
      \node at (acd.south) [label] {\Large a};
      \draw [red] (abcd) -- (b);
      \draw [-, dashed] (abcd) -- (acd);
      }

      \uncover<4>{
      \node (cd) at (acd.east) [anchor=west] {\includegraphics[height=\pht]{stock/person}};
      \node at (cd.south) [label] {\Large c};
      \draw [blue] (acd.south) to [bend right=75] (cd.south);
      }

      \uncover<5->{
      \node (cd) [below right=0.5 and 0.5 of acd] {\includegraphics[height=\pht]{stock/person}};
      \node at (cd.south) [label] {\Large c};
      \node (a) [below left=0.25 and 0.25 of acd] {\includegraphics[height=\pht]{stock/person}};
      \node at (a.south) [label] {\Large a};
      \draw [blue] (acd) -- (cd);
      \draw [-, dashed] (acd) -- (a);
      }

      \uncover<6>{
      \node (d) at (cd.east) [anchor=west] {\includegraphics[height=\pht]{stock/person}};
      \node at (d.south) [label] {\Large d};
      \draw [green] (cd.south) to [bend right=75] (d.south);
      }

      \uncover<7->{
      \node (c) [below left=0.25 and 0.25 of cd] {\includegraphics[height=\pht]{stock/person}};
      \node at (c.south) [label] {\Large c};
      \node (d) [below right=0.4 and 0.4 of cd] {\includegraphics[height=\pht]{stock/person}};
      \node at (d.south) [label] {\Large d};
      \draw [-, dashed] (cd) -- (c);
      \draw [green!80!black] (cd) -- (d);
      }

      \coordinate [below right=0.5 and 3 of b] (abcd);
      \coordinate [below right=1 and 1 of abcd] (b);
      \coordinate [above left=1 and 0.25 of abcd] (acd);
      \coordinate [above left=1 and 0.25 of acd] (a);
      \coordinate [right=of acd] (cd);
      \coordinate [right=1.5 of cd] (c);
      \coordinate [above right=of cd] (d);

      \uncover<1-2>{
      \node (x) at (abcd) {\includegraphics[height=\pht]{stock/virus}};
      \node at (x.center) {\large a};
      }

      \uncover<2>{
      \node (x) [below right=0.3 of abcd, anchor=north west] {\includegraphics[height=\pht]{stock/virus}}; 
      \node at (x.center) {\large b};
      }

      \uncover<3-4>{
      \node (x) [above=0 of acd, anchor=south] {\includegraphics[height=\pht]{stock/virus}}; 
      \node at (x.center) {\large a};
      }

      \uncover<3->{
      \node (x) [below right=-0.2 of b, anchor=north west] {\includegraphics[height=\pht]{stock/virus}}; 
      \node at (x.center) {\large b};

      \draw (abcd) -- (acd);
      \draw (abcd) -- (b);
      }

      \uncover<4>{
      \node (x) [right=0.3 of acd, anchor=west] {\includegraphics[height=\pht]{stock/virus}}; 
      \node at (x.center) {\large c};
      }

      \uncover<5-6>{
      \node (x) at (cd) [anchor=west] {\includegraphics[height=\pht]{stock/virus}};  
      \node at (x.center) {\large c};
      }

      \uncover<5->{
      \node (x) at (a) [anchor=south] {\includegraphics[height=\pht]{stock/virus}};
      \node at (x.center) {\large a};

      \draw (acd) -- (cd);
      \draw (acd) -- (a);
      }

      \uncover<6>{
      \node (x) [above right=0.4\pht and 0.4\pht of cd] [anchor=south west] {\includegraphics[height=\pht]{stock/virus}};
      \node at (x.center) {\large d};
      }

      \uncover<7->{
      \node (x) at (c) [anchor=west] {\includegraphics[height=\pht]{stock/virus}};
      \node at (x.center) {\large c};

      \node (x) at (d) [anchor=south west] {\includegraphics[height=\pht]{stock/virus}};
      \node at (x.center) {\large d};

      \draw (cd) -- (c);
      \draw (cd) -- (d);
      }
    \end{tikzpicture}
  \end{center}
\end{frame}