### 2016-07-03 doc/defence/kpca.tex
Updates to presentation

\begin{frame}{yes they are}
  \centerline{\includegraphics[height=0.8\textheight]{kernel-kpca}}
\end{frame}