### 2016-07-05 doc/defence/kpca.tex
Presentation updates

\begin{frame}{$\alpha$, $I$, and $N$ are strongly or weakly identifiable}
  \centerline{\includegraphics[height=0.8\textheight]{kernel-kpca}}
\end{frame}