### 2016-07-05 doc/defence/objects.tex

\begin{frame}{Transmission trees connect contact networks to viral phylogenies}
    \begin{tikzpicture}
      [
        every path/.style={very thick, ->, >=stealth}
      ]
      \uncover<1-2>{
        \node (a) at (0, 0) [anchor=center] {\includegraphics[width=\textwidth, trim=0 0 5cm 0, clip]{objects}};
        \node [below=0 of a] {contact network \qquad transmission tree \qquad viral phylogeny};
      }
      \only<1>{ 
        \fill [white] (-1.6, 2.1) rectangle (2.4, -2.7); 
      }
    \end{tikzpicture}
\end{frame}