### 2016-07-05 doc/defence/phylo.tex

\begin{frame}{Viral phylogenies relate individuals' viral genotypes}
  \begin{center}
    \begin{tikzpicture}[
        every node/.style = {inner sep=0pt},
        every path/.style = {-, very thick},
        label/.style = {anchor=south, color=white, inner sep=4pt}
      ]
      \def\pht{1cm}

      \coordinate (abcd);
      \coordinate [below right=1 and 1 of abcd] (b);
      \coordinate [above left=1 and 0.25 of abcd] (acd);
      \coordinate [above left=1 and 0.25 of acd] (a);
      \coordinate [right=of acd] (cd);
      \coordinate [right=1.5 of cd] (c);
      \coordinate [above right=of cd] (d);

      \node (x) [right=0.5 of b, anchor=west, opacity=0.4] {\includegraphics[height=\pht]{stock/person}}; 
      \node (v) at (x.south west) {\includegraphics[height=\pht]{stock/virus}}; 
      \node at (v.center) {\large b};

      \draw (abcd) -- (acd);
      \draw (abcd) -- (b);

      \node (x) [above=0.7 of a, anchor=south west, opacity=0.4] {\includegraphics[height=\pht]{stock/person}};
      \node (va) at (x.south west) {\includegraphics[height=\pht]{stock/virus}};
      \node at (va.center) {\large a};

      \draw (acd) -- (cd);
      \draw (acd) -- (a);

      \node (x) [above right=0 and 0.8 of c, opacity=0.4] {\includegraphics[height=\pht]{stock/person}};
      \node (v) at (x.south west) {\includegraphics[height=\pht]{stock/virus}};
      \node at (v.center) {\large c};

      \node (x) [above right=0.8 of d, opacity=0.4] {\includegraphics[height=\pht]{stock/person}};
      \node (v) at (x.south west) {\includegraphics[height=\pht]{stock/virus}};
      \node at (v.center) {\large d};

      \draw (cd) -- (c);
      \draw (cd) -- (d);

      \uncover<2->{
        \node (t1) [below left=of a, inner ysep=6pt] {viruses within individuals};
        \draw [->, >=stealth] (t1.north) -- (va.south west);
      }

      \uncover<3->{
        \node (t2) [below=of t1, inner ysep=6pt] {evolutionary relationships};
        \coordinate [left=0.5 of abcd] (tmp);
        \draw [->, >=stealth] (t2.north) -- (tmp);
      }

    \end{tikzpicture}
  \end{center}
\end{frame}