### 2016-07-05 doc/defence/pointest.tex

\begin{frame}{$\alpha$ and $I$ are identifiable with ABC}
    \includegraphics[width=\textwidth]{abc-boxplot}
\end{frame}