### 2016-07-05 doc/defence/posterior1.tex

\begin{frame}{Marginal posterior distributions for $\alpha$ show least spread}
  \begin{minipage}[p][\textheight][t]{\textwidth}
    \begin{center}
    preferential attachment power \hfill prevalence
      \includegraphics[width=0.8\textwidth]{abc-posterior-example.pdf}

      \vspace{-0.25cm}
      edges per vertex \hfill total nodes
    \end{center}
  \end{minipage}
\end{frame}