### 2016-07-06 doc/paper/abstract.tex
Add paper and tex

Models of the spread of disease in a population often make the simplifying
assumption that the population is homogeneously mixed, or is divided into
homogeneously mixed compartments. However, human populations have complex
structures formed by social contacts, which can have a significant influence on
the rate of epidemic spread. Contact network models capture this structure by
explicitly representing each contact which could possibly lead to a
transmission. We developed a method based on approximate Bayesian
computation (ABC) for estimating structural parameters of the contact network
underlying an observed viral phylogeny. The method combines adaptive sequential
Monte Carlo for ABC, Gillespie simulation for propagating epidemics though
networks, and a kernel-based tree similarity score. We used the method to fit
the Barab\'{a}si-Albert network model to simulated transmission trees, and also
applied it to viral phylogenies estimated from \del{five}\add{ten} published
HIV sequence datasets. On simulated data, we found that the preferential
attachment power and the number of infected nodes in the network can often be
accurately estimated. On the other hand, the mean degree of the network, as
well as the total number of nodes, were not estimable with ABC. \del{We
observed substantial heterogeneity in the parameter estimates on real datasets,
with point estimates for the preferential attachment power ranging from 0.06 to
1.05.} \add{We observed sub-linear preferential attachment (PA) power in all
datasets, as well as higher PA power in networks of injection drug users.}
These results underscore the importance of considering contact structures when
performing phylodynamic inference. Our method offers the potential to
quantitatively investigate the contact network structure underlying viral
epidemics.