### 2016-07-14 doc/defence/abc.tex
Presentation

\begin{frame}{Simulated trees can be used to fit network models}
  \begin{minipage}[p][0.8\textheight][t]{\textwidth}
    \only<2>{\includegraphics[width=0.9\textwidth, trim=0 3in 0 0, clip]{abc-idea.pdf}}
    \only<3>{\includegraphics[width=0.9\textwidth, trim=0 1.7in 0 0, clip]{abc-idea.pdf}}
    \only<4>{\includegraphics[width=0.9\textwidth, trim=0 0.8in 0 0, clip]{abc-idea.pdf}}
    \only<5->{\includegraphics[width=0.9\textwidth, trim=0 0in 0 0, clip]{abc-idea.pdf}}

    \vspace{0.5cm}
    \only<6>{\centerline{\Large approximate Bayesian computation}}
  \end{minipage}
\end{frame}