### 2016-07-14 doc/defence/conclusions.tex

\begin{frame}
  \begin{itemize}
    \setlength{\itemsep}{12pt}
    \item \textit{Netabc} uses approximate Bayesian computation with sequential
      Monte Carlo, epidemic simulation, and the tree kernel to fit contact
      network models to phylogenetic data.
    \pause
    \item In simulated data, viral phylogenies were informative about the
      $\alpha$, $I$, and $N$ parameters of the Barab\'asi-Albert contact
      network model.
    \pause
    \item The networks underlying real-world HIV epidemics exhibit sub-linear
      preferential attachment, and PA values may be higher for IDU networks.
  \end{itemize}
\end{frame}