### 2016-07-14 doc/defence/identifiable.tex

\begin{frame}{Are network parameters identifiable from tree shape?}
  \begin{minipage}[p][0.6\textheight][t]{\textwidth}
    \only<2>{\includegraphics[width=0.9\textwidth, trim=0 2.8in 0 0, clip]{kernel-idea.pdf}}
    \only<3>{\includegraphics[width=0.9\textwidth, trim=0 1.7in 0 0, clip]{kernel-idea.pdf}}
    \only<4>{\includegraphics[width=0.9\textwidth, trim=0 0.8in 0 0, clip]{kernel-idea.pdf}}
    \only<5->{\includegraphics[width=0.9\textwidth, trim=0 0in 0 0, clip]{kernel-idea.pdf}}
  \end{minipage}
\end{frame}