### 2016-07-14 doc/defence/kpca.tex

\begin{frame}{$\alpha$, $I$, and $N$ are strongly or weakly identifiable; $m$ is not}
  \begin{center}
  \vspace{-0.5cm}
  \begin{tikzpicture}
    \node at (0, 0) [anchor=center] {\includegraphics[height=0.8\textheight, trim=0.75cm 0.75cm 0 0, clip]{kernel-kpca}};
    \only<1>{ \fill [white] (0, 0) rectangle (4.3, 3.6); }
    \only<1-2> { \fill [white] (0, 0) rectangle (-4.2, -3.6); }
    \only<1-3> { \fill [white] (0, 0) rectangle (4.3, -3.6); }
    \only<4>{ }
  \end{tikzpicture}
  \end{center}
\end{frame}