### 2016-07-14 doc/defence/pointest.tex

\begin{frame}{$\alpha$ and $I$ are identifiable with ABC}
  \vspace{-0.5cm}
  \begin{tikzpicture}
    \node at (0, 0) [anchor=center] {\includegraphics[width=\textwidth]{abc-boxplot}};
    \only<1> { 
      \fill [white] (-1.2, 0.5) rectangle (2, 3.8);
      \fill [white] (-1.2, -4) rectangle (2, -3);
    }
    \only<1-2> { 
      \fill [white] (2, 0.5) rectangle (6, 3.8); 
      \fill [white] (2, -4) rectangle (6, -3);
    } 
    \only<1-3>{ \fill [white] (-5.5, -3) rectangle (6, 0.5); }
    \only<4>{ }
  \end{tikzpicture}
\end{frame}