### 2016-07-14 doc/defence/posterior1.tex

\begin{frame}{Marginal posterior distributions for $\alpha$ show least spread}
  \vspace{-0.5cm}
  \begin{tikzpicture}
    \node at (0, 0) [anchor=center] {\includegraphics[trim=1cm 0 0 0, clip, width=0.8\textwidth]{{abc-posterior/1.0_1000_3_5000_0}.pdf}};
    \node at (-4, 3) [anchor=center, text width=2cm, fill=white, inner sep=2pt] {preferential attachment};
    \node at (4.5, 3) [anchor=center, text width=2cm, fill=white, inner sep=2pt] {number infected};
    \node at (-4.2, -1) [anchor=center, text width=2cm, align=center, fill=white, inner ysep=2pt, inner xsep=0pt] {edges per vertex};
    \node at (4.2, -0.7) [anchor=center, text width=1.5cm, align=center, fill=white, inner sep=2pt] {total nodes};

    \only<1>{ \fill [white] (0, 0) rectangle (4.8, 3.8);}
    \only<1-2>{ \fill [white] (0, 0) rectangle (-5, -3.8);}
    \only<1-3>{ \fill [white] (0, 0) rectangle (4.8, -3.8);}
    \only<4>{ }
  \end{tikzpicture}
\end{frame}