### 2016-07-14 doc/defence/supp.tex

\begin{frame}{Marginalization improves estimates of $I$, $m$, and $N$}
  \centerline{\includegraphics[height=0.8\textheight]{abc-marginal}}
\end{frame}

\begin{frame}{Higher SMC settings do not improve parameter estimates}
  \centerline{\includegraphics[height=0.8\textheight]{better-smc}}
\end{frame}

\begin{frame}{$I$ and $N$ may be identifiable only up to epidemic phase}
  \centerline{\includegraphics[height=0.8\textheight]{derivatives}}
\end{frame}

\begin{frame}{Grid search estimates are accurate with wide intervals}
  \centerline{\includegraphics[height=0.8\textheight]{gridsearch-example}}
\end{frame}

\begin{frame}{Higher $\alpha$ causes unbalanced trees}
  \includegraphics[width=\textwidth]{kernel-alpha-tree}
\end{frame}

\begin{frame}{Higher $I$ causes longer terminal branch lengths}
  \includegraphics[width=\textwidth]{kernel-I-tree}
\end{frame}

\begin{frame}{Tree shape not visibly affected by $m$}
  \includegraphics[width=\textwidth]{kernel-m-tree}
\end{frame}

\begin{frame}{Higher $N$ causes longer internal branch lengths}
  \includegraphics[width=\textwidth]{kernel-N-tree}
\end{frame}

\begin{frame}{Tree kernel and other statistics for parameter classification}
  \includegraphics[width=\textwidth]{kernel-rsquared}
\end{frame}

\begin{frame}{Estimates of $\alpha$ are robust to HIV gene}
  \centerline{\includegraphics[height=0.8\textheight]{realdata-genes}}
\end{frame}

\begin{frame}{$\alpha$ estimates are robust to phylogeny reconstruction method}
  \centerline{\includegraphics[height=0.8\textheight]{realdata-methods}}
\end{frame}

\begin{frame}{Power law fit captures slope of BA degree distributions}
  \centerline{\includegraphics[height=0.8\textheight]{powerlaw}}
\end{frame}

\begin{frame}{SMC was validated on two mixture distributions}
  \begin{columns}
    \column{0.5\textwidth}
    \includegraphics[width=\textwidth]{smc-test}
    \column{0.5\textwidth}
    \includegraphics[width=\textwidth]{smc-test-bimodal}
  \end{columns}
\end{frame}