### 2016-07-27 doc/thesis/acknowledgements.tex
Defence version of thesis with changes

First and foremost I am deeply indebted to my supervisor, Dr. Art Poon, for
mentoring me through the majority of my academic career, and somehow always
managing to make time for me. My time in his lab has been challenging and
rewarding, and I feel well prepared for opportunities to come. I also thank my
labmates, especially Jeff, Richard, and Thuy. I'll miss our Thursday lunches.

I am grateful for the guidance of my committee members Dr. Alexandre
Bouchard-C\^ot\'e, Dr. Richard Harrigan, and Dr. Sarah Otto. Thanks also to my
rotation supervisors Dr. Sara Mostafavi and Dr. Ryan Morin, who accepted me
into their labs and always had time for my many questions.

I owe thanks to my fellow graduate students of the bioinformatics training
program, especially Ogan, who helped keep me on an even keel when times got
tough.

Finally, I dedicate this thesis to my parents, who have supported me morally
and financially, endured my frustrations, and celebrated my successes.